---
title: "Annotate and assess peak usage"
author: "Briana Mittleman"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(reshape2)
```

I will annotate the PAS that passed [the liftover](liftoverPAS.html). These PAS are in 
../data/cleanPeaks_lifted


Map PAS to these annoations:

```{bash,eval=F}
mkdir ../data/cleanPeaks_anno
bedtools map -a ../data/cleanPeaks_lifted/AllPAS_postLift.sort.bed -b  /project2/gilad/briana/genome_anotation_data/hg38_refseq_anno/hg38_ncbiRefseq_Formatted_Allannotation.sort.bed -c 4 -S -o distinct > ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnno.bed 
```

Chose annotation if PAS in multiple and create bed. I will have to lift this back to chimp then make saf files for both to do the feature count  



```{bash,eval=F}
python chooseAnno2Bed.py ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnno.bed  ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnnoPARSED.bed 
```


Lift this so I have it with chimp coordinates:  

```{bash,eval=F}
sbatch LiftOrthoPAS2chimp.sh
```


bed 2 SAF

```{bash,eval=F}
python bed2SAF_gen.py ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnnoPARSED.bed  ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnnoPARSED.SAF

python bed2SAF_gen.py ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnnoPARSED_chimpLoc.bed  ../data/cleanPeaks_anno/AllPAS_postLift.sort_LocAnnoPARSED_chimpLoc.SAF
```

Use feature counts to quantify:  


```{bash,eval=F}
mkdir  ../Human/data/CleanLiftedPeaks_FC/
mkdir ../Chimp/data/CleanLiftedPeaks_FC/

sbatch quantLiftedPAS.sh 

```

Fix header:  

```{bash,eval=F}
python fixFChead_bothfrac.py ../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human ../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc


python fixFChead_bothfrac.py ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc

#make file ID

python makeFileID.py ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp ../Chimp/data/CleanLiftedPeaks_FC/ChimpFileID.txt

python makeFileID.py ../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human ../Human/data/CleanLiftedPeaks_FC/HumanFileID.txt
```

Make phenotypes from these: 
```{bash,eval=F}
mkdir ../Human/data/phenotype/
mkdir ../Chimp/data/phenotype/
python makePheno.py  ../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc ../Human/data/CleanLiftedPeaks_FC/HumanFileID.txt ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt

python makePheno.py  ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc ../Chimp/data/CleanLiftedPeaks_FC/ChimpFileID.txt ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt
```



Convert these to numeric:  

```{bash,eval=F}
Rscript pheno2countonly.R -I ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt -O ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnly.txt

Rscript pheno2countonly.R -I ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt -O ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnly.txt


python convertNumeric.py ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnly.txt ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt

python convertNumeric.py ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnly.txt ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt
```


Plot usages to see if 5% is a good cutoff for this analysis as well.  

```{r}
HumanAnno=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndH=colnames(HumanAnno)[9:ncol(HumanAnno)]

HumanUsage=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt", col.names = IndH)

HumanMean=as.data.frame(cbind(HumanAnno[,1:8], Human=rowMeans(HumanUsage)))

HumanUsage_anno=as.data.frame(cbind(HumanAnno[,1:8],HumanUsage ))
```

```{r}
ChimpAnno=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndC=colnames(ChimpAnno)[9:ncol(ChimpAnno)]

ChimpUsage=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt", col.names = IndC)

ChimpMean=as.data.frame(cbind(ChimpAnno[,1:8], Chimp=rowMeans(ChimpUsage)))

ChimpUsage_anno=as.data.frame(cbind(ChimpAnno[,1:8],ChimpUsage ))
```


Mean both: 

```{r}
BothMean=ChimpMean %>% full_join(HumanMean, by=c("chr","start","end","gene"   ,"strand", "loc", "disc","PAS" )) 

BothMeanM=melt(BothMean,id.vars =c("chr","start","end","gene"   ,"strand", "loc", "disc","PAS" ),variable.name = "Species", value.name = "MeanUsage" ) %>% filter(loc !="008559")
```

Plot:

```{r}
ggplot(BothMeanM, aes(x=loc, y=MeanUsage,by=Species,fill=Species)) + geom_boxplot()  + scale_fill_brewer(palette = "Dark2")
```


```{r}
ggplot(BothMeanM, aes(x=MeanUsage,by=Species,col=Species)) + stat_ecdf(geom = "point", alpha=.25)  + scale_color_brewer(palette = "Dark2") + labs(title="Cumulative Distribution plot for PAS Usage", x="Mean Usage- both fractions", y="F(Mean Usage)") 

```


This is good. Globally the usages are similar across species.  

```{r}


ggplot(BothMeanM, aes(x=log10(MeanUsage + .00001),by=Species,fill=Species)) + geom_histogram(alpha=.5, bins=30,position="dodge")  + scale_fill_brewer(palette = "Dark2") + geom_vline(xintercept = log10(0.05))

```

Looks like 5% in either species is a good set.  


Filter to PAS with 5% usage  


BothMean mean in human or chimp > 0.5 
```{r} 
BothMean_5= BothMean %>% filter(Chimp >=0.05 | Human >= 0.05)  


```

```{r}
BothMean_5M=melt(BothMean_5,id.vars =c("chr","start","end","gene"   ,"strand", "loc", "disc","PAS" ),variable.name = "Species", value.name = "MeanUsage" ) %>% filter(loc !="008559")

ggplot(BothMean_5M, aes(x=loc, y=MeanUsage,by=Species,fill=Species)) + geom_boxplot()  + scale_fill_brewer(palette = "Dark2")
```

```{r}
ggplot(BothMean_5M, aes(x=MeanUsage,by=Species,col=Species)) + stat_ecdf(geom = "point", alpha=.25)  + scale_color_brewer(palette = "Dark2") + labs(title="Cumulative Distribution plot for PAS Usage at 5%", x="Mean Usage- both fractions", y="F(Mean Usage)") 

```
```{r}
ggplot(BothMean_5M, aes(x=MeanUsage,by=Species,fill=Species)) + geom_histogram(alpha=.5, bins=30, position = "dodge")  + scale_fill_brewer(palette = "Dark2")
```

Write this out this way and as a bed files with human and chimp scores: 


```{bash,eval=F}
mkdir ../data/Peaks_5perc
mkdir ../data/Pheno_5perc
```

```{r}
BothMean_5_out=BothMean_5 %>% dplyr::select(PAS,disc, gene, loc,chr, start, end,Chimp, Human)
write.table(BothMean_5_out, "../data/Peaks_5perc/Peaks_5perc_either_bothUsage.txt", row.names = F, col.names = T, quote = F)

BothMean_5_out_noUN=BothMean_5 %>% dplyr::select(PAS,disc, gene, loc,chr, start, end,Chimp, Human) %>% filter(!grepl("Un",chr))

write.table(BothMean_5_out_noUN, "../data/Peaks_5perc/Peaks_5perc_either_bothUsage_noUnchr.txt", row.names = F, col.names = T, quote = F)
#write bed with human coord for igv
BothMean_5_bed=BothMean_5 %>% dplyr::select(chr, start, end, PAS, Human, strand)
write.table(BothMean_5_bed, "../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed", row.names = F, col.names = T, quote = F)

```

I can filter the phenotypes and PAS with this list. 

```{r}
ggplot(BothMean_5_out, aes(x=disc, fill=disc))+  geom_bar(aes(y = (..count..)/sum(..count..)))+ scale_fill_brewer(palette = "Dark2")


BothMean_5_outmean= BothMean_5_out %>% mutate(meanUsage=(Human+Chimp)/2)
ggplot(BothMean_5M, aes(x=disc, by= Species, fill=Species, y=MeanUsage)) + geom_boxplot() + scale_y_log10()+ scale_fill_brewer(palette = "Dark2")


```



```{r}


ChimpUsage_anno_5perc= ChimpUsage_anno %>% filter(PAS %in% BothMean_5$PAS)

write.table(ChimpUsage_anno_5perc, "../data/Pheno_5perc/Chimp_Pheno_5perc.txt", row.names = F, col.names = T, quote = F)

HumaUsage_anno_5perc= HumanUsage_anno %>% filter(PAS %in% BothMean_5$PAS)

write.table(HumaUsage_anno_5perc, "../data/Pheno_5perc/Human_Pheno_5perc.txt", row.names = F, col.names = T, quote = F)

```


