---
title: "Human Total vs Nuclear"
author: "Briana Mittleman"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(reshape2)
```
In this anaylsis I will complete the Chimp total vs nuclear analysis. This analysis is similar to the analysis in the [apaQTL project](https://brimittleman.github.io/apaQTL/DiffIsoAnalysis.html)

I need the human 5% both fraction feature counts. 

Thee feature counts are in  ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc. I need to subset these for those in the annotations.  Keep the PAS in this file: ../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed. I will do this with a python script.  

```{bash,eval=F}
mkdir ../Chimp/data/CleanLiftedPeaks4LC/
python prepareCleanLiftedFC_5perc4LC.py ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc ../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed ../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc
```


This will only look at PAS on chromosomes 1-22 no extra haplotpyes. 

```{bash,eval=F}
mkdir ../Chimp/data/DiffIso_Chimp/

python subset_diffisopheno_Chimp_tvN.py 1
python subset_diffisopheno_Chimp_tvN.py 2
python subset_diffisopheno_Chimp_tvN.py 3
python subset_diffisopheno_Chimp_tvN.py 4
python subset_diffisopheno_Chimp_tvN.py 5
python subset_diffisopheno_Chimp_tvN.py 6
python subset_diffisopheno_Chimp_tvN.py 7
python subset_diffisopheno_Chimp_tvN.py 8
python subset_diffisopheno_Chimp_tvN.py 9
python subset_diffisopheno_Chimp_tvN.py 10
python subset_diffisopheno_Chimp_tvN.py 11
python subset_diffisopheno_Chimp_tvN.py 12
python subset_diffisopheno_Chimp_tvN.py 13
python subset_diffisopheno_Chimp_tvN.py 14
python subset_diffisopheno_Chimp_tvN.py 15
python subset_diffisopheno_Chimp_tvN.py 16
python subset_diffisopheno_Chimp_tvN.py 18
python subset_diffisopheno_Chimp_tvN.py 19
python subset_diffisopheno_Chimp_tvN.py 20
python subset_diffisopheno_Chimp_tvN.py 21
python subset_diffisopheno_Chimp_tvN.py 22
```


Make sample groups: 

```{bash,eval=F}
python makeSamplyGroupsChimp_TvN.py
```


Run Leafcutter:

```{bash,eval=F}
sbatch runChimpDiffIso.sh
```


Concatinate results:  

```{bash,eval=F}
awk '{if(NR>1)print}' ../Chimp/data/DiffIso_Chimp/TN_diff_isoform_chr*.txt_effect_sizes.txt > ../Chimp/data/DiffIso_Chimp/TN_diff_isoform_allChrom.txt_effect_sizes.txt


awk '{if(NR>1)print}' ../Chimp/data/DiffIso_Chimp/TN_diff_isoform_chr*.txt_cluster_significance.txt > ../Chimp/data/DiffIso_Chimp/TN_diff_isoform_allChrom.txt_significance.txt

```


Significant clusters:  

```{r}
sig=read.table("../Chimp/data/DiffIso_Chimp/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sig$p.adjust=as.numeric(as.character(sig$p.adjust))
```

```{r}
qqplot(-log10(runif(nrow(sig))), -log10(sig$p.adjust),ylab="-log10 Total Adjusted Leafcutter pvalue", xlab="-log 10 Uniform expectation", main="Chimp: Leafcutter differencial isoform analysis between fractions")
abline(0,1)
```
```{r}
tested_genes=nrow(sig)
tested_genes
```

```{r}
sig_genes=sig %>% filter(p.adjust<.05)
number_sig_genes=nrow(sig_genes)
number_sig_genes
```

Effect Sizes  

```{r}
effectsize=read.table("../Chimp/data/DiffIso_Chimp/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('intron',  'logef' ,'Nuclear', 'Total','deltaPAU')) %>% filter(intron != "intron")

effectsize$deltaPAU=as.numeric(as.character(effectsize$deltaPAU))
effectsize$logef=as.numeric(as.character(effectsize$logef))
```


```{r}
plot(sort(effectsize$deltaPAU),main="Chimp: Leafcutter delta PAU", ylab="Delta PAU", xlab="PAS Index")
```
```{r}
effectsize_deltaPAU= effectsize %>% filter(abs(deltaPAU) > .2) 
nrow(effectsize_deltaPAU)
```

```{r}
effectSize_highdiffGenes=effectsize_deltaPAU %>% separate(intron, into=c("chrom", "start", "end", "GeneName"), sep=":") %>% dplyr::select(GeneName) %>% unique()


```

```{r}
effectsize_deltaPAU_Genes= effectsize_deltaPAU %>% separate(intron, into=c("chrom", "start", "end","gene"),sep=":") %>% group_by(gene) %>% summarise(nperGene=n()) 

nrow(effectsize_deltaPAU_Genes)
```

Pull in the PAS to see where these PAS are located.  

```{r}
PAS=read.table("../data/Peaks_5perc/Peaks_5perc_either_bothUsage.txt", stringsAsFactors = F, header = T) %>% filter(loc !="008559")

PAS$start=as.character(PAS$start)
PAS$end=as.character(PAS$end)
```


Increased usage in nuclear:

```{r}
effectsize_deltaPAU_nuclear= effectsize_deltaPAU %>% filter(deltaPAU < -0.2) %>% separate(intron,into=c("chr", "start", "end", "gene"), sep=":" )%>% inner_join(PAS, by=c("gene", "chr","start", "end"))

```

```{r}
ggplot(effectsize_deltaPAU_nuclear,aes(x=loc,fill=loc)) + geom_histogram(stat="count") + scale_fill_brewer(palette = "Dark2") +labs(title="Chimp: Location of PAS used more in Nuclear")
```


Increased usage in total:  

```{r}
effectsize_deltaPAU_total= effectsize_deltaPAU %>% filter(deltaPAU > 0.2)  %>% separate(intron,into=c("chr", "start", "end", "gene"), sep=":" )%>% inner_join(PAS, by=c("gene", "chr","start", "end"))
```

```{r}
ggplot(effectsize_deltaPAU_total,aes(x=loc,fill=loc)) + geom_histogram(stat="count") + scale_fill_brewer(palette = "Dark2") +labs(title="Chimp: Location of PAS used more in Total")
```



This is in line with expectation. Now I can compare this to the number of discovered PAS by plotting the proportion of PAS used more in each fraction by location. 

```{r}
PAS_loc =PAS%>% group_by(loc) %>% summarise(nloc=n())
```

```{r}
effectsize_deltaPAU_nuclear_loc=effectsize_deltaPAU_nuclear %>% group_by(loc) %>% summarise(nlocNuclear=n()) 

effectsize_deltaPAU_total_loc=effectsize_deltaPAU_total %>% group_by(loc) %>% summarise(nloctotal=n()) 

effectsize_deltaPAUProp_tot=effectsize_deltaPAU_total_loc %>% inner_join(PAS_loc, by="loc") %>% mutate(Proportion_tot=nloctotal/nloc)

effectsize_deltaPAUProp_nuc=effectsize_deltaPAU_nuclear_loc %>% inner_join(PAS_loc, by="loc") %>% mutate(Proportion_nuc=nlocNuclear/nloc)


effectsize_deltaPAUProp_both= effectsize_deltaPAUProp_nuc %>% inner_join(effectsize_deltaPAUProp_tot, by=c("loc","nloc")) %>% dplyr::rename(Nuclear=Proportion_nuc, Total=Proportion_tot) %>% select(loc, Nuclear, Total) 
effectsize_deltaPAUProp_both_melt= effectsize_deltaPAUProp_both %>% melt(id.vars="loc", variable.name="Fraction", value.name = "Proportion") 
effectsize_deltaPAUProp_both_melt$Fraction=as.character(effectsize_deltaPAUProp_both_melt$Fraction)
```

Plot this:  
```{r}
ggplot(effectsize_deltaPAUProp_both_melt, aes(x=loc, y=Proportion, by=Fraction, fill=Fraction)) + geom_bar(stat="identity", position="dodge") + scale_fill_brewer(palette = "Dark2")+ labs(title="Chimp: Proportion of PAS differential used by location",x="") +scale_x_discrete(labels = c('Coding','5kb downstream','Intronic',"3' UTR", "5' UTR")) 
```
