---
title: "Differential APA in total and nuclear"
author: "Briana Mittleman"
date: "12/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I ran a differential isoform analysis to compare human and chimp PAS usage in both total and nuclear fractions. 

```{r}
library(workflowr)
library(tidyverse)
library(VennDiagram)
```

###Chimp  

Are the total and nuclear genes the same: 

```{r}
ChimpTotal=read.table("../data/DiffIso_Total/SignifianceChimpPAS_2_Total.txt", header = T, stringsAsFactors = F)
ChimpNuclear=read.table("../data/DiffIso_Nuclear/SignifianceChimpPAS_2_Nuclear.txt", header=T, stringsAsFactors = F)

```



```{r}
overlap <- intersect(ChimpTotal$gene, ChimpNuclear$gene) 
Total.unique <- setdiff(ChimpTotal$gene, ChimpNuclear$gene) # n 1st NOT 2nd
Nuclear.unique <- setdiff(ChimpNuclear$gene, ChimpTotal$gene)
full.list <- unique(c(ChimpTotal$gene,ChimpNuclear$gene))
```

Plot: 

```{r}
grid.newpage()
venn.plot.chimp <- draw.pairwise.venn(area1 = length(ChimpTotal$gene),
                           area2 = length(ChimpNuclear$gene),
                           cross.area = length(overlap),
                           c("Chimp Total", "Chimp Nuclear"), scaled = TRUE,
                           fill = c("mediumorchid", "orange"),
                           cex = 1,
                           cat.cex = 1,
                           cat.pos = c(320, 25),
                           cat.dist = .05,
                           lty = "blank",
                           alpha = rep(0.5, 2),
                           fontface =rep("plain", 3)) 
```



###Human  

Are the total and nuclear genes the same: 

```{r}
HumanTotal=read.table("../data/DiffIso_Total/SignifianceHumanPAS_2_Total.txt", header = T, stringsAsFactors = F)
HumanNuclear=read.table("../data/DiffIso_Nuclear/SignifianceHumanPAS_2_Nuclear.txt", header=T, stringsAsFactors = F)

```



```{r}
overlapH <- intersect(HumanTotal$gene, HumanNuclear$gene) 
Total.uniqueH <- setdiff(HumanTotal$gene, HumanNuclear$gene) # n 1st NOT 2nd
Nuclear.uniqueH <- setdiff(HumanNuclear$gene, HumanTotal$gene)
full.listH <- unique(c(HumanTotal$gene,HumanNuclear$gene))
```

Plot: 

```{r}
grid.newpage()
venn.plot.Human <- draw.pairwise.venn(area1 = length(HumanTotal$gene),
                           area2 = length(HumanNuclear$gene),
                           cross.area = length(overlapH),
                           c("Human Total", "Human Nuclear"), scaled = TRUE,
                           fill = c("skyblue", "pink1"),
                           cex = 1,
                           cat.cex = 1,
                           cat.pos = c(320, 25),
                           cat.dist = .05,
                           lty = "blank",
                            alpha = rep(0.5, 2),
                           fontface =rep("plain", 3))
```


More in  human total and more in chimp nuclear 
```{r}

overlapHTCN <- intersect(HumanTotal$gene,ChimpNuclear$gene) 
grid.newpage()
venn.plot.HTCN <- draw.pairwise.venn(area1 = length(HumanTotal$gene),
                           area2 = length(ChimpNuclear$gene),
                           cross.area = length(overlapHTCN),
                           c("Human Total", "Chimp Nuclear" ), scaled = TRUE,
                           fill = c("skyblue", "orange"),
                           cex = 1,
                           cat.cex = 1,
                           cat.pos = c(320, 25),
                           cat.dist = .05,
                           lty = "blank",
                            alpha = rep(0.5, 2),
                           fontface =rep("plain", 3))
```

More in  human nuclear and more in chimp total 
```{r}

overlapHNCT <- intersect(HumanNuclear$gene,ChimpTotal$gene) 
grid.newpage()
venn.plot.HNCT <- draw.pairwise.venn(area1 = length(HumanNuclear$gene),
                           area2 = length(ChimpTotal$gene),
                           cross.area = length(overlapHNCT),
                           c("Human Nuclear", "Chimp Total" ), scaled = TRUE,
                           fill = c("pink1", "mediumorchid"),
                           cex = 1,
                           cat.cex = 1,
                           cat.pos = c(320, 25),
                           cat.dist = .05,
                           lty = "blank",
                            alpha = rep(0.5, 2),
                           fontface =rep("plain", 3)) 
```


Plot with all:  

**draw.quad.venn**

area1=HT
area2=HN
area3=CT
area4=CN


Reduce(intersect,  list(v1 = c("a","b","c","d"), 
+                         v2 = c("a","b","e"), 
+                         v3 = c("a","f","g"))) 


I need all of the intersections for this plot.  
```{r}
area1=length(HumanTotal$gene)
area2= length(HumanNuclear$gene)
area3=length(ChimpTotal$gene)
area4=length(ChimpNuclear$gene)
n12= length(intersect(HumanTotal$gene, HumanNuclear$gene))
n13= length(intersect(HumanTotal$gene, ChimpTotal$gene) )
n14=length(intersect(HumanTotal$gene, ChimpNuclear$gene) )
n23=length(intersect(HumanNuclear$gene, ChimpTotal$gene) )
n24=length(intersect(HumanNuclear$gene,ChimpNuclear$gene))
n34=length(intersect(ChimpTotal$gene,ChimpNuclear$gene))
n123=length(Reduce(intersect, list(v1=HumanTotal$gene,v2= HumanNuclear$gene , v3=ChimpTotal$gene)))
n124=length(Reduce(intersect, list(v1=HumanTotal$gene,v2= HumanNuclear$gene , v4=ChimpNuclear$gene)))
n134=length(Reduce(intersect, list(v1=HumanTotal$gene, v3=ChimpTotal$gene , v4=ChimpNuclear$gene)))
n234=length(Reduce(intersect, list(v2= HumanNuclear$gene, v3=ChimpTotal$gene , v4=ChimpNuclear$gene)))
n1234=length(Reduce(intersect, list(v1=HumanTotal$gene,v2= HumanNuclear$gene, v3=ChimpTotal$gene , v4=ChimpNuclear$gene)))


```


Draw it:

```{r}
grid.newpage()
venn.plot.all=draw.quad.venn(area1,area2,area3,area4,n12,n13,n14,n23,n24,n34,n123,n124,n134,n234,n1234,category = c("Human Total", "Human Nuclear", "Chimp Total", "Chimp Nuclear"), fill = c("skyblue", "pink1", 
    "mediumorchid", "orange"), alpha = rep(0.5, 4), label.col = rep("black", 15), cex = rep(1, 15),fontface =rep("plain", 15),lty = "blank")
```

