---
title: "Conservation and dominance structure"
author: "Briana Mittleman"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here I can combine 2 lines of analysis. I will look to see if hte dominant PAS are more conserved than the non dominant PAS.   

```{r}
library(workflowr)
library(tidyverse)
library(ggpubr)
```

Load results on conservation with phylop.  

```{r}
PASMeta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% dplyr::select(PAS, chr, start,end, gene, loc, Human, Chimp)
phylores=read.table("../data/PhyloP/PAS_phyloP.txt", col.names = c("chr","start","end", "phyloP"), stringsAsFactors = F) %>% drop_na() %>% inner_join(PASMeta, by=c("chr", "start", "end"))
```

For the domiant PAS I will be inclusive on ties. 
```{r}

allPAS= read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T) 
ChimpPASwMean =allPAS %>% dplyr::select(-Human)
HumanPASwMean =allPAS %>% dplyr::select(-Chimp)

Chimp_Dom2= ChimpPASwMean %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) 
nrow(Chimp_Dom2%>% filter(nPer>1) )

Human_Dom2= HumanPASwMean %>%
  group_by(gene) %>%
  top_n(1,Human) %>% 
  mutate(nPer=n()) 
```

Add this information.  

```{r}
phylores_dom = phylores %>% mutate(ChimpDom=ifelse(PAS %in% Chimp_Dom2$PAS, "Yes","No"), HumanDom=ifelse(PAS %in% Human_Dom2$PAS, "Yes", "No"))
```


Plot:  

```{r}
ggplot(phylores_dom, aes(x=ChimpDom,y=phyloP, fill=ChimpDom)) + geom_boxplot() + scale_fill_brewer(palette = "Dark2") + stat_compare_means() + labs(x="Is PAS dominant in Chimp?", title="Chimp Dominant PAS are more conserved than non dominant PAS") + theme(legend.position = "none")
```

```{r}
ggplot(phylores_dom, aes(x=HumanDom,y=phyloP, fill=HumanDom)) + geom_boxplot() + scale_fill_brewer(palette = "Dark2") + stat_compare_means() + labs(x="Is PAS dominant in Human?", title="Human Dominant PAS are more conserved than non dominant PAS") + theme(legend.position = "none")

```

Seperate this by mean usage. 

Use the mean usag for human with dominance for human and vice versa 
```{r}

ggplot(phylores_dom, aes(x=Human, y=phyloP, by=HumanDom, color=HumanDom)) + geom_point(alpha=.3) + geom_smooth(method="lm") +  scale_color_brewer(palette = "Dark2") + labs(title="Human Dominance and Usage")

ggplot(phylores_dom, aes(x=Chimp, y=phyloP, by=ChimpDom, color=ChimpDom)) + geom_point(alpha=.3) + geom_smooth(method="lm") +  scale_color_brewer(palette = "Dark2")+ labs(title="Chimp Dominance and Usage")
```
This is not the best way to visualize this I should group it by deciles. 
```{r}
phylores_domDec= phylores_dom %>% mutate(decileHuman = ntile(Human, 10),decileChimp = ntile(Chimp, 10))

phylores_domDec$decileHuman=as.factor(phylores_domDec$decileHuman)
phylores_domDec$decileChimp=as.factor(phylores_domDec$decileChimp)


ggplot(phylores_domDec, aes(x=decileHuman, y=phyloP, by=HumanDom, fill=HumanDom)) + geom_boxplot() +  scale_fill_brewer(palette = "Dark2") + labs(title="Human Dominance and Usage") 

ggplot(phylores_domDec, aes(x=decileChimp, y=phyloP, by=ChimpDom, fill=ChimpDom)) + geom_boxplot() +  scale_fill_brewer(palette = "Dark2")+ labs(title="Chimp Dominance and Usage") 
```
This is decile but I want up to the decile.

I can do this with cumulative ifelse statements first. I will start with up to certain cutoffs. 


```{r}
phylores_dom_cumulative= phylores_dom %>% mutate(CutoffHuman=ifelse(Human <=.1, "ten", ifelse(Human<=.2, "twenty", ifelse(Human <=.3, "thirty", ifelse(Human <=.4 , "fourty", ifelse(Human <=.5, "fifty", ifelse(Human<=.6, "sixty", ifelse(Human <=.7, "seventy", ifelse(Human <=.8, "eighty", ifelse(Human <=.9, "neinty", "onehundred"))))))))))%>% mutate(CutoffChimp=ifelse(Chimp <=.1, "ten", ifelse(Chimp<=.2, "twenty", ifelse(Chimp <=.3, "thirty", ifelse(Chimp <=.4 , "fourty", ifelse(Chimp <=.5, "fifty", ifelse(Chimp<=.6, "sixty", ifelse(Chimp <=.7, "seventy", ifelse(Chimp <=.8, "eighty", ifelse(Chimp <=.9, "neinty", "onehundred"))))))))))

phylores_dom_cumulative$CutoffHuman=factor(phylores_dom_cumulative$CutoffHuman, levels=c("ten", "twenty", "thirty", "fourty","fifty", "sixty", "seventy", "eighty", "neinty", "onehundred"))
phylores_dom_cumulative$CutoffChimp=factor(phylores_dom_cumulative$CutoffChimp,levels=c("ten", "twenty", "thirty", "fourty","fifty", "sixty", "seventy", "eighty", "neinty", "onehundred"))
```

```{r}
ggplot(phylores_dom_cumulative, aes(x=CutoffHuman, y=phyloP, by=HumanDom, fill=HumanDom)) + geom_boxplot() +  scale_fill_brewer(palette = "Dark2", name="PAS Dominant \nin Human") + labs(title="Conservation of Human Dominant PAS \n by different usage cutoffs",x="Human Usage Cutoff, PAS with usage up to X%") 

ggplot(phylores_dom_cumulative, aes(x=CutoffChimp, y=phyloP, by=ChimpDom, fill=ChimpDom)) + geom_boxplot() +  scale_fill_brewer(palette = "Dark2", name="PAS Dominant \nin Chimp") + labs(title="Conservation of Chimp Dominant PAS \n by different usage cutoffs",x="Chimp Usage Cutoff, PAS with usage up to X%") 
```

When they have the same dominant:  

I need to do the no ties version for this.  
```{r}
Chimp_Dom= ChimpPASwMean %>%
  group_by(gene) %>%
  arrange(desc(Chimp)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom= HumanPASwMean %>%
  group_by(gene) %>%
  arrange(desc(Human)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)

BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")
SameDom= BothDom %>% filter(ChimpPAS==HumanPAS)
DiffDom=BothDom %>% filter(ChimpPAS!=HumanPAS) 


```

same domiant:  
```{r}
phylores_dom_samediff=phylores_dom %>% mutate(SameDom=ifelse(PAS %in%SameDom$ChimpPAS, "Yes", "No"), DiffDomChimp=ifelse(PAS %in% DiffDom$ChimpPAS, "Yes", "No" ), DiffDomHuman=ifelse(PAS %in% DiffDom$HumanPAS, "Yes","No"))
```

```{r}
ggplot(phylores_dom_samediff, aes(x=SameDom, y=phyloP, fill=SameDom)) + geom_boxplot() +stat_compare_means()+ scale_fill_brewer(palette = "Dark2")+ labs(x="Is PAS the same dominant in both species?", title="Same dominant are more conserved") + theme(legend.position = "none")
```
Plot 3 plot- same dominant, diff dom in chimp, diff dom human  

```{r}
BothDomPhy= BothDom %>%ungroup() %>%  select(HumanPAS) %>% rename("PAS"=HumanPAS) %>% inner_join(phylores_dom, by="PAS") %>% mutate(set="SameDominant")
DiffHumanPhy=DiffDom %>%ungroup() %>%  select(HumanPAS) %>% rename("PAS"=HumanPAS) %>% inner_join(phylores_dom, by="PAS") %>% mutate(set="DiffDominant_Human")
DiffChimpPhy=DiffDom %>%ungroup() %>%  select(ChimpPAS) %>% rename("PAS"=ChimpPAS) %>% inner_join(phylores_dom, by="PAS") %>% mutate(set="DiffDominant_Chimp")
NotDom=phylores_dom %>% filter(ChimpDom=="No", HumanDom=="Yes") %>% mutate(set="NotDominant")


AlldomrePhy=BothDomPhy %>% bind_rows(DiffHumanPhy) %>% bind_rows(DiffChimpPhy) %>% bind_rows(NotDom)

ggplot(AlldomrePhy, aes(x=set, y=phyloP, fill=set)) + geom_boxplot()  +stat_compare_means() + scale_fill_brewer(palette ="Dark2") + labs(x="", title="PAS that are dominant in both species are most conserved") +theme(legend.position = "none")


AlldomrePhy %>% group_by(set) %>% summarise(meanPhylop=round(mean(phyloP),2))
```

This is interesting. the chimp dominant.  
