---
title: "Number of Conserved PAS"
author: "Briana Mittleman"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(workflowr)
library(tidyverse)
```

I look at genes with the same dominant PAS but I want to look at the distribution of PAS in a gene. Are the number of PAS used for each gene conserved. I can use different usage filters then see if the distributions are the same or different when the gene has the same dominant site.  

One question is whether we see examples of 1 dominant site in one species and the other has more than one site used at similar proportions. 


I can start by looking at the genes with shared dominant PAS to see if the usages in each species are similar.  

I will filter for genes with more than 1 PAS 

```{r}

PASMeta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)
PASpregene=PASMeta %>% group_by(gene) %>% summarize(nPAS=n())
PASmore2=PASpregene %>% filter(nPAS>1)
DiffDom=read.table("../data/DominantPAS_DF/Nuclear_DiffDom.txt",header = T,stringsAsFactors = F) %>% filter(gene %in% PASmore2$gene)
SameDom=read.table("../data/DominantPAS_DF/Nuclear_SameDom.txt",header = T,stringsAsFactors = F) %>% mutate(DiffinDom=Chimp-Human) %>% filter(gene %in% PASmore2$gene)

```

Plot this: - is more in human: 

```{r}
ggplot(SameDom,aes(x=DiffinDom))+ geom_histogram(bins=100)
```

A bit shifted toward chimp. 

```{r}
summary(SameDom$DiffinDom)
```
I can look at examples at each extreme.  


Fore both species, i want to get the number of PAS used at certain cutoffs, 0%-50%

```{r}
Cutoff=seq(0,.5,.1)
cutoffCol=c()
nPAS=seq(1,5,1)
ChimpGenes=c()
HumanGenes=c()
nPAScol=c()
for (i in Cutoff){
  for (n in nPAS){
    human= PASMeta %>% filter(gene %in% PASmore2$gene, Human >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    HumanGenes=c(HumanGenes,human)
    chimp= PASMeta %>% filter(gene %in% PASmore2$gene,Chimp >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    ChimpGenes=c(ChimpGenes,chimp)
    nPAScol=c(nPAScol,n)
    cutoffCol=c(cutoffCol,i)
  }
}

DFdata=as.data.frame(cbind(nPAScol,cutoffCol,ChimpGenes, HumanGenes))
DFdata_gather=DFdata %>% gather("Species", "NGene", -nPAScol, -cutoffCol)
DFdata_gather$nPAScol=as.factor(DFdata_gather$nPAScol)
DFdata_gather$cutoffCol=as.factor(DFdata_gather$cutoffCol)

```

```{r}
ggplot(DFdata_gather,aes(x=cutoffCol, by=nPAScol, y=NGene,fill=nPAScol))+ geom_bar(position = "dodge", stat="identity") +facet_grid(~Species) + scale_fill_brewer(palette="Dark2", name="Number of PAS") + labs(title="Number of PAS per gene by usage",y="Number of Genes", x="Usage is at least")
```

Do the same thing for same dominant:  

```{r}

PASMetaSame= PASMeta %>% filter(gene %in% SameDom$gene)
cutoffCol_same=c()
ChimpGenes_same=c()
HumanGenes_same=c()
nPAScol_same=c()
for (i in Cutoff){
  for (n in nPAS){
    human= PASMetaSame %>% filter( Human >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    HumanGenes_same=c(HumanGenes_same,human)
    chimp= PASMetaSame %>% filter(Chimp >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    ChimpGenes_same=c(ChimpGenes_same,chimp)
    nPAScol_same=c(nPAScol_same,n)
    cutoffCol_same=c(cutoffCol_same,i)
  }
}

DFdata_same=as.data.frame(cbind(nPAScol_same,cutoffCol_same,ChimpGenes_same, HumanGenes_same))
DFdata_gather_same=DFdata_same %>% gather("Species", "NGene", -nPAScol_same, -cutoffCol_same)
DFdata_gather_same$nPAScol_same=as.factor(DFdata_gather_same$nPAScol_same)
DFdata_gather_same$cutoffCol_same=as.factor(DFdata_gather_same$cutoffCol_same)
```

```{r}
ggplot(DFdata_gather_same,aes(x=cutoffCol_same, by=nPAScol_same, y=NGene,fill=nPAScol_same))+ geom_bar(position = "dodge", stat="identity") +facet_grid(~Species) + scale_fill_brewer(palette="Dark2", name="Number of PAS") + labs(title="Number of PAS per gene by usage \n same domiant",y="Number of Genes", x="Usage is at least")
```


Different dominannt: 


```{r}
PASMetaDiff= PASMeta %>% filter(gene %in% DiffDom$gene)
cutoffCol_Diff=c()
ChimpGenes_Diff=c()
HumanGenes_Diff=c()
nPAScol_Diff=c()
for (i in Cutoff){
  for (n in nPAS){
    human= PASMetaDiff %>% filter( Human >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    HumanGenes_Diff=c(HumanGenes_Diff,human)
    chimp= PASMetaDiff %>% filter(Chimp >= i) %>% group_by(gene) %>% summarise(nPASingene=n()) %>% filter(nPASingene==n) %>% nrow() 
    ChimpGenes_Diff=c(ChimpGenes_Diff,chimp)
    nPAScol_Diff=c(nPAScol_Diff,n)
    cutoffCol_Diff=c(cutoffCol_Diff,i)
  }
}

DFdata_Diff=as.data.frame(cbind(nPAScol_Diff,cutoffCol_Diff,ChimpGenes_Diff, HumanGenes_Diff))
DFdata_gather_Diff=DFdata_Diff %>% gather("Species", "NGene", -nPAScol_Diff, -cutoffCol_Diff)
DFdata_gather_Diff$nPAScol_Diff=as.factor(DFdata_gather_Diff$nPAScol_Diff)
DFdata_gather_Diff$cutoffCol_Diff=as.factor(DFdata_gather_Diff$cutoffCol_Diff)
```


```{r}
ggplot(DFdata_gather_Diff,aes(x=cutoffCol_Diff, by=nPAScol_Diff, y=NGene,fill=nPAScol_Diff))+ geom_bar(position = "dodge", stat="identity") +facet_grid(~Species) + scale_fill_brewer(palette="Dark2", name="Number of PAS") + labs(title="Number of PAS per gene by usage \n different domiant",y="Number of Genes", x="Usage is at least")
```


Number of PAS in same vs diffferent:  

```{r}
NpasSame=PASMetaSame%>% group_by(gene) %>% summarise(nPAS=n()) %>% group_by(nPAS) %>% summarise(Same=n(),PropSame=Same/nrow(PASMetaSame)) %>% select(-Same)
NpasDiff=PASMetaDiff %>% group_by(gene) %>% summarise(nPAS=n()) %>% group_by(nPAS) %>% summarise(Different=n(), PropDiff=Different/nrow(PASMetaDiff)) %>% select(-Different)

NpasBoth= NpasSame %>% inner_join(NpasDiff) %>% gather("GeneSet", "Prop", -nPAS)
NpasBoth$nPAS=as.factor(NpasBoth$nPAS)

ggplot(NpasBoth,aes(fill=GeneSet, y=Prop, x=nPAS))+ geom_bar(position = "dodge", stat="identity") + scale_fill_brewer(palette="Dark2", labels=c("Different Dominant", "Same Dominant"),name="") + labs(y="Proportion of Genes", x="Number of PAS", title="Number of PAS distribution by Dominance structure")
```

