---
title: "Conservation with PhyloP 20"
author: "Briana Mittleman"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggpubr)
library(cowplot)
library(workflowr)
```

Copy and edit extractPhylopGeneral.py for phylop 20 way.  
```{bash,eval=F}
#extractPhylopGeneral_20way.py

python extractPhylopGeneral_20way.py  ../data/PhyloP/PAS_200downpregions2.bed ../data/PhyloP/PAS_phyloP20_200downstream2.txt 

python extractPhylopGeneral_20way.py ../data/PhyloP/PAS_200downpregions3.bed ../data/PhyloP/PAS_phyloP20_200downstream3.txt 
python extractPhylopGeneral_20way.py  ../data/PhyloP/PAS_200downpregions.bed ../data/PhyloP/PAS_phyloP20_200downstream.txt 

python extractPhylopGeneral_20way.py  ../data/PAS_doubleFilter/PAS_doublefilter_either_HumanCoordHummanUsage.sort.bed ../data/PhyloP/PAS_Phylop20.txt

python extractPhylopGeneral_20way.py ../data/PhyloP/PAS_200upregions.bed ../data/PhyloP/PAS_phyloP20_200upstream.txt
python extractPhylopGeneral_20way.py ../data/PhyloP/PAS_200upregions2.bed ../data/PhyloP/PAS_phyloP20_200upstream2.txt
python extractPhylopGeneral_20way.py ../data/PhyloP/PAS_200upregions3.bed ../data/PhyloP/PAS_phyloP20_200upstream3.txt
```


```{r}
PASMeta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% dplyr::select(PAS, chr, start,end, gene, loc)

phylores=read.table("../data/PhyloP/PAS_Phylop20.txt", col.names = c("chr","start","end","PAS", "phyloP"), stringsAsFactors = F) %>% drop_na()


Down=read.table("../data/PhyloP/PAS_phyloP20_200downstream.txt",col.names = c("chr", "start", "end", "PAS", "Down"),stringsAsFactors = F) %>% select(PAS, Down)%>% drop_na()
Down2=read.table("../data/PhyloP/PAS_phyloP20_200downstream2.txt",col.names = c("chr", "start", "end", "PAS", "Down2"),stringsAsFactors = F) %>% select(PAS, Down2)%>% drop_na()
Down3=read.table("../data/PhyloP/PAS_phyloP20_200downstream3.txt",col.names = c("chr", "start", "end", "PAS", "Down3"),stringsAsFactors = F) %>% select(PAS, Down3)%>% drop_na()


Up=read.table("../data/PhyloP/PAS_phyloP_200upstream.txt",col.names = c("chr", "start", "end", "PAS", "Up"),stringsAsFactors = F) %>% select(PAS, Up)%>% drop_na()
Up2=read.table("../data/PhyloP/PAS_phyloP_200upstream2.txt",col.names = c("chr", "start", "end", "PAS", "Up2"),stringsAsFactors = F) %>% select(PAS, Up2)%>% drop_na()
Up3=read.table("../data/PhyloP/PAS_phyloP_200upstream3.txt",col.names = c("chr", "start", "end", "PAS", "Up3"),stringsAsFactors = F) %>% select(PAS, Up3)%>% drop_na()

ResUpdownAll= phylores %>% inner_join(Down, by="PAS")%>% inner_join(Down2, by="PAS") %>% inner_join(Down3, by="PAS") %>% inner_join(Up, by="PAS") %>% inner_join(Up3, by="PAS")%>% inner_join(Up2, by="PAS") %>% select(-chr, -start,-end)


ResUpdownAll_gather= ResUpdownAll %>% gather("Set", "PhyloP", -PAS)

ResUpdownAll_gather$Set=factor(ResUpdownAll_gather$Set, levels=c("Up3", "Up2","Up", "phyloP","Down", "Down2", "Down3" ))
```


```{r}
ggplot(ResUpdownAll_gather, aes(x=Set, by=Set, fill=Set, y=PhyloP)) + geom_boxplot() + scale_fill_brewer(palette = "Dark2") + scale_x_discrete(labels=c("-600", "-400", "-200", '0','200','400','600')) + labs(x="Basepairs", title="PAS are more conserved than surrounding regions") +theme_classic()+ theme(legend.position = "none")
```

```{r}
PASMetaSm=PASMeta %>% select(PAS,loc)
ResUpdownAll_loc=ResUpdownAll_gather %>% inner_join(PASMetaSm,by="PAS") %>% mutate(region=ifelse(Set=="phyloP", "Yes", "No"))
cons20loc=ggplot(ResUpdownAll_loc, aes(x=Set, by=Set,fill=region, y=PhyloP)) + geom_boxplot(notch = T) + scale_fill_brewer(palette = "RdYlBu") + scale_x_discrete(labels=c("-600", "-400", "-200", '0','200','400','600')) + labs(x="Basepairs from PAS", title="PAS are more conserved than surrounding regions", y= "Phylop 20") + guides(fill = FALSE) + theme_classic() + facet_grid(~loc) + theme(axis.text.x =element_text(angle = 90))
```
```{r}
cons20=ggplot(ResUpdownAll_loc, aes(x=Set, by=Set, fill=region, y=PhyloP)) + geom_boxplot() + scale_fill_brewer(palette = "RdYlBu") + scale_x_discrete(labels=c("-600", "-400", "-200", '0','200','400','600')) + labs(x="Basepairs", title="PAS are more conserved than surrounding regions", y="Phylop 20") +theme_classic()+ theme(legend.position = "none")
```

```{r}
phylop20both=plot_grid(cons20,cons20loc, nrow = 2,scale = c(.9,.9), labels=c("a","b"))
phylop20both
```

```{r}

pdf("../output/revisions/phylop20.pdf", height=8, width=6)
phylop20both
dev.off()
```

Compare significant and non significant PAS:  

```{r}
DiffUsage=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherPAS_2_Nuclear.txt", header = T, stringsAsFactors = F)
DiffUsagePAS=DiffUsage %>% inner_join(PASMeta, by=c("gene","chr", "start", "end"))



NucReswPhy=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header = T, stringsAsFactors = F) %>% inner_join(phylores, by=c("chr","start","end"))


```

```{r}

ggplot(NucReswPhy,aes(y=phyloP, x=SigPAU2,fill=SigPAU2)) + geom_boxplot() + stat_compare_means()+ scale_fill_brewer(palette = "Dark2", name="Signficant")

ggplot(NucReswPhy,aes(x=phyloP, by=SigPAU2, fill=SigPAU2)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Dark2", name="Signficant PAS") + labs(title="Mean PhyloP scores for tested PAS") + annotate("text",label="Wilcoxan, p=0.0016",x=0,y=.75)

```

more <0 than expected:

```{r}
x=nrow(NucReswPhy %>% filter(SigPAU2=="Yes", phyloP<0))
m= nrow(NucReswPhy %>% filter(phyloP<0))
n=nrow(NucReswPhy %>% filter(phyloP>=0))
k=nrow(NucReswPhy %>% filter(SigPAU2=="Yes"))


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))
#actual:
x
#pval
phyper(x-1,m,n,k,lower.tail=F)


b=nrow(NucReswPhy %>% filter(SigPAU2=="Yes", phyloP<0))
n=nrow(NucReswPhy %>% filter(SigPAU2=="Yes"))
B=nrow(NucReswPhy %>% filter(phyloP<0))
N=nrow(NucReswPhy)

(b/n)/(B/N)
```
 This is even more enriched than the 100 way phyloP. 
