---
title: "DAPA enrichment in apaQTL genes"
author: "Briana Mittleman"
date: "1/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to ask if dAPA genes are enriched in the genes I found to have apaQTLs in my previous paper. This is similar to the analysis where I asked if dAPA genes are enriched for egenes. This is a more relevent analysis. I am asking if genes with species level differenences are enriched for those found to be genetically regulated.  


```{r}
library(workflowr)
library(tidyverse)
```

First I need to get the apaQTL genes from my other project. These are the nuclear apaQTLs at 5% fdr.  

```{bash,eval=F}
mkdir ../data/apaQTLGenes
#NuclearapaQTLGenes.txt
```

```{r}
QTL_genes=read.table("../data/apaQTLGenes/NuclearapaQTLGenes.txt", header = F, stringsAsFactors = F, col.names = "genes")
```


Now I will pull in the genes from the dAPA analysis in the nuclear fraction. I need both the sigificant genes as well as all of those tested.    

```{r}
sigNuclear_genes=read.table("../data/DiffIso_Nuclear/SignifianceEitherGENES_Nuclear.txt", stringsAsFactors = F, col.names = "genes")

NuclearAPAtested=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") %>% separate(cluster, into=c("chrom", "gene"), sep=":") %>% dplyr::select(gene)
NuclearAPAtested_genes=as.vector(NuclearAPAtested$gene)
```

First look at the overlap.  


```{r}
actual=QTL_genes %>% inner_join(sigNuclear_genes,by="genes") %>% nrow()
```


I will use permutations selecting 1502 genes from all of those tested to see if this is significant.  

```{r}
permuteGenes <- function(InputGenes, nGenes, nTests){
  #InputGenes=NuclearAPAtested_genes
  #nGenes=nrow(sigNuclear_genes)
  #nTests=10000
  qtls=QTL_genes
  qtlOverlap=c()
  for (n in 1:nTests){
    genesTest=sample(InputGenes, nGenes)
    overlapqtl=intersect(genesTest, qtls$genes)
    qtlOverlap= c(qtlOverlap, length(overlapqtl))
  }
  DF=as.data.frame(cbind(qtlOverlap))
  return(DF)
}

```


Run this:  
```{r}
overlapSample=permuteGenes(NuclearAPAtested_genes,nrow(sigNuclear_genes), 10000)
```

an empirical pvalue is the number of sample overlaps greater than the actual overlap. 


```{r}
empricalpval=overlapSample %>% filter(qtlOverlap>=actual) %>% nrow() / 10000
```


Plot the results:  

```{r}
ggplot(overlapSample,aes(x=qtlOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =actual, col="red" ) +labs(x="Number of Overlaps", title="Nuclear dAPA overlap with apaQTL Genes") +  annotate("text", x = 110, y = 400, label = paste("Empirical Pvalue=", empricalpval), col="black") +  annotate("text", x = 168, y = 400, label = "Actual Overlap", col="red")
```


```{r}
overlapgenes=QTL_genes %>% inner_join(sigNuclear_genes,by="genes")
```

