---
title: "Differential APA and eQTLs"
author: "Briana Mittleman"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(workflowr)
```


To start to catalog my differential apa gene, I want to evaluate how many of them have an eQTL. I will use the eQTLs from Li et al. I have the explained and unexplained eGenes. 

I will put them in ../data/Li_eqtls

```{bash,eval=F}
mkdir ../data/Li_eqtls
```

I want to make plots that have bars for explained eQTL, unexplained eQTL, not an eQTL for total differences and nuclear differences.  

```{r}
explained=read.table("../data/Li_eqtls/explainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
unexplained=read.table("../data/Li_eqtls/UnexplainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
```


###Nuclear:

```{r}
sigNuclear=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigNuclear$p.adjust=as.numeric(as.character(sigNuclear$p.adjust))

sigNuclear_genes=sigNuclear %>% filter(p.adjust<.05) %>% separate(cluster, into=c("chrom", "genes"), sep=":") %>% dplyr::select(genes) %>% unique()
```


```{r}
explainedNuclear= sigNuclear_genes %>% inner_join(explained, by = "genes")
unexplainedNuclear=sigNuclear_genes %>% inner_join(unexplained, by="genes")
NeitherNuclear= sigNuclear_genes %>%  anti_join(explained, by="genes") %>% anti_join(unexplained, by="genes")

cat=c("Explained", "Unexaplained", "Neither")
num=c(nrow(explainedNuclear), nrow(unexplainedNuclear), nrow(NeitherNuclear))

Nuclearqtl=as.data.frame(cbind(cat,num))

Nuclearqtl$cat=factor(Nuclearqtl$cat, levels=c("Explained","Unexaplained", "Neither"), ordered=T)
Nuclearqtl$num=as.numeric(as.character(Nuclearqtl$num ))

ggplot(Nuclearqtl,aes(x=cat, y=num)) +geom_bar(stat="identity") + geom_text(aes(label=num), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Nuclear fraction and human eQTL", y="Number of genes", x="Overlap")
```
Make this plot with proportion:

```{r}
NuclearqtlProp=Nuclearqtl %>% mutate(Prop=num/nrow(sigNuclear_genes))

ggplot(NuclearqtlProp,aes(x=cat, y=Prop)) +geom_bar(stat="identity") + geom_text(aes(label=round(Prop,digits = 3)), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Nuclear fraction and human eQTL", y="Proportion of dAPA genes", x="Overlap")
```

###Total 


```{r}
sigTotal=read.table("../data/DiffIso_Total/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigTotal$p.adjust=as.numeric(as.character(sigTotal$p.adjust))

sigTotal_genes=sigTotal %>% filter(p.adjust<.05) %>% separate(cluster, into=c("chrom", "genes"), sep=":") %>% dplyr::select(genes) %>% unique()
```


```{r}
explainedTotal= sigTotal_genes %>% inner_join(explained, by = "genes")
unexplainedTotal=sigTotal_genes %>% inner_join(unexplained, by="genes")
NeitherTotal= sigTotal_genes %>%  anti_join(explained, by="genes") %>% anti_join(unexplained, by="genes")

catTotal=c("Explained", "Unexaplained", "Neither")
numTotal=c(nrow(explainedTotal), nrow(unexplainedTotal), nrow(NeitherTotal))

Totalqtl=as.data.frame(cbind(catTotal,numTotal))

Totalqtl$catTotal=factor(Totalqtl$catTotal, levels=c("Explained","Unexaplained", "Neither"), ordered=T)
Totalqtl$numTotal=as.numeric(as.character(Totalqtl$numTotal ))

ggplot(Totalqtl,aes(x=catTotal, y=numTotal)) +geom_bar(stat="identity") + geom_text(aes(label=numTotal), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Total fraction and human eQTL", y="Number of genes", x="Overlap")

```

Make this plot with proportion:

```{r}
TotalqtlProp=Totalqtl %>% mutate(Prop=num/nrow(sigTotal_genes))

ggplot(TotalqtlProp,aes(x=cat, y=Prop)) +geom_bar(stat="identity") + geom_text(aes(label=round(Prop,digits = 3)), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Total fraction and human eQTL", y="Proportion of dAPA genes", x="Overlap")
```

###Test enrichment
To test if these values are enriched I need to write a function that randomly chooses the same number of genes and assess the same overlaps. I will use all of the genes that we tested for APA.

```{r}
NuclearAPAtested=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") %>% separate(cluster, into=c("chrom", "gene"), sep=":") %>% dplyr::select(gene)
NuclearAPAtested_genes=as.vector(NuclearAPAtested$gene)
TotalAPAtested=read.table("../data/DiffIso_Total/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") %>% separate(cluster, into=c("chrom", "gene"), sep=":") %>% dplyr::select(gene)
TotalAPAtested_genes=as.vector(TotalAPAtested$gene)
```


I can run it 100 times to have error bars.  

```{r}
permuteGenes <- function(InputGenes, nGenes, nTests){
  #InputGenes=NuclearAPAtested_genes
  #nGenes=nrow(sigNuclear_genes)
  #nTests=100
  explained=read.table("../data/Li_eqtls/explainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
  unexplained=read.table("../data/Li_eqtls/UnexplainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
  explainedOverlap=c()
  unexplainedOverlap=c()
  neitherOverlap=c()
  for (n in 1:nTests){
    genesTest=sample(InputGenes, nGenes)
    overlapE=intersect(genesTest, explained$genes) 
    overlapUN=intersect(genesTest, unexplained$genes) 
    diffs <- Reduce(setdiff,list(A = genesTest, B = explained$genes,C = unexplained$genes))
    explainedOverlap= c(explainedOverlap, length(overlapE))
    unexplainedOverlap=c(unexplainedOverlap,length(overlapUN))
    neitherOverlap=c(neitherOverlap, length(diffs))
  }
  DF=as.data.frame(cbind(explainedOverlap,unexplainedOverlap,neitherOverlap))
  return(DF)
}
```


Perform permutations 1000 times
```{r}
Nuclear_100tests=permuteGenes(NuclearAPAtested_genes, nrow(sigNuclear_genes),1000 )

Total_100tests=permuteGenes(TotalAPAtested_genes, nrow(sigTotal_genes),1000 )
```

empirical pvalues
```{r}
nuclearEx=Nuclear_100tests %>% filter(explainedOverlap>=nrow(explainedNuclear)) %>% nrow() / 1000
nuclearUnEx=Nuclear_100tests %>% filter(unexplainedOverlap>=nrow(unexplainedNuclear)) %>% nrow() / 1000

totalEx=Total_100tests %>% filter(explainedOverlap>=nrow(explainedTotal)) %>% nrow() / 1000
totalUnEx=Total_100tests %>% filter(unexplainedOverlap>=nrow(unexplainedTotal)) %>% nrow() / 1000
```

Plot as histograms 

Nuclear Explained
```{r}

ggplot(Nuclear_100tests,aes(x=explainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(explainedNuclear), col="red" )+labs(x="Number of Overlaps", title="Nuclear dAPA overlap with explained eGenes") + annotate("text", x = 490, y = 40, label = paste("EmpPvalue=", nuclearEx), col="red")

```

 
Nuclear Unexplained
```{r}
ggplot(Nuclear_100tests,aes(x=unexplainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(unexplainedNuclear), col="red" ) +labs(x="Number of Overlaps", title="Nuclear dAPA overlap with unexplained eGenes")+ annotate("text", x = 376, y = 40, label = paste("EmpPvalue=", nuclearUnEx), col="red")
```
 

Total Explained

```{r}

ggplot(Total_100tests,aes(x=explainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(explainedTotal), col="red" )+labs(x="Number of Overlaps", title="Total dAPA overlap with explained eGenes")+ annotate("text", x = 425, y = 40, label = paste("EmpPvalue=", totalEx), col="red")

```

Total Unexplained


```{r}
ggplot(Total_100tests,aes(x=unexplainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(unexplainedTotal), col="red" ) +labs(x="Number of Overlaps", title="Total dAPA overlap with unexplained eGenes")+ annotate("text", x = 325, y = 30, label = paste("EmpPvalue=", totalEx), col="red")
```
###extra code:


```{r}

# NuclearEx_mean=mean(Nuclear_100tests$explainedOverlap)
# NuclearEx_sd=sd(Nuclear_100tests$explainedOverlap)
# NuclearUn_mean=mean(Nuclear_100tests$unexplainedOverlap)
# NuclearUn_sd=sd(Nuclear_100tests$unexplainedOverlap)
# NuclearN_mean=mean(Nuclear_100tests$neitherOverlap)
# NuclearN_sd=sd(Nuclear_100tests$neitherOverlap)
# 
# Nuclear_100testsDf=as.data.frame(cbind(cat=c("Explained","Unexplained", "Neither"), Mean=c(NuclearEx_mean,NuclearUn_mean,NuclearN_mean), SD=c(NuclearEx_sd,NuclearUn_sd,NuclearN_sd), actual=Nuclearqtl$num))
# 
# 
# Nuclear_100testsDf$cat=factor(Nuclear_100testsDf$cat, levels=c("Explained","Unexplained", "Neither"), ordered=T)
# Nuclear_100testsDf$Mean=as.numeric(as.character(Nuclear_100testsDf$Mean))
# Nuclear_100testsDf$SD=as.numeric(as.character(Nuclear_100testsDf$SD))
# Nuclear_100testsDf$actual=as.numeric(as.character(Nuclear_100testsDf$actual))
# 
# ggplot(Nuclear_100testsDf,aes(x=cat,y=Mean)) + geom_bar(stat = "identity", alpha=.5) + geom_errorbar(aes(x=cat, ymin=Mean-SD, ymax=Mean+SD), width=0.2, size=1) + geom_point(aes(x=cat, y=actual), col="red")

```



Total Explained

```{r}

ggplot(Total_100tests,aes(x=explainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(explainedTotal), col="red" )+labs(x="Number of Overlaps", title="Total dAPA overlap with explained eGenes")

```

Total Unexplained


```{r}
ggplot(Total_100tests,aes(x=unexplainedOverlap)) + geom_histogram(stat="count") + geom_vline(xintercept =nrow(unexplainedTotal), col="red" ) +labs(x="Number of Overlaps", title="Total dAPA overlap with unexplained eGenes")
```

###extra code:


```{r}

# NuclearEx_mean=mean(Nuclear_100tests$explainedOverlap)
# NuclearEx_sd=sd(Nuclear_100tests$explainedOverlap)
# NuclearUn_mean=mean(Nuclear_100tests$unexplainedOverlap)
# NuclearUn_sd=sd(Nuclear_100tests$unexplainedOverlap)
# NuclearN_mean=mean(Nuclear_100tests$neitherOverlap)
# NuclearN_sd=sd(Nuclear_100tests$neitherOverlap)
# 
# Nuclear_100testsDf=as.data.frame(cbind(cat=c("Explained","Unexplained", "Neither"), Mean=c(NuclearEx_mean,NuclearUn_mean,NuclearN_mean), SD=c(NuclearEx_sd,NuclearUn_sd,NuclearN_sd), actual=Nuclearqtl$num))
# 
# 
# Nuclear_100testsDf$cat=factor(Nuclear_100testsDf$cat, levels=c("Explained","Unexplained", "Neither"), ordered=T)
# Nuclear_100testsDf$Mean=as.numeric(as.character(Nuclear_100testsDf$Mean))
# Nuclear_100testsDf$SD=as.numeric(as.character(Nuclear_100testsDf$SD))
# Nuclear_100testsDf$actual=as.numeric(as.character(Nuclear_100testsDf$actual))
# 
# ggplot(Nuclear_100testsDf,aes(x=cat,y=Mean)) + geom_bar(stat = "identity", alpha=.5) + geom_errorbar(aes(x=cat, ymin=Mean-SD, ymax=Mean+SD), width=0.2, size=1) + geom_point(aes(x=cat, y=actual), col="red")

```

 
