---
title: "Differential APA and eQTLs"
author: "Briana Mittleman"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(workflowr)
```


To start to catalog my differential apa gene, I want to evaluate how many of them have an eQTL. I will use the eQTLs from Li et al. I have the explained and unexplained eGenes. 

I will put them in ../data/Li_eqtls

```{bash,eval=F}
mkdir ../data/Li_eqtls
```

I want to make plots that have bars for explained eQTL, unexplained eQTL, not an eQTL for total differences and nuclear differences.  

```{r}
explained=read.table("../data/Li_eqtls/explainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
unexplained=read.table("../data/Li_eqtls/UnexplainedEgenes.txt",col.names = c("genes"),stringsAsFactors = F)
```


###Nuclear:

```{r}
sigNuclear=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigNuclear$p.adjust=as.numeric(as.character(sigNuclear$p.adjust))

sigNuclear_genes=sigNuclear %>% filter(p.adjust<.05) %>% separate(cluster, into=c("chrom", "genes"), sep=":") %>% dplyr::select(genes) %>% unique()
```


```{r}
explainedNuclear= sigNuclear_genes %>% inner_join(explained, by = "genes")
unexplainedNuclear=sigNuclear_genes %>% inner_join(unexplained, by="genes")
NeitherNuclear= sigNuclear_genes %>%  anti_join(explained, by="genes") %>% anti_join(unexplained, by="genes")

cat=c("Explained", "Unexaplained", "Neither")
num=c(nrow(explainedNuclear), nrow(unexplainedNuclear), nrow(NeitherNuclear))

Nuclearqtl=as.data.frame(cbind(cat,num))

Nuclearqtl$cat=factor(Nuclearqtl$cat, levels=c("Explained","Unexaplained", "Neither"), ordered=T)
Nuclearqtl$num=as.numeric(as.character(Nuclearqtl$num ))

ggplot(Nuclearqtl,aes(x=cat, y=num)) +geom_bar(stat="identity") + geom_text(aes(label=num), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Nuclear fraction and human eQTL", y="Number of genes", x="Overlap")
```
Make this plot with proportion:

```{r}
NuclearqtlProp=Nuclearqtl %>% mutate(Prop=num/nrow(sigNuclear_genes))

ggplot(NuclearqtlProp,aes(x=cat, y=Prop)) +geom_bar(stat="identity") + geom_text(aes(label=round(Prop,digits = 3)), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Nuclear fraction and human eQTL", y="Proportion of dAPA genes", x="Overlap")
```

###Total 


```{r}
sigTotal=read.table("../data/DiffIso_Total/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigTotal$p.adjust=as.numeric(as.character(sigTotal$p.adjust))

sigTotal_genes=sigTotal %>% filter(p.adjust<.05) %>% separate(cluster, into=c("chrom", "genes"), sep=":") %>% dplyr::select(genes) %>% unique()
```


```{r}
explainedTotal= sigTotal_genes %>% inner_join(explained, by = "genes")
unexplainedTotal=sigTotal_genes %>% inner_join(unexplained, by="genes")
NeitherTotal= sigTotal_genes %>%  anti_join(explained, by="genes") %>% anti_join(unexplained, by="genes")

catTotal=c("Explained", "Unexaplained", "Neither")
numTotal=c(nrow(explainedTotal), nrow(unexplainedTotal), nrow(NeitherTotal))

Totalqtl=as.data.frame(cbind(catTotal,numTotal))

Totalqtl$catTotal=factor(Totalqtl$catTotal, levels=c("Explained","Unexaplained", "Neither"), ordered=T)
Totalqtl$numTotal=as.numeric(as.character(Totalqtl$numTotal ))

ggplot(Totalqtl,aes(x=catTotal, y=numTotal)) +geom_bar(stat="identity") + geom_text(aes(label=numTotal), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Total fraction and human eQTL", y="Number of genes", x="Overlap")

```

Make this plot with proportion:

```{r}
TotalqtlProp=Totalqtl %>% mutate(Prop=num/nrow(sigTotal_genes))

ggplot(TotalqtlProp,aes(x=cat, y=Prop)) +geom_bar(stat="identity") + geom_text(aes(label=round(Prop,digits = 3)), vjust=1.6, color="white", size=3.5) + labs(title="Differential APA in Total fraction and human eQTL", y="Proportion of dAPA genes", x="Overlap")
```


