---
title: "dIC and other metrics"
author: "Briana Mittleman"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I used simpson to call differences in information content. 

```{r}
library(workflowr)
library(tidyverse)

```

I want to look at DE, dP, dt, dAPA, and dIC. I will do simple 2by2 tables at first.  

For each set I will only consider the genes that I can test for those analylsis. 

##Load all set:

APA
```{r}
Meta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)  %>% dplyr::select(PAS, chr, start,end, loc)
DiffIso= read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header = T,stringsAsFactors = F) %>% inner_join(Meta, by=c("chr", 'start','end')) 

#gene level:  

SigGenesDI=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt",header = T,stringsAsFactors = F)
DiffIsoGene= DiffIso %>% select(gene) %>% unique() %>% mutate(dAPA=ifelse(gene %in% SigGenesDI$gene, "Yes", "No"))


DiffIsoGene %>% group_by(dAPA) %>% summarise(n())
```


DE  

```{r}
nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F) %>% dplyr::select(Gene_stable_ID, Gene.name)

DiffExp=read.table("../data/DiffExpression/DEtested_allres.txt",stringsAsFactors = F,header = F, col.names = c("Gene_stable_ID" ,"logFC" ,"AveExpr" , "t" ,  "P.Value" ,  "adj.P.Val", "B"  )) %>% inner_join(nameID,by="Gene_stable_ID") %>% dplyr::rename('gene'=Gene.name) %>% dplyr::select(-Gene_stable_ID) %>% mutate(DE=ifelse(adj.P.Val<.05, "Yes", "No"))


DiffExp %>% group_by(DE) %>% summarise(n())

DiffExpSmall= DiffExp %>% select(gene,DE)
```


DTE 

```{r}
Ribo=read.table("../data/Wang_ribo/Additionaltable5_translationComparisons.txt",header = T, stringsAsFactors = F) %>% rename("Gene_stable_ID"= ENSG) %>% inner_join(nameID,by="Gene_stable_ID") %>% dplyr::select(Gene.name, HvC.beta, HvC.pvalue, HvC.FDR) %>% rename("gene"=Gene.name) %>% mutate(dTE=ifelse(HvC.FDR <0.05, "Yes","No"))


Ribo %>% group_by(dTE) %>% summarise(n())

RiboSmall= Ribo %>% select(gene,dTE)


```


DP  
(pval is adjusted already)
```{r}
Prot= read.table("../data/Khan_prot/ProtData_effectSize.txt",header = T)  %>% mutate(dP=ifelse(pval<0.05, "Yes", "No"))

Prot %>% group_by(dP) %>% summarise(n())

ProtSmall=Prot %>% select(gene, dP)
```


Simpson Info Content  

```{r}
dICdata= read.table("../data/IndInfoContent/SimpsonMedianSignificance.txt", header = T, stringsAsFactors = F) %>% rename(dIC=sIC)

dICdata %>% group_by(dIC) %>% summarise(n())

dICSmall=dICdata %>% select(gene, dIC)

dICdata10= read.table("../data/IndInfoContent/SimpsonMedianSignificance_10FDR.txt", header = T, stringsAsFactors = F) %>% rename(dIC10=sIC10)


dICdata10 %>% group_by(dIC10) %>% summarise(n())

dIC10Small=dICdata10 %>% select(gene, dIC10)
```

##5% FDR  

###dAPA and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dICanddAPA= dICSmall %>% inner_join(DiffIsoGene, by="gene") 


dICanddAPA22=dICanddAPA %>% group_by(dIC, dAPA) %>% summarise(n=n()) %>% spread(dAPA, n) %>% column_to_rownames("dIC")
dICanddAPA22


```

Do this with proportion:  
```{r}
dICanddAPA %>% group_by(dIC, dAPA) %>% summarise(n=n()) %>% mutate(nG=nrow(dICanddAPA),Prop=n/nG)  %>% select(dIC, dAPA, Prop) %>% spread(dAPA, Prop)
```

Enrichment: 

```{r}
x=nrow(dICanddAPA %>% filter(dIC=="Yes", dAPA=="Yes"))
m=nrow(dICanddAPA %>% filter(dAPA=="Yes"))
n=nrow(dICanddAPA %>% filter(dAPA=="No"))
k=nrow(dICanddAPA %>% filter(dIC=="Yes"))
N=nrow(dICanddAPA)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

###dE and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dICanddE= dICSmall %>% inner_join(DiffExpSmall, by="gene") 


dICanddE22=dICanddE %>% group_by(dIC, DE) %>% summarise(n=n()) %>% spread(DE, n) %>% column_to_rownames("dIC")
dICanddE22


```
Do this with proportion:  
```{r}
dICanddE %>% group_by(dIC, DE) %>% summarise(n=n()) %>% mutate(nG=nrow(dICanddE),Prop=n/nG)  %>% select(dIC, DE, Prop) %>% spread(DE, Prop)
```


```{r}
x=nrow(dICanddE %>% filter(dIC=="Yes", DE=="Yes"))
m=nrow(dICanddE %>% filter(DE=="Yes"))
n=nrow(dICanddE %>% filter(DE=="No"))
k=nrow(dICanddE %>% filter(dIC=="Yes"))
N=nrow(dICanddE)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```



###dT and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dICanddT= dICSmall %>% inner_join(RiboSmall, by="gene") 


dICanddT22=dICanddT %>% group_by(dIC, dTE) %>% summarise(n=n()) %>% spread(dTE, n) %>% column_to_rownames("dIC")
dICanddT22



```
Do this with proportion:  
```{r}
dICanddT %>% group_by(dIC, dTE) %>% summarise(n=n()) %>% mutate(nG=nrow(dICanddT),Prop=n/nG)  %>% select(dIC, dTE, Prop) %>% spread(dTE, Prop)
```


```{r}
x=nrow(dICanddT %>% filter(dIC=="Yes", dTE=="Yes"))
m=nrow(dICanddT %>% filter(dTE=="Yes"))
n=nrow(dICanddT %>% filter(dTE=="No"))
k=nrow(dICanddT %>% filter(dIC=="Yes"))
N=nrow(dICanddT)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

###dP and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dICanddP= dICSmall %>% inner_join(ProtSmall, by="gene") 


dICanddP22=dICanddP %>% group_by(dIC, dP) %>% summarise(n=n()) %>% spread(dP, n) %>% column_to_rownames("dIC")
dICanddP22



```
Do this with proportion:  
```{r}
dICanddP %>% group_by(dIC, dP) %>% summarise(n=n()) %>% mutate(nG=nrow(dICanddP),Prop=n/nG)  %>% select(dIC, dP, Prop) %>% spread(dP, Prop)
```


```{r}
x=nrow(dICanddP %>% filter(dIC=="Yes", dP=="Yes"))
m=nrow(dICanddP %>% filter(dP=="Yes"))
n=nrow(dICanddP %>% filter(dP=="No"))
k=nrow(dICanddP %>% filter(dIC=="Yes"))
N=nrow(dICanddP)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```
Not enough power for this one.  


##10%FDR  


###dAPA and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dIC10anddAPA= dIC10Small %>% inner_join(DiffIsoGene, by="gene") 


dIC10anddAPA22=dIC10anddAPA %>% group_by(dIC10, dAPA) %>% summarise(n=n()) %>% spread(dAPA, n) %>% column_to_rownames("dIC10")
dIC10anddAPA22



```

Do this with proportion:  
```{r}
dIC10anddAPA %>% group_by(dIC10, dAPA) %>% summarise(n=n()) %>% mutate(nG=nrow(dIC10anddAPA),Prop=n/nG)  %>% select(dIC10, dAPA, Prop) %>% spread(dAPA, Prop)
```

Enrichment: 

```{r}
x=nrow(dIC10anddAPA %>% filter(dIC10=="Yes", dAPA=="Yes"))
m=nrow(dIC10anddAPA %>% filter(dAPA=="Yes"))
n=nrow(dIC10anddAPA %>% filter(dAPA=="No"))
k=nrow(dIC10anddAPA %>% filter(dIC10=="Yes"))
N=nrow(dIC10anddAPA)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

###dE and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dIC10anddE= dIC10Small %>% inner_join(DiffExpSmall, by="gene") 


dIC10anddE22=dIC10anddE %>% group_by(dIC10, DE) %>% summarise(n=n()) %>% spread(DE, n) %>% column_to_rownames("dIC10")
dIC10anddE22



```
Do this with proportion:  
```{r}
dIC10anddE %>% group_by(dIC10, DE) %>% summarise(n=n()) %>% mutate(nG=nrow(dIC10anddE),Prop=n/nG)  %>% select(dIC10, DE, Prop) %>% spread(DE, Prop)
```


```{r}
x=nrow(dIC10anddE %>% filter(dIC10=="Yes", DE=="Yes"))
m=nrow(dIC10anddE %>% filter(DE=="Yes"))
n=nrow(dIC10anddE %>% filter(DE=="No"))
k=nrow(dIC10anddE %>% filter(dIC10=="Yes"))
N=nrow(dIC10anddE)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```



###dT and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dIC10andT= dIC10Small %>% inner_join(RiboSmall, by="gene") 


dIC10andT22=dIC10andT %>% group_by(dIC10, dTE) %>% summarise(n=n()) %>% spread(dTE, n) %>% column_to_rownames("dIC10")
dIC10andT22


```
Do this with proportion:  
```{r}
dIC10andT %>% group_by(dIC10, dTE) %>% summarise(n=n()) %>% mutate(nG=nrow(dIC10andT),Prop=n/nG)  %>% select(dIC10, dTE, Prop) %>% spread(dTE, Prop)
```


```{r}
x=nrow(dIC10andT %>% filter(dIC10=="Yes", dTE=="Yes"))
m=nrow(dIC10andT %>% filter(dTE=="Yes"))
n=nrow(dIC10andT %>% filter(dTE=="No"))
k=nrow(dIC10andT %>% filter(dIC10=="Yes"))
N=nrow(dIC10andT)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

###dP and dIC  

I will start with dIC and dAPA, I expect a pretty high overlap for this.  

```{r}
dIC10anddP= dIC10Small %>% inner_join(ProtSmall, by="gene") 


dIC10anddP22=dIC10anddP %>% group_by(dIC10, dP) %>% summarise(n=n()) %>% spread(dP, n) %>% column_to_rownames("dIC10")
dIC10anddP22



```
Do this with proportion:  
```{r}
dIC10anddP %>% group_by(dIC10, dP) %>% summarise(n=n()) %>% mutate(nG=nrow(dIC10anddP),Prop=n/nG)  %>% select(dIC10, dP, Prop) %>% spread(dP, Prop)
```


```{r}
x=nrow(dIC10anddP %>% filter(dIC10=="Yes", dP=="Yes"))
m=nrow(dIC10anddP %>% filter(dP=="Yes"))
n=nrow(dIC10anddP %>% filter(dP=="No"))
k=nrow(dIC10anddP %>% filter(dIC10=="Yes"))
N=nrow(dIC10anddP)
phyper(x-1,m,n,k,lower.tail=F)
(x/k)/(m/N)
```
Not enough power for this one.  

