---
title: "Distribution of dominance"
author: "Briana Mittleman"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(workflowr)
library(tidyverse)
```

I changed the way I look at Dominance. I am looking at the difference between the top used and next used PAS. I want to look at the overall distribution of this metric. This will help when I look on same vs different dominant PAS as well. I will create a python script that writes out the top PAS the gene and how far the next PAS usage is away.  

The code will be adopted from the FindDomXCutoff.py.  

```{bash,eval=F}
#test
python GetTopminus2Usage.py  ../data/DomDefGreaterX/TestFile_ZSWIM7.txt  ../data/DomDefGreaterX/TestDomAll_ZSWIM7.txt Human

#run

python GetTopminus2Usage.py ../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt ../data/DomDefGreaterX/Human_AllGenes_DiffTop.txt Human

python GetTopminus2Usage.py ../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt ../data/DomDefGreaterX/Chimp_AllGenes_DiffTop.txt Chimp
```


Again I want the one 1 PAS genes: 

```{r}
MetaPAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)
MetaCol=colnames(MetaPAS)
Human1PASGene= MetaPAS %>% filter(Human>0) %>% group_by(gene) %>% summarise(nPAS=n()) %>% filter(nPAS==1)
Chimp1PASGene= MetaPAS %>% filter(Chimp>0) %>% group_by(gene) %>% summarise(nPAS=n()) %>% filter(nPAS==1)

MetaPAS_human1= MetaPAS %>% filter(gene %in% Human1PASGene$gene, Human >0) %>% mutate(Set="Human1")
MetaPAS_chimp1= MetaPAS %>% filter(gene %in% Chimp1PASGene$gene, Chimp >0)  %>% mutate(Set="Chimp1")

MetaPAS_human1 %>% anti_join(MetaPAS_chimp1,by="gene") %>% nrow()

MetaPAS_chimp1 %>% anti_join(MetaPAS_human1,by="gene") %>% nrow()
```



```{r}
HumanRes=read.table("../data/DomDefGreaterX/Human_AllGenes_DiffTop.txt", col.names = c("Human_PAS", "gene","Human_DiffDom"),stringsAsFactors = F)

ChimpRes=read.table("../data/DomDefGreaterX/Chimp_AllGenes_DiffTop.txt", col.names = c("Chimp_PAS", "gene","Chimp_DiffDom"),stringsAsFactors = F)

BothRes=HumanRes %>% inner_join(ChimpRes,by="gene")

nrow(BothRes)

BothRes_g= BothRes %>% select(gene, Chimp_DiffDom, Human_DiffDom) %>% gather("species", "diff", -gene)
```

I am going to lose the few genes that have 1 PAS in only one species. 


```{r}
ggplot(BothRes_g,aes(x=diff,fill=species, by=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Dark2", labels=c("Chimp","Human")) + labs(title="Chimp Dominant PAS are 'more Dominant'", x="Top PAS Usage - Second PAS Usage")

wilcox.test(BothRes$Human_DiffDom, BothRes$Chimp_DiffDom)
```

I can play with cutoffs through this to get a set of genes where I can look at same and different PAS.  

```{r}
BothRes_40=BothRes %>% filter(Chimp_DiffDom >=0.4 | Human_DiffDom>=0.4) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"))
nrow(BothRes_40)
BothRes_40_same=BothRes_40 %>% filter(Human_PAS==Chimp_PAS)
nrow(BothRes_40_same)
BothRes_40_diff=BothRes_40 %>% filter(Human_PAS!=Chimp_PAS)
nrow(BothRes_40_diff)
```

Most of these share the same dominant but it is interesting to look at the distribution of the dominant PAS usage distributions. 

```{r}
BothRes_g_same40= BothRes_g %>% filter(gene %in% BothRes_40_same$gene)

ggplot(BothRes_g_same40,aes(x=diff,fill=species, by=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Dark2", labels=c("Chimp","Human")) + labs(title="Chimp Dominant PAS are 'more Dominant \n Same Dominant PAS at .4 cutoff'", x="Top PAS Usage - Second PAS Usage")

```

```{r}
ggplot(BothRes_40, aes(x=Human_DiffDom,Chimp_DiffDom, col= Set ))  + geom_point(alpha=.2)+ geom_abline(slope=1, intercept = 0) + geom_density2d() + scale_color_brewer(palette = "Dark2") + labs(title="Level of dominance for each PAS passing .4 filter")
```

Look at the number of PAS:  (just do number in either species for now)

```{r}
PASnum= MetaPAS %>% select(PAS, gene) %>%  group_by(gene) %>% summarise(nPAS=n())



BothRes_40_num= BothRes_40 %>% inner_join(PASnum,by="gene")


ggplot(BothRes_40_num, aes(x=Human_DiffDom,Chimp_DiffDom, col= Set ))  + geom_point(alpha=.2)+ geom_abline(slope=1, intercept = 0) + geom_density2d() + scale_color_brewer(palette = "Dark2") + labs(title="Level of dominance for each PAS passing .4 filter") + facet_grid(~nPAS)
```


```{r}
BothRes_40_num %>% group_by(nPAS, Set) %>% summarise(n=n()) %>% spread(Set, n)
```
The plot is only inforamative up to about 6. Filter this: 

```{r}
BothRes_40_num_filt= BothRes_40_num %>% filter(nPAS<=6)

ggplot(BothRes_40_num_filt, aes(x=Human_DiffDom,Chimp_DiffDom, col= Set ))  + geom_point(alpha=.2)+ geom_abline(slope=1, intercept = 0) + geom_density2d() + scale_color_brewer(palette = "Dark2") + labs(title="Level of dominance for each PAS passing .4 filter") + facet_grid(~nPAS)

```

Incorporate usage 

```{r}
metausage= MetaPAS %>% select(PAS, Human, Chimp)
BothRes_g_same40= BothRes %>% select(Human_PAS, gene, Human_DiffDom, Chimp_DiffDom) %>% rename(PAS=Human_PAS) %>% gather("Species", "Diff", -gene, -PAS) %>% filter(gene %in% BothRes_40_same$gene) %>% inner_join(metausage, by="PAS")
```

Plot the usage correlation 

```{r}
ggplot(BothRes_g_same40, aes(x=Human, y=Chimp, col=Diff)) + geom_point(alpha=.3)+ facet_grid(~Species)
```

```{r}
BothRes_g_same40s= BothRes_g_same40 %>% spread(Species, Diff) %>% gather("species", "usage", -PAS,-Chimp_DiffDom, -Human_DiffDom, -gene)

ggplot(BothRes_g_same40s, aes(x=Human_DiffDom,y=Chimp_DiffDom, col=usage)) + geom_point() + facet_grid(~species)
```

There is more of a usage effect in the Human usage. 

Write out the results to use for downstream analysis: 

```{bash,eval=F}
mkdir ../data/DomStructure_4/
```

```{r}
write.table(BothRes_40, "../data/DomStructure_4/InclusiveDominantPASat4.txt", col.names =T, row.names = F, quote = F)
```



Look at this metric for all cutoffs: 

```{r}

BothRes=HumanRes %>% inner_join(ChimpRes,by="gene")

BothRes_10=BothRes %>% filter(Chimp_DiffDom >=0.1 | Human_DiffDom>=0.1) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=10) 
BothRes_20=BothRes %>% filter(Chimp_DiffDom >=0.2 | Human_DiffDom>=0.2) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=20)
BothRes_30=BothRes %>% filter(Chimp_DiffDom >=0.3 | Human_DiffDom>=0.3) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=30)
BothRes_40=BothRes %>% filter(Chimp_DiffDom >=0.4 | Human_DiffDom>=0.4) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=40)
BothRes_50=BothRes %>% filter(Chimp_DiffDom >=0.5 | Human_DiffDom>=0.5) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=50)
BothRes_60=BothRes %>% filter(Chimp_DiffDom >=0.6 | Human_DiffDom>=0.6) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=60)
BothRes_70=BothRes %>% filter(Chimp_DiffDom >=0.7 | Human_DiffDom>=0.7) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=70)
BothRes_80=BothRes %>% filter(Chimp_DiffDom >=0.8 | Human_DiffDom>=0.8) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=80)
BothRes_90=BothRes %>% filter(Chimp_DiffDom >=0.9 | Human_DiffDom>=0.9) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=90)

BothResAll=BothRes_10 %>% bind_rows(BothRes_20) %>% bind_rows(BothRes_30) %>% bind_rows(BothRes_40) %>% bind_rows(BothRes_50) %>% bind_rows(BothRes_60) %>% bind_rows(BothRes_70) %>% bind_rows(BothRes_80) %>% bind_rows(BothRes_90)
```


I want the number and proportion of same vs different at every cutoff: 

```{r}

MetaPAS_genes= MetaPAS %>% group_by(gene) %>% summarise(n())
BothResAll_g= BothResAll %>% group_by(cut, Set) %>% summarise(nEach=n()) %>% ungroup() %>% group_by(cut) %>% mutate(nDom=sum(nEach), PropSame=nEach/nDom) %>% filter(Set=="Same") %>%ungroup() %>%  mutate(NTested=nrow(MetaPAS_genes), PropDom=nDom/NTested)

BothResAll_g$cut=as.factor(BothResAll_g$cut)
```


Plot:

```{r}

numberDom=ggplot(BothResAll_g,aes(x=cut,y=PropSame,fill=cut)) + geom_bar(stat="identity") +geom_text(aes(label=nDom), position=position_dodge(width=0.9), vjust=2) + scale_fill_brewer(palette = "RdYlBu")+theme(legend.position = "none") + labs(title="Most genes with a domiant PAS share the same dominant PAS", y="Proportion of Genes in Set", x="Domianance Cutoff")

numberDom
```

```{r}

propDom=ggplot(BothResAll_g,aes(x=cut,y=PropDom,fill=cut)) + geom_bar(stat="identity") +geom_text(aes(label=nDom), position=position_dodge(width=0.9), vjust=2) + scale_fill_brewer(palette = "RdYlBu")+theme(legend.position = "none") + labs(title="Proportion of Tested Genes with a Dominant PAS", y="Proportion of Tested Genes", x="Domianance Cutoff")
```

```{r}
library(cowplot)

plot_grid(propDom, numberDom)
```

```{r}

pdf("../output/PropSamesdom.pdf", height=8, width=8)
numberDom
dev.off()

```


```{r PropSamesdom, include=FALSE, dev="pdf", fig.height=8, fig.width=11, crop=FALSE}
numberDom
```
