---
title: "Differential splicing Without bad individuals"
author: "Briana Mittleman"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I will run the differential splicing pipeline I previously ran but without Chimp 4973 and Human 18498. 


```{r}
library(tidyverse)
library(reshape2)
```



I now have my RNA seq for each species. I can write a script that runs the junctions for each species.  


```{bash,eval=F}
mkdir ../Human/data/RNAseq/sort_removebad
mkdir ../Human/data/RNAseq/DiffSplice_removBad

mkdir ../Chimp/data/RNAseq/sort_removebad
mkdir ../Chimp/data/RNAseq/DiffSplice_removeBad
#move good files to this
sbatch converBam2Junc_removeBad.sh
```

```{bash,eval=F}
sbatch quantJunc_removeBad.sh
```


Now I need to do reciprocal liftover with the clusters.  

* /project2/gilad/briana/Comparative_APA/Human/data/RNAseq/DiffSplice_removeBad/
* /project2/gilad/briana/Comparative_APA/Chimp/data/RNAseq/DiffSplice_removeBad/

Chain files are in /data/chainFiles/
 * panTro5ToHg38.over.chain  
 * hg38ToPanTro5.over.chain  
 
 I first need to make bedfiles with these clusters.  
 
 The clusters all have _NA I dont this this is correct.  
 
 
```{bash,eval=F}

gunzip ../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_perind.counts.gz

gunzip ../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_perind.counts.gz


python cluster2bed.py ../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_perind.counts ../Human/data/RNAseq/DiffSplice_removeBad/humanJunc.bed

python cluster2bed.py ../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_perind.counts ../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc.bed


#I need to name the clusters before I can do the lift. (this is like the naming in apa 1-n(clusters))
python nameClusters.py ../Human/data/RNAseq/DiffSplice_removeBad/humanJunc.bed ../Human/data/RNAseq/DiffSplice_removeBad/humanJuncNamed.bed

python nameClusters.py ../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc.bed ../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJuncNamed.bed

sbatch clusterLiftprimary_removebad.sh

sbatch clusterLiftReverse_removebad.sh
```
 
 
 Evaluate results:
 
 (this code is from the lift for the PAS)
```{r}
unliftedH=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_unlifted.bed",stringsAsFactors = F) %>% nrow()
unliftedC=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_unlifted.bed",stringsAsFactors = F) %>% nrow()

liftedH=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp.bed",stringsAsFactors = F) %>% nrow()
liftedC=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman.bed",stringsAsFactors = F) %>% nrow()

primaryUnC=c("Chimp","Unlifted", unliftedC)
primaryUnH=c("Human","Unlifted", unliftedH)

primaryLH=c("Human","Lifted", liftedH)
primaryLC=c("Chimp","Lifted", liftedC)

header=c("species", "liftStat", "PAS")
primaryDF= as.data.frame(rbind(primaryLH,primaryLC, primaryUnH,primaryUnC)) 
colnames(primaryDF)=header


primaryDF$PAS=as.numeric(as.character(primaryDF$PAS)) 



primaryDF= primaryDF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)
```


```{r}
ggplot(primaryDF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results", y="Isoforms")

ggplot(primaryDF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results")
```



Look at the lifted: 

```{r}
OriginalHuman=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc.bed",stringsAsFactors = F) 
liftedHuman=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp.bed",stringsAsFactors = F) 

OriginalChimp=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc.bed",stringsAsFactors = F)
liftedChimp=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman.bed",stringsAsFactors = F)
```

Reverse lift: 
```{r}
re_unliftedH=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp_B2Human_unlifted.bed",stringsAsFactors = F) %>% nrow()
re_unliftedC=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman_B2Chimp_unlifted.bed",stringsAsFactors = F) %>% nrow()

re_liftedH=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp_B2Human.bed",stringsAsFactors = F) %>% nrow()
re_liftedC=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman_B2Chimp.bed",stringsAsFactors = F) %>% nrow()

re_UnC=c("Chimp","Unlifted", re_unliftedC)
re_UnH=c("Human","Unlifted", re_unliftedH)

re_LH=c("Human","Lifted", re_liftedH)
re_LC=c("Chimp","Lifted", re_liftedC)

header=c("species", "liftStat", "PAS")
re_DF= as.data.frame(rbind(re_LH,re_LC, re_UnH,re_UnC)) 
colnames(re_DF)=header


re_DF$PAS=as.numeric(as.character(re_DF$PAS)) 



re_DF= re_DF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)
```


```{r}
ggplot(re_DF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Reverse Liftover Results", y="Isoforms")
ggplot(re_DF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2")+ labs(title="Reverse Liftover Results")
```

How many lifted both ways?
```{r}
#human
re_liftedH/nrow(OriginalHuman)
#chimp
re_liftedC/nrow(OriginalChimp)
```



The next step will be to find the corresponding clusters. This is important because I will need to get the quantifications for the same introns and clusters. To do this I will need to write code that looks for the intron location from the primary lift in the reverse lift. 

For now I will only look at those introns identified in both species. I need to do this because I need junctions we have quantifications for in both species.  

I can make files with the human and chimp coordintats for the clusters that lift both ways. I will have to number each cluster 


Human cluser: 
```{r}

humanRevlift=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp_B2Human.bed",stringsAsFactors = F,col.names = c("Hchr","Hstart", "Hend", "cluster", "score", "strand")) %>% select(-strand)

#number clusters
humanRevlift %>% select(cluster) %>% unique() %>% nrow()

humanRevlift$score=as.character(humanRevlift$score)
humanRevlift= humanRevlift %>%  mutate(Name=paste("Human", score, sep="_")) %>% select(-score)


humanInChimp=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_inChimp.bed",stringsAsFactors = F,col.names = c("Cchr","Cstart", "Cend", "cluster", "score", "strand"))%>% select(-strand)
humanInChimp$score=as.character(humanInChimp$score)
humanInChimp= humanInChimp %>%  mutate(Name=paste("Human", score, sep="_")) %>% select(-score)


humanliftedBoth=humanRevlift %>% inner_join(humanInChimp, by=c("cluster", "Name"))
```

Chimp clusters: 

```{r}
chimpRevLift=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman_B2Chimp.bed",stringsAsFactors = F,col.names = c("Cchr","Cstart", "Cend", "cluster", "score", "strand")) %>% select(-strand)
chimpRevLift$score=as.character(chimpRevLift$score)
chimpRevLift= chimpRevLift %>%  mutate(Name=paste("Chimp", score, sep="_")) %>% select(-score)
chimpRevLift %>% select(cluster) %>% unique() %>% nrow()


chimpInHuman=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_inHuman.bed",stringsAsFactors = F,col.names = c("Hchr","Hstart", "Hend", "cluster", "score", "strand"))%>% select(-strand) 
chimpInHuman$score=as.character(chimpInHuman$score)
chimpInHuman= chimpInHuman %>%  mutate(Name=paste("Chimp", score, sep="_")) %>% select(-score)

chimpliftedBoth=chimpRevLift %>% inner_join(chimpInHuman, by=c("cluster", "Name"))
```


Try to join these by the human and chimp coordinates  

```{r}
AllClusters=chimpliftedBoth %>% inner_join(humanliftedBoth, by=c("Cchr", "Cstart","Cend", "Hchr", "Hstart", "Hend")) %>% mutate(ChimpName=paste(Cchr,Cstart,Cend, cluster.x, sep=":" ),HumanName=paste(Hchr,Hstart,Hend, cluster.y, sep=":" ) )

nrow(AllClusters)

AllClusters %>% select(cluster.x) %>% unique() %>% nrow()
AllClusters %>% select(cluster.y) %>% unique() %>% nrow()
```
This means there are ~1k isoforms from about 500 genes.  
I will have to go back and figure out how to call clusters for more genes.  

I need to reformat these back into the counts format. 

```{r}
#chr1:17055:17233:clu_1

AllClustersNames=AllClusters %>% select(HumanName, ChimpName)

ChimpCluster=read.table("../Chimp/data/RNAseq/DiffSplice_removeBad/chimpJunc_perind_numers.counts.gz") %>% rownames_to_column(var="ChimpName")
FilteredChimpCluster= ChimpCluster %>% inner_join(AllClustersNames, by="ChimpName")

#map human onto these

HumanCluster=read.table("../Human/data/RNAseq/DiffSplice_removeBad/humanJunc_perind_numers.counts.gz") %>% rownames_to_column(var="HumanName")
FilteredClusterBoth=HumanCluster %>% inner_join(FilteredChimpCluster, by="HumanName") %>% select(-ChimpName) 

FilteredClusterBothfixed=FilteredClusterBoth[!duplicated(FilteredClusterBoth$HumanName),]






#create group file- this should have the name of the bams and the group
Bams=as.data.frame(colnames(FilteredClusterBothfixed)) %>% mutate(Species=ifelse(grepl("H",colnames(FilteredClusterBothfixed)), "Human", "Chimp")) %>% slice(2:n())

# mkdir ../data/DiffSplice_removeBad/

write.table(Bams, "../data/DiffSplice_removeBad/groups_file.txt", col.names = F, row.names = F, quote = F, sep="\t" )


write.table(FilteredClusterBothfixed, "../data/DiffSplice_removeBad/BothSpec_perind.counts", col.names = T, row.names = F, quote = F, sep="\t" )


```

Remove the first name in header and zip the file: 

**(manually)**
```{bash,eval=F}
vi ../data/DiffSplice_removeBad/BothSpec_perind.counts

gzip ../data/DiffSplice_removeBad/BothSpec_perind.counts
```

Run leafcutter with python 2 

```{bash,eval=F}
sbatch DiffSplice_removebad.sh

```

Look at results: 

```{r}
sig=read.table("../data/DiffSplice_removeBad/_cluster_significance.txt",sep="\t" ,header =T,stringsAsFactors = F) %>% filter(status=="Success") 

sig$p.adjust=as.numeric(as.character(sig$p.adjust))


qqplot(-log10(runif(nrow(sig))), -log10(sig$p.adjust),ylab="-log10 Total Adjusted Leafcutter pvalue", xlab="-log 10 Uniform expectation", main="Leafcutter Differential Splicing")
abline(0,1)

sig %>% filter(p.adjust < .05 ) %>% nrow()


```

