---
title: "Differential Information Content"
author: "Briana Mittleman"
date: "4/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library("edgeR")
library(cowplot)
library(coin)
library("R.utils")
library(rcompanion)
library("scales")
library(gplots)
library("limma")
library("RColorBrewer")
```


##Get info at indiv. 
In order to look for genes with differential information content (dIC) between species, I need to create a method for calculating indiviudal level info content measures. 

It will be like what I did previously but I will work with individual ratios rather than means. I will use python and wrap the code around each individual. I can make it in a long dataframe at first to write out the gene, individual, and values, then I can spread the output in R after.  

It will take in the ratio file.  

```{r}
PASuse=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", stringsAsFactors = F ,header = T)
```


```{bash,eval=F}
mkdir ../data/DoubleFilterUsageNumeric
mkdir ../data/IndInfoContent
```

Chimp usage:  

```{r}
ChimpAnno=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndC=colnames(ChimpAnno)[9:ncol(ChimpAnno)]

ChimpUsage=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt", col.names = IndC)
ChimpUsage_nuclear= ChimpUsage %>% select(contains("_N"))

ChimpUsage_anno=as.data.frame(cbind(ChimpAnno[,1:8],ChimpUsage_nuclear )) %>% filter(PAS %in% PASuse$PAS) %>% select(-chr, -start, -end, -strand, -loc, -disc)
```


Human Usage:

```{r}
HumanAnno=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndH=colnames(HumanAnno)[9:ncol(HumanAnno)]

HumanUsage=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt", col.names = IndH)
HumanUsage_nuclear= HumanUsage %>% select(contains("_N"))

HumanUsage_anno=as.data.frame(cbind(HumanAnno[,1:8],HumanUsage_nuclear )) %>% filter(PAS %in% PASuse$PAS) %>% select(-chr, -start, -end, -strand, -loc, -disc)
```


```{r}
bothUsage=HumanUsage_anno %>% inner_join(ChimpUsage_anno, by=c("gene","PAS"))

write.table(bothUsage,"../data/DoubleFilterUsageNumeric/BothSpecies_NumericUsage_DoubleFilt.txt", col.names = T, row.names = F, quote = F)

write.table(ChimpUsage_anno,"../data/DoubleFilterUsageNumeric/Chimp_NumericUsage_DoubleFilt.txt", col.names = T, row.names = F, quote = F)

write.table(HumanUsage_anno,"../data/DoubleFilterUsageNumeric/Human_NumericUsage_DoubleFilt.txt", col.names = T, row.names = F, quote = F)
```

```{r}
bothUsage %>% select(gene) %>% unique() %>% nrow

ChimpUsage_anno%>% group_by(gene) %>% summarise(nPAS=n()) %>% filter(nPAS>1) %>% nrow


HumanUsage_anno %>% group_by(gene) %>% summarise(nPAS=n()) %>% filter(nPAS>1) %>% nrow
```


test the code: 

```{bash,eval=F}
head ../data/DoubleFilterUsageNumeric/Human_NumericUsage_DoubleFilt.txt >  ../data/IndInfoContent/test.txt

python InfoContentbyInd.py ../data/IndInfoContent/test.txt ../data/IndInfoContent/testout.txt


head ../data/DoubleFilterUsageNumeric/BothSpecies_NumericUsage_DoubleFilt.txt >  ../data/IndInfoContent/bothtest.txt

python InfoContentbyInd.py ../data/IndInfoContent/bothtest.txt ../data/IndInfoContent/bothtestOut.txt
```

```{r}
test=read.table("../data/IndInfoContent/testout.txt", col.names = c("gene", "ind", "shannon", "eq", "simpson"), stringsAsFactors = F)
testBoth=read.table("../data/IndInfoContent/bothtestOut.txt",col.names = c("gene", "ind", "shannon", "eq", "simpson"), stringsAsFactors = F)
```

Try it:  

```{bash,eval=F}
python InfoContentbyInd.py ../data/DoubleFilterUsageNumeric/Chimp_NumericUsage_DoubleFilt.txt  ../data/IndInfoContent/ChimpInfoContent_long.txt


python InfoContentbyInd.py ../data/DoubleFilterUsageNumeric/Human_NumericUsage_DoubleFilt.txt  ../data/IndInfoContent/HumanInfoContent_long.txt
```


```{r}
chimpAllInfoLong=read.table("../data/IndInfoContent/ChimpInfoContent_long.txt", col.names = c("gene", "ind", "shannon", "eq", "simpson"), stringsAsFactors = F ) %>% mutate(species="Chimp")

humanAllInfoLong=read.table("../data/IndInfoContent/HumanInfoContent_long.txt", col.names = c("gene", "ind", "shannon", "eq", "simpson"), stringsAsFactors = F )%>% mutate(species="Human")


```

##Simpson  
Spread the simpson to do differences between species:  

```{r}
humanSimp= humanAllInfoLong %>% select(-eq,-shannon,- species) %>% spread(ind, simpson)

chimpSimp= chimpAllInfoLong %>% select(-eq,-shannon,- species) %>% spread(ind, simpson)

BothSimp= chimpSimp %>% inner_join(humanSimp, by="gene")
```

design matrix:  

```{r}

humD=humanAllInfoLong %>% select(ind, species) %>% unique() 
chimpD=chimpAllInfoLong %>% select(ind, species) %>% unique()
design=bind_rows(chimpD,humD)
design$species=as.factor(design$species)
design
```

Run some QC on this like we would be DE.  

```{r}
SimpMatric=BothSimp %>% column_to_rownames(var="gene")

head(SimpMatric)

colors <- colorRampPalette(c(brewer.pal(9, "Blues")[1],brewer.pal(9, "Blues")[9]))(100)

pal <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), brewer.pal(12, "Set3"))
labels <- paste(design$species,design$ind, sep=" ")


```

```{r}
#PCA function (original code from Julien Roux)
#Load in the plot_scores function
plot_scores <- function(pca, scores, n, m, cols, points=F, pchs =20, legend=F){
  xmin <- min(scores[,n]) - (max(scores[,n]) - min(scores[,n]))*0.05
  if (legend == T){ ## let some room (35%) for a legend                                                                                                                                                 
    xmax <- max(scores[,n]) + (max(scores[,n]) - min(scores[,n]))*0.50
  }
  else {
    xmax <- max(scores[,n]) + (max(scores[,n]) - min(scores[,n]))*0.05
  }
  ymin <- min(scores[,m]) - (max(scores[,m]) - min(scores[,m]))*0.05
  ymax <- max(scores[,m]) + (max(scores[,m]) - min(scores[,m]))*0.05
  plot(scores[,n], scores[,m], xlab=paste("PC", n, ": ", round(summary(pca)$importance[2,n],3)*100, "% variance explained", sep=""), ylab=paste("PC", m, ": ", round(summary(pca)$importance[2,m],3)*100, "% variance explained", sep=""), xlim=c(xmin, xmax), ylim=c(ymin, ymax), type="n")
  if (points == F){
    text(scores[,n],scores[,m], rownames(scores), col=cols, cex=1)
  }
  else {
    points(scores[,n],scores[,m], col=cols, pch=pchs, cex=1.3)
  }
}
```
```{r}
# Clustering (original code from Julien Roux)
cors <- cor(SimpMatric, method="spearman", use="pairwise.complete.obs")

```
Correaltion plot
```{r}

heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels,ColSideColors=pal[as.integer(as.factor(design$species))], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```

PCA: 

```{r}
# Perform PCA

pca_genes <- prcomp(t(SimpMatric), scale = F)
scores <- pca_genes$x


#Make PCA plots with the factors colored by species

### PCs 1 and 2 Raw Data
for (n in 1:1){
  col.v <- pal[as.integer(design$species)]
  plot_scores(pca_genes, scores, n, n+1, col.v)
}
```
```{r}
plotDensities(SimpMatric, col=pal[as.numeric(design$species)], legend="topleft")
```


First pass I will just run a wilcoxan test and correct the pvalues.  

```{r}
BothSimpM=as.matrix(SimpMatric)
outPut=as.data.frame(cbind(gene=BothSimp$gene, pval=rep(0,nrow(BothSimp)), effect=rep(0,nrow(BothSimp))))
outPut$pval=as.numeric(as.character(outPut$pval))
outPut$effect=as.numeric(as.character(outPut$effect))
for (i in seq(1:nrow(BothSimp))){
  chimp=as.vector(BothSimpM[i,1:6])
  human=as.vector(BothSimpM[i,7:11])
  pval=wilcox.test(human, chimp, alt = "two.sided", paired=F)$p.value
  #effect=wilcoxonR(x =as.vector(BothSimpM[i,1:11]),  g = design$species)
  outPut[i,2] <- pval
  #outPut[i,3] <- effect
}
```



```{r}
padj=p.adjust(outPut$pval, method ="fdr")
#plot(sort(padj))



outPutad=outPut %>% mutate(fdr=padj)

outPutadSig= outPutad %>% filter(fdr<0.05)
```


Plot pvalues:  

```{r}
plot(sort(outPutad$pval))

summary(outPutad$pval)
```



```{r}
plot(sort(outPutad$fdr),main="dIC FDR distribution", ylab="Sorted FDR")
```



Look at mean and variance:

```{r}
humanSimp= humanAllInfoLong %>% dplyr::select(-eq,-shannon)
chimpSimp= chimpAllInfoLong %>% dplyr::select(-eq,-shannon)

BothSimplong=humanSimp %>% bind_rows(chimpSimp)

SimpMeanVar= BothSimplong %>% group_by(species, gene) %>% summarise(mean=mean(simpson), var=var(simpson))

ggplot(SimpMeanVar, aes(y=mean, x=species))+ geom_boxplot()

ggplot(SimpMeanVar, aes(y=var, x=species))+ geom_boxplot()
```


```{r}
SimpMeanVar= BothSimplong %>% group_by(species, gene) %>% summarise(mean=mean(simpson)) %>% spread(species, mean)
SimpHuman=read.table("../data/InfoContent/Human_SimpsonInfoContent.txt", header = T, stringsAsFactors = F) %>% rename(simpson_Human=simpson) %>% mutate(simpOpp_Human=1-simpson_Human)
SimpChimp=read.table("../data/InfoContent/Chimp_SimpsonInfoContent.txt", header = T, stringsAsFactors = F)%>% rename(simpson_Chimp=simpson)%>% mutate(simpOpp_Chimp=1-simpson_Chimp)

BothSimp= SimpHuman %>% inner_join(SimpChimp, by=c("gene", "numPAS")) %>% filter(numPAS > 1)



allSImpson=BothSimp %>% inner_join(SimpMeanVar, by="gene")

cor.test(allSImpson$simpOpp_Human, allSImpson$Human)

cor.test(allSImpson$simpOpp_Chimp, allSImpson$Chimp)
```

Explore the genes:  
```{r}
#outPutadSig
allSImpson_sig=allSImpson %>% mutate(sIC=ifelse(gene %in% outPutadSig$gene, "Yes","No"))


ggplot(allSImpson_sig, aes(x=Human, y=Chimp, col=sIC)) + geom_point()

ggplot(allSImpson_sig, aes(x=Human, y=Chimp, col=sIC)) + geom_density_2d()

```


Ok this makes sense. Let me look at medians.  

```{r}
SimpMed= BothSimplong %>% group_by(species, gene) %>% summarise(Median=median(simpson)) %>% spread(species, Median)

allSImpsonMed=BothSimp %>% inner_join(SimpMed, by="gene") %>% mutate(sIC=ifelse(gene %in% outPutadSig$gene, "Yes","No"))


med5=ggplot(allSImpsonMed, aes(x=Human, y=Chimp, col=sIC)) + geom_point() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="5% FDR dIC plotted on median values")+scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))

med5

mean5=ggplot(allSImpsonMed, aes(x=simpOpp_Human, y=simpOpp_Chimp, col=sIC)) + geom_point() + labs(x="Human Simpson (calculated on mean usage)", y="Chimp Simpson (calculated on mean usage)", title="5% FDR dIC plotted on \nvalues calculated by mean usage") + scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))
mean5
ggplot(allSImpsonMed, aes(x=Human, y=Chimp, col=sIC)) + geom_density_2d()
```
```{r}
allSImpsonMed %>% group_by(sIC) %>% summarise(n=n())
```

Write out to look at overlap with phenotypes  

```{r}
write.table(allSImpsonMed,"../data/IndInfoContent/SimpsonMedianSignificance.txt", col.names = T, row.names = F, quote=F)
```

This is an underpowered analysis because with pvalue ranges are restricted with low samples sizes. I will also write out a set at 10% FDR.  

```{r}
outPutadSig10= outPutad %>% filter(fdr<0.1)

nrow(outPutadSig10)
allSImpsonMed10=BothSimp %>% inner_join(SimpMed, by="gene") %>% mutate(sIC10=ifelse(gene %in% outPutadSig10$gene, "Yes","No"))


write.table(allSImpsonMed10,"../data/IndInfoContent/SimpsonMedianSignificance_10FDR.txt", col.names = T, row.names = F, quote=F)
```

```{r}

med10=ggplot(allSImpsonMed10, aes(x=Human, y=Chimp, col=sIC10)) + geom_point() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="10% FDR dIC plotted on median values")+scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))
med10
mean10=ggplot(allSImpsonMed10, aes(x=simpOpp_Human, y=simpOpp_Chimp, col=sIC10)) + geom_point() + labs(x="Human Simpson (calculated on mean usage)", y="Chimp Simpson (calculated on mean usage)", title="10% FDR dIC plotted on values\n calculated by mean usage") + scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))
mean10

```


```{r}
plot_grid(med5,mean5,med10,mean10)
```


Color dIC only, dAPA, and both.  

```{r}
Meta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) 


Meta_genes= Meta %>% select(gene) %>% unique()

Meta_PAS=Meta %>%  select(PAS,gene)

dAPAGenes=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt", header = T, stringsAsFactors = F)
dAPAPAS=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header = T, stringsAsFactors = F) %>% inner_join(Meta, by=c("chr","start", "end","gene")) %>% select(PAS,gene,SigPAU2 ) 

dAPAPAS_genes= dAPAPAS %>% select(gene) %>% unique()

dAPATestedGenes= dAPAPAS  %>% select(gene) %>% unique() %>% mutate(dAPA=ifelse(gene %in% dAPAGenes$gene,"Yes", "No")) 

dAPAandDic= allSImpsonMed %>% inner_join(dAPATestedGenes,by="gene") %>% mutate(Both=ifelse(sIC=="Yes" & dAPA=="Yes", "Yes","No"),OnlyIC=ifelse(sIC=="Yes" & dAPA=="No", "Yes","No"),OnlyAPA=ifelse(sIC=="No" & dAPA=="Yes", "Yes","No"))

```


```{r}
ggplot(dAPAandDic, aes(x=Human, y=Chimp, col=Both)) + geom_point() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="Median IC colored by genes in both sets")+scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold")) + facet_grid(~Both)
```

```{r}
ggplot(dAPAandDic, aes(x=Human, y=Chimp, col=OnlyAPA)) + geom_point() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="Median IC colored by Only dAPA")+scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"))+ facet_grid(~OnlyAPA)
```
```{r}
ggplot(dAPAandDic, aes(x=Human, y=Chimp, col=OnlyIC)) + geom_point() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="Median IC colored by Only dIC")+scale_color_brewer(palette = "Set1")+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold")) + facet_grid(~OnlyIC)
```


plot together: 

```{r}


dAPAandDic_both= dAPAandDic %>% filter(Both=="Yes") %>% mutate(set="Both") %>% select(gene,Human, Chimp, set)


dAPAandDic_onlyIC= dAPAandDic %>% filter(OnlyIC=="Yes") %>% mutate(set="OnlyIC") %>% select(gene,Human, Chimp, set)

dAPAandDic_onlyAPA= dAPAandDic %>% filter(OnlyAPA=="Yes") %>% mutate(set="OnlyAPA") %>% select(gene,Human, Chimp, set)


dAPAandDic_g= dAPAandDic_both %>% bind_rows(dAPAandDic_onlyIC,dAPAandDic_onlyAPA)
```

```{r}

useCOl <- c("#d73027", "#4575b4","#FFB833")


colordots=ggplot(dAPAandDic_g, aes(x=Human, y=Chimp, col=set)) + geom_point(alpha=.8) + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="Median IC colored by dAPA structure")+scale_color_manual(values = useCOl)+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"), legend.position = "bottom")

colordots
```

```{r}
densityplot=ggplot(dAPAandDic_g, aes(x=Human, y=Chimp, col=set)) + geom_density2d() + labs(x="Median Human Simpson", y="Median Simpson Chimp", title="Median IC colored by dAPA structure")+scale_color_manual(values = useCOl)+theme(axis.text=element_text(size=10),axis.title=element_text(size=10,face="bold"),legend.position = "bottom" )
```


```{r}
plot_grid(colordots,densityplot)
```

