---
title: "Categories from Khan"
author: "Briana Mittleman"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will use the directional selection categories from Khan et al.  

```{r}
library(tidyverse)
```


Model from Khan et al. 

model.num.rna: : 1 = mRNA expression level pattern consistent with directional selection along human lineage, 2 = mRNA expression level pattern consistent with directional selection along chimpanzee lineage, 3 = undetermined pattern, 4 = patterns with no significant difference between mean expression levels; 5 = evidence for relaxation of constraint along human lineage, 6 = evidence of relaxation of constraint along chimpanzee lineage

model.num.protein: 1 = protein expression level pattern consistent with directional selection along human lineage, 2 = protein expression level pattern consistent with directional selection along chimpanzee lineage, 3 = undetermined pattern, 4 = patterns with no significant difference between mean expression levels; 5 = evidence for relaxation of constraint along human lineage, 6 = evidence of relaxation of constraint along chimpanzee lineage 


```{r}

KhanData=read.csv("../data/Khan_prot/Khan_TableS4.csv",stringsAsFactors = F)  %>% select(gene.symbol,contains("model") ) %>% rename("gene"=gene.symbol, "Protein"=model.num.protein, "RNA"=model.num.rna)

KhanData_g=KhanData %>% gather("Set", "Model", -gene)

KhanData_g$Model= as.factor(KhanData_g$Model)
```



```{r}
ggplot(KhanData_g, aes(x=Model, by=Set, fill=Set)) +geom_bar(stat="count",position = "dodge")


```
1. directional human 
2. directional in chimp 
3, undetermined
4. no mean difference
5. relaxed in human
6. related in chimp 

Proportion  

```{r}
KhanData_group=KhanData_g %>% group_by(Set, Model) %>% summarise(Nset=n(), Prortion=Nset/nrow(KhanData))

ggplot(KhanData_group, aes(x=Model, by=Set, fill=Set, y=Prortion)) +geom_bar(stat="identity",position = "dodge")
```

This is their results. I will overlap this with the genes I found differences in. 

```{r}
DiffIso=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header = T, stringsAsFactors = F) %>% inner_join(KhanData, by="gene")

nrow(DiffIso)
nrow(DiffIso %>% select(gene) %>% unique())


```
We have data for 12392 PAS in 2532 genes.  

I want to get this down to gene level

```{r}
GenewDiffIso= DiffIso %>% group_by(gene,SigPAU2) %>% summarise(nEach=n()) %>% filter(SigPAU2=="Yes")

KhanData_withAPAinfo= KhanData %>% mutate(dAPA=ifelse(gene %in%GenewDiffIso$gene, "Yes", "No" ))

```

Plot the proportion with a dAPA in each 
```{r}
KhanData_gwAPA=KhanData_withAPAinfo %>% gather("Set", "Model", -gene, -dAPA)
KhanData_gwAPA$Model= as.factor(KhanData_gwAPA$Model)

KhanData_gwAPA$Set=factor(KhanData_gwAPA$Set, levels=c("RNA", "Protein"))
ggplot(KhanData_gwAPA, aes(x=Model, by=dAPA, fill=dAPA)) +geom_bar(stat="count", position = "stack") + facet_grid(~Set) + scale_fill_brewer(palette = "Dark2") + labs(y="Number of Genes") + scale_x_discrete( labels=c("Selection Human","Selection Chimp","Undetermined","No mean difference","Relaxation in Human","Relaxation in Chimp"))+theme(axis.text.x=element_text(angle=90, hjust=0), text= element_text(size=16))
```

Look at the genes with directional selection:

```{r}
DirSelectionRNA=KhanData_gwAPA %>% filter(Set=="RNA", Model %in% c(1,2))

DirSelectionRNA %>% group_by(Model,dAPA) %>% summarise(n())

DirSelectionProt=KhanData_gwAPA %>% filter(Set=="Protein", Model %in% c(1,2))

DirSelectionProt %>% group_by(Model,dAPA) %>% summarise(n())
```

Relaxed selection:  

```{r}
RelSelectionRNA=KhanData_gwAPA %>% filter(Set=="RNA", Model %in% c(5,6))

RelSelectionRNA %>% group_by(Model,dAPA) %>% summarise(n())

RelSelectionProt=KhanData_gwAPA %>% filter(Set=="Protein", Model %in% c(5,6))

RelSelectionProt %>% group_by(Model,dAPA) %>% summarise(n())
```





Enrichement:  

Write a loop that gets the pvalue and enrichment for each of these:  


```{r}
Model=seq(1,6)
EnrichmentRNA=c()
PvalueRNA=c()
for (i in seq(1:6)){
  x=nrow(KhanData_gwAPA %>% filter(Set=="RNA", dAPA=="Yes", Model==i))
  m=nrow(KhanData_gwAPA %>% filter(Set=="RNA", Model==i))
  n=nrow(KhanData_gwAPA %>% filter(Set=="RNA", Model!=i))
  k=nrow(KhanData_gwAPA %>% filter(Set=="RNA", dAPA=="Yes"))
  N=nrow(KhanData_gwAPA %>% filter(Set=="RNA"))
  PvalueRNA=c(PvalueRNA, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichmentRNA=c(EnrichmentRNA, enrich)
}

EnrichProt=c()
PvalueProt=c()

for (i in seq(1:6)){
  x=nrow(KhanData_gwAPA %>% filter(Set=="Protein", dAPA=="Yes", Model==i))
  m=nrow(KhanData_gwAPA %>% filter(Set=="Protein", Model==i))
  n=nrow(KhanData_gwAPA %>% filter(Set=="Protein", Model!=i))
  k=nrow(KhanData_gwAPA %>% filter(Set=="Protein", dAPA=="Yes"))
  N=nrow(KhanData_gwAPA %>% filter(Set=="Protein"))
  PvalueProt=c(PvalueProt, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichProt=c(EnrichProt, enrich)
}




```


```{r}
EnrichmentResults=as.data.frame(cbind(Model, EnrichmentRNA,EnrichProt,PvalueRNA, PvalueProt))

EnrichmentG= EnrichmentResults %>% select(Model, EnrichmentRNA, EnrichProt) %>% rename("RNA"=EnrichmentRNA, "Protein"=EnrichProt) %>% gather("Set", "Enrichment", -Model)
PvalG= EnrichmentResults %>% select(Model, PvalueRNA, PvalueProt) %>% rename("RNA"=PvalueRNA, "Protein"=PvalueProt) %>% gather("Set", "Pvalue", -Model)

Alldata=EnrichmentG %>% inner_join(PvalG, by=c("Model","Set"))

Alldata$Set=factor(Alldata$Set, levels=c("RNA", "Protein"))
Alldata$Model=factor(Alldata$Model)

ggplot(Alldata,aes(x=Model, y=Enrichment,fill=Set)) + geom_bar(stat = "identity") + geom_hline(yintercept = 1) + geom_text(aes(label=round(Pvalue,2), vjust=0))+ facet_grid(~Set) + scale_fill_brewer(palette = "Dark2")  + scale_x_discrete( labels=c("Selection Human","Selection Chimp","Undetermined","No mean difference","Relaxation in Human","Relaxation in Chimp"))+theme(axis.text.x=element_text(angle=90, hjust=0), text= element_text(size=16)) +labs(title="Enrichment of dAPA genes in directional selection sets")
```


I need a way to relate this to something about APA  

Examples:  

Look at all of the relaxed 


```{r}
RelSelectionRNA %>% filter(dAPA=="Yes", Model=="6")
```

```{r}
RelSelectionProt%>% filter(dAPA=="Yes", Model=="6")
```

IRF5
OTUB1

```{r}
RelSelectionRNA %>% filter(dAPA=="Yes", Model=="5")
```

```{r}
DirSelectionRNA %>% filter(dAPA=="Yes", Model=="2")
```


```{r}
DirSelectionProt %>% filter(dAPA=="Yes", Model=="2")
```

