---
title: "Dominant PAS Location Nuclear Double Filter"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I want to find the most dominant PAS for each gene in each species. I am interested in genes where the dominant PAS in human and chimp are intronic vs utr respectively. I will do this first in the nuclear fraction.  

I will compare these genes with those identified in the differential APA analysis. This will be helpful to narrow down the genes I want to visualize.  

```{r}
library(workflowr)
library(tidyverse)
```

These are the PAS 
```{r}
allPAS= read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T) 
ChimpPASwMean =allPAS %>% dplyr::select(-Human)
HumanPASwMean =allPAS %>% dplyr::select(-Chimp)
```


Find the dominant PAS per gene:

I will remove genes with ties for now
```{r}
Chimp_Dom= ChimpPASwMean %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom= HumanPASwMean %>% 
  group_by(gene) %>% 
  top_n(1, Human) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)


#merge

BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")
```

Look at how many have the same dominat and where these are:  

```{r}
SameDom=BothDom %>% filter(ChimpPAS==HumanPAS) 

ggplot(SameDom, aes(x=HumanLoc))+ geom_histogram(stat="count") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species")

```   


Plot this as boxplots as well. 

```{r}
SameDom_gather=SameDom %>% dplyr::select(gene, HumanLoc, Chimp,Human) %>% gather(species, value, -c(gene,HumanLoc))

ggplot(SameDom_gather,aes(x=HumanLoc, y=value,fill=species)) + geom_boxplot() + scale_fill_brewer(palette = "Dark2") + labs(x="PAS Location", y="Mean Usage accross individuals", title="Mean usage for Genes with matching Dominant PAS")
```
Write out the intronic for location analysis.  


```{bash,eval=F}
mkdir ../data/DominantPAS_DF
```

```{r}
SameDom_intron=SameDom %>% filter(ChimpLoc=="intron")
write.table(SameDom_intron, "../data/DominantPAS_DF/SameDominantIntronic.txt", col.names = T, row.names = F, quote = F)
```




Different PAS but in the same location:   
```{r}
DiffDom=BothDom %>% filter(ChimpPAS!=HumanPAS) 

ggplot(DiffDom,aes(x=HumanLoc)) +  geom_histogram(stat="count") + labs(x="PAS Location", y= "Number of Genes", title="Dominat PAS in same genic location but different PAS")

```
```{r}
DiffDom_g= DiffDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(DiffDom_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS") + scale_fill_brewer(palette = "Dark2")+theme(legend.position='bottom')
```



This is interesting but could be based on the annotations. I will look at the genes with human in intron and chimp in 3' UTR.  

```{r}
DiffDom_humanIntronChimpUTR= DiffDom %>% filter(ChimpLoc=="utr3", HumanLoc=="intron")

nrow(DiffDom_humanIntronChimpUTR)
```

Opposite Direction
```{r}

DiffDom_humanUTRChimpInton= DiffDom %>% filter(ChimpLoc=="intron", HumanLoc=="utr3")
nrow(DiffDom_humanUTRChimpInton)
```


```{r}
prop.test(x=c(nrow(DiffDom_humanIntronChimpUTR),nrow(DiffDom_humanUTRChimpInton)), n=c(nrow(DiffDom), nrow(DiffDom)))
```

I will look to see if these genes are those I see with differential APA.  



```{r}
write.table(DiffDom_humanIntronChimpUTR, "../data/DominantPAS_DF/Nuclear_HumanIntronicChimpUTR.txt", col.names = T, row.names = F, quote = F)
write.table(DiffDom_humanUTRChimpInton, "../data/DominantPAS_DF/Nuclear_HumanUTRChimpIntronic.txt", col.names = T, row.names = F, quote = F)

write.table(SameDom, "../data/DominantPAS_DF/Nuclear_SameDom.txt", col.names = T, row.names = F, quote = F)
```
