---
title: "Dominant PAS after Normalization"
author: "Briana Mittleman"
date: "4/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am worried that the different dominance structure is effected by mena and variance differences. To get around this. I want to use the normalized values for each PAS from the leafcutter model. I will use the nuclear differences.  

```{r}
library(reshape2)
library(tidyverse)
```

Load PAS

```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% mutate(chrom=paste(chr,start,end,gene, sep=":"))
```

```{r}
effectsize=read.table("../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('chrom',  'logef' ,'ChimpNorm', 'HumanNorm','deltaPAU')) %>% filter(chrom != "intron") %>% inner_join(PAS, by="chrom")

effectsize$HumanNorm=as.numeric(effectsize$HumanNorm)


effectsize$ChimpNorm=as.numeric(effectsize$ChimpNorm)
```

Look at the correlation between raw usage and normalized  

```{r}
ggplot(effectsize,aes(x=HumanNorm, y=Human)) + geom_point(alpha=.5) + geom_abline(aes(slope=1, intercept=0),col="red")

summary(lm(effectsize$HumanNorm~effectsize$Human))
```

```{r}
ggplot(effectsize,aes(x=ChimpNorm, y=Chimp)) + geom_point(alpha=.5) + geom_abline(aes(slope=1, intercept=0),col="red")

summary(lm(effectsize$ChimpNorm~effectsize$Chimp))
```

Looks like more change in the chimps.  


##Pick dominant PAS  

```{r}
Chimp_Dom= effectsize %>%
  group_by(gene) %>%
  arrange(desc(ChimpNorm)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,ChimpNorm) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom= effectsize %>%
  group_by(gene) %>%
  arrange(desc(HumanNorm)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,HumanNorm) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)

BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")
```



```{r}
SameDom= BothDom %>% filter(ChimpPAS==HumanPAS)
nrow(SameDom)
```

6006 have the same dominant PAS  

```{r}
SameDom_g= SameDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(SameDom_g, aes(x=Location, by=Species, fill=Species))+ geom_bar(stat="count",position = "Dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species\normalized usage")+ scale_fill_brewer(palette = "Dark2")
```
```{r}
DiffDom=BothDom %>% filter(ChimpPAS!=HumanPAS)
nrow(DiffDom)
```

This gives more genes with different dominant PAS  

```{r}
DiffDom_g= DiffDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


ggplot(DiffDom_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS using normalized usage\n (n=2416)") + scale_fill_brewer(palette = "Dark2")
```
This flips the distribution. It over corrects.  


###Dominant from original usage in filtered set  

```{r}
Chimp_Dom_Nonnorm= effectsize %>%
  group_by(gene) %>%
  arrange(desc(Chimp)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom_Nonnorm= effectsize %>%
  group_by(gene) %>%
  arrange(desc(Human)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)

BothDom_Nonnorm= Chimp_Dom_Nonnorm %>% inner_join(Human_Dom_Nonnorm,by="gene")
```

```{r}
SameDom_Nonnorm= BothDom_Nonnorm %>% filter(ChimpPAS==HumanPAS)
nrow(SameDom_Nonnorm)
```

6477 have the same dominant PAS  

```{r}
SameDom_Nonnorm_g= SameDom_Nonnorm %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(SameDom_Nonnorm_g, aes(x=Location, by=Species, fill=Species))+ geom_bar(stat="count",position = "Dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species\non normalized usage")+ scale_fill_brewer(palette = "Dark2")
```


```{r}
DiffDom_Nonnorm=BothDom_Nonnorm %>% filter(ChimpPAS!=HumanPAS)
nrow(DiffDom_Nonnorm)
```


```{r}
DiffDom_Nonnorm_g= DiffDom_Nonnorm %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


ggplot(DiffDom_Nonnorm_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS using non normalized usage (n=2067)") + scale_fill_brewer(palette = "Dark2")
```
