---
title: "Dominant PAS in Total Fraction"
author: "Briana Mittleman"
date: "12/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this analysis I want to find the most dominant PAS for each gene in each species. I am interested in genes where the dominant PAS in human and chimp are intronic vs utr respectively. 

I will compare these genes with those identified in the differential APA analysis. This will be helpful to narrow down the genes I want to visualize.  

```{r}
library(workflowr)
library(tidyverse)
```

These are the PAS 
```{r}
ChimpPAS= read.table("../data/Pheno_5perc/Chimp_Pheno_5perc.txt", header = T) %>% dplyr::select(-contains("_N"))


HumanPAS= read.table("../data/Pheno_5perc/Human_Pheno_5perc.txt", header = T) %>% dplyr::select(-contains("_N"))
```

Prepare the mean vector: 

```{r}
ChimpMean=rowMeans(ChimpPAS[,9:ncol(ChimpPAS)])

ChimpPASwMean=cbind(ChimpPAS[,1:8],ChimpMean)


HumanMean=rowMeans(HumanPAS[,9:ncol(HumanPAS)])

HumanPASwMean=cbind(HumanPAS[,1:8],HumanMean)
```

Find the dominant PAS per gene:

I will remove genes with ties for now
```{r}
Chimp_Dom= ChimpPASwMean %>%
  group_by(gene) %>%
  top_n(1,ChimpMean) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,ChimpMean) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom= HumanPASwMean %>% 
  group_by(gene) %>% 
  top_n(1, HumanMean) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,HumanMean) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)


#merge

BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")
```

Look at how many have the same dominat and where these are:  

```{r}
SameDom=BothDom %>% filter(ChimpPAS==HumanPAS,HumanLoc!="008559") 

ggplot(SameDom, aes(x=HumanLoc))+ geom_histogram(stat="count") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species")

```
Plot this as boxplots as well. 

```{r}
SameDom_gather=SameDom %>% dplyr::select(gene, HumanLoc, ChimpMean,HumanMean) %>% gather(species, value, -c(gene,HumanLoc))

ggplot(SameDom_gather,aes(x=HumanLoc, y=value,fill=species)) + geom_boxplot() + scale_fill_brewer(palette = "Dark2") + labs(x="PAS Location", y="Mean Usage accross individuals", title="Mean usage for Genes with matching Dominant PAS")
```


Different PAS but in the same location:   
```{r}
DiffDom_sameLoc=BothDom %>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc==HumanLoc) 

ggplot(DiffDom_sameLoc,aes(x=HumanLoc)) +  geom_histogram(stat="count") + labs(x="PAS Location", y= "Number of Genes", title="Dominat PAS in same genic location but different PAS")

```
Now I can look at those that are in different locations. 

```{r}
DiffDom_diffLoc=BothDom %>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc) 


ggplot(DiffDom_diffLoc,aes(x=ChimpLoc))+ geom_histogram(stat="count")+ labs(x="Chimp Dominant Location", y="Number of Genes", title="Location of dominant PAS when they are differnet between species")
ggplot(DiffDom_diffLoc,aes(x=HumanLoc))+ geom_histogram(stat="count")+labs(x="Human Dominant Location", y="Number of Genes", title="Location of dominant PAS when they are differnet between species")
```

This is interesting but could be based on the annotations. I will look at the genes with human in intron and chimp in 3' UTR.  

```{r}
DiffDom_diffLoc_humanIntronChimpUTR= DiffDom_diffLoc %>% filter(ChimpLoc=="utr3", HumanLoc=="intron")

nrow(DiffDom_diffLoc_humanIntronChimpUTR)
```

I will look to see if these genes are those I see with differential APA.  

**How do i test if this number of genes is enriched??** 
