---
title: "DominatPASandDIffAPA"
author: "Briana Mittleman"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I will look to see if any of the genes I found with differences in the most dominant PAS were called as differentially used. I will look first at the gene level then move to the PAS level.  

```{r}
library(tidyverse)
library(workflowr)
```

###Nuclear  


```{r}
HumanIntronChimpUTR=read.table( "../data/DominantPAS/Nuclear_HumanIntronicChimpUTR.txt", header = T, stringsAsFactors = F)
HumanUTRChimpChimp= read.table( "../data/DominantPAS/Nuclear_HumanUTRChimpIntronic.txt",  header = T, stringsAsFactors = F)
```


```{r}
sig=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sig$p.adjust=as.numeric(as.character(sig$p.adjust))
sig_genes=sig %>% filter(p.adjust<.05) %>% separate(cluster, into=c("chrom","gene"), sep=":")
```

```{r}
OverlapHumanIntronChimpUTR=length(intersect(HumanIntronChimpUTR$gene, sig_genes$gene) )
OverlapHumanIntronChimpUTR/nrow(HumanIntronChimpUTR)
```

```{r}
OverlapHumanUTRChimpChimp= length(intersect(HumanUTRChimpChimp$gene, sig_genes$gene) )
OverlapHumanUTRChimpChimp/nrow(HumanUTRChimpChimp)
```


Compare this to the genes with the same dominant PAS.  

```{r}
NuclearSame= read.table( "../data/DominantPAS/Nuclear_SameDom.txt",  header = T, stringsAsFactors = F)

OverlapSame=length(intersect(NuclearSame$gene, sig_genes$gene) )
OverlapSame/nrow(NuclearSame)
```


###Total  

```{r}
TotalHumanIntronChimpUTR=read.table( "../data/DominantPAS/Total_HumanIntronicChimpUTR.txt", header = T, stringsAsFactors = F)
TotalHumanUTRChimpChimp= read.table( "../data/DominantPAS/Total_HumanUTRChimpIntronic.txt",  header = T, stringsAsFactors = F)
```

```{r}
TotalOverlapHumanIntronChimpUTR=length(intersect(TotalHumanIntronChimpUTR$gene, sig_genes$gene) )
TotalOverlapHumanIntronChimpUTR/nrow(TotalHumanIntronChimpUTR)
```

```{r}
TotalOverlapHumanUTRChimpChimp= length(intersect(TotalHumanUTRChimpChimp$gene, sig_genes$gene) )
TotalOverlapHumanUTRChimpChimp/nrow(TotalHumanUTRChimpChimp)
```

Compare this to the genes with the same dominant PAS.  

```{r}
TotalSame= read.table( "../data/DominantPAS/Total_SameDom.txt",  header = T, stringsAsFactors = F)

TotalOverlapSame=length(intersect(TotalSame$gene, sig_genes$gene) )
TotalOverlapSame/nrow(TotalSame)
```


###Analysis
Looks like the dominant human intronic and chimp 3' utr are more likely to be called differential APA.  


