---
title: "No ties for dominant"
author: "Briana Mittleman"
date: "4/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I had just been taking the first pas as the top one if there is a tie. This shouldnt matter because they should match between species. I will try removing these genes.  

```{r}
library(workflowr)
library(tidyverse)
```

```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% mutate(chrom=paste(chr,start,end,gene, sep=":"))
```



This flips the distribution. It over corrects.  


###Dominant from original usage in filtered set  

```{r}
Chimp_Dom_ties= PAS %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==2) 
nrow(Chimp_Dom_ties)

Chimp_Dom_Noties= PAS %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)
nrow(Chimp_Dom_Noties)

Human_Dom_ties= PAS %>%
  group_by(gene) %>%
  top_n(1,Human) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==2) 
nrow(Human_Dom_ties)

Human_Dom_Noties= PAS %>%
  group_by(gene) %>%
  top_n(1,Human) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)
nrow(Human_Dom_Noties)

```

Inner join no ties  

```{r}
BothDom_noties= Human_Dom_Noties %>% inner_join(Chimp_Dom_Noties, by="gene")
```

Same dominant: 

```{r}
SameDom= BothDom_noties %>%  filter(ChimpPAS==HumanPAS)
nrow(SameDom)

SameDom_g= SameDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(SameDom_g, aes(x=Location, by=Species, fill=Species))+ geom_bar(stat="count",position = "Dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species \nremove ties (n=7,638)") + scale_fill_brewer(palette = "Dark2")
```
```{r}
DiffDom=BothDom_noties %>% filter(ChimpPAS!=HumanPAS)
nrow(DiffDom)
```

```{r}
DiffDom_g= DiffDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


ggplot(DiffDom_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS \n No ties (n=2014)") + scale_fill_brewer(palette = "Dark2")
```
##1 PAS  
What about if we remove any PAS only has 1 PAS  

```{r}
PAS_filt=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% group_by(gene) %>% mutate(nPAS=n()) %>% filter(nPAS!=1)
nrow(PAS_filt)
nrow(PAS)
```
Lose about 1k PAS.  

```{r}
Chimp_Dom= PAS_filt %>%
  group_by(gene) %>%
  arrange(desc(Chimp)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)


Human_Dom= PAS_filt %>%
  group_by(gene) %>%
  arrange(desc(Human)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)


BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")

```

same dominant: 

```{r}
SameDom= BothDom %>%  filter(ChimpPAS==HumanPAS)
nrow(SameDom)

SameDom_g= SameDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(SameDom_g, aes(x=Location, by=Species, fill=Species))+ geom_bar(stat="count",position = "Dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species \nremove 1 PAS genes (n=6,496)") + scale_fill_brewer(palette = "Dark2")
```
```{r}
DiffDom=BothDom %>% filter(ChimpPAS!=HumanPAS)
nrow(DiffDom)

DiffDom_g= DiffDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


ggplot(DiffDom_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS \n remove 1 PAS genes (n=2092)") + scale_fill_brewer(palette = "Dark2")
```

Doesnt help either.  
