---
title: "Are Most used PAS dAPA"
author: "Briana Mittleman"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to see if the most used PAS are also the differentially used.  

```{r}
library(tidyverse)
library(workflowr)
library(cowplot)
```
Caveat- ties.  

Dominant PAS
```{r}
allPAS= read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T) 
ChimpPASwMean =allPAS %>% dplyr::select(-Human)
HumanPASwMean =allPAS %>% dplyr::select(-Chimp)


Chimp_Dom= ChimpPASwMean %>%
  group_by(gene) %>%
  arrange(desc(Chimp)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)


Chimp_Dom2= ChimpPASwMean %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) 
nrow(Chimp_Dom2%>% filter(nPer>1) )


  

Human_Dom= HumanPASwMean %>%
  group_by(gene) %>%
  arrange(desc(Human)) %>% 
  slice(1) %>% 
  group_by(gene) %>% 
  mutate(npas=n()) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)

Human_Dom2= HumanPASwMean %>%
  group_by(gene) %>%
  top_n(1,Human) %>% 
  mutate(nPer=n()) 

nrow(Human_Dom2 %>% filter(nPer>1) )

```

198 genes with a tie in chimp and 161 with a tie in human. Picks top for both for analysis. 


##No PAS usage cutoff for dominant  

```{r}
BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")

SameDom= BothDom %>% filter(ChimpPAS==HumanPAS,HumanLoc!="008559")
nrow(SameDom)


SameDom_g= SameDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)
ggplot(SameDom_g, aes(x=Location, by=Species, fill=Species))+ geom_bar(stat="count",position = "Dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species")


ggplot(SameDom, aes(x=HumanLoc))+ geom_bar(stat="count") + labs(x="Location", y="Number of Genes", title="Dominant PAS for genes with matching by species")
```
```{r}
DiffDom=BothDom %>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(DiffDom)

DiffDom_g= DiffDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plotnones=ggplot(DiffDom_g,aes(by=Species, x=Location, fill=Species))+ geom_histogram(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant PAS") + scale_fill_brewer(palette = "Dark2")
```

Diff dominant different locations:  

```{r}
DiffDomDiffLoc=BothDom %>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(DiffDomDiffLoc)

DiffDomDiffLoc_g= DiffDomDiffLoc %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plotnone=ggplot(DiffDomDiffLoc_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations") + scale_fill_brewer(palette = "Dark2")
```

##Implement filters  

I want to filter these and see if these distribution changes. First I will say the most used has to be over 50% in both species.   

```{r}
BothDom_50= BothDom %>% filter(Chimp>=.5, Human>.5)
nrow(BothDom_50)
BothDom_40= BothDom %>% filter(Chimp>=.4, Human>.4)
nrow(BothDom_40)
BothDom_30= BothDom %>% filter(Chimp>=.3, Human>.3)
nrow(BothDom_30)
BothDom_20= BothDom %>% filter(Chimp>=.2, Human>.2)
nrow(BothDom_20)
BothDom_10= BothDom %>% filter(Chimp>=.1, Human>.1)
nrow(BothDom_10)

cutoffs=c(".5",'.4','.3','.2','.1', "0")
nPAS=c(nrow(BothDom_50),nrow(BothDom_40),nrow(BothDom_30),nrow(BothDom_20),nrow(BothDom_10), nrow(BothDom))
nPAS



```
###Different Dominant, different Location 
```{r}
BothDom_50_diff=BothDom_50%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(BothDom_50_diff)

BothDom_50_diff_g= BothDom_50_diff %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot50=ggplot(BothDom_50_diff_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations (50%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_40_diff=BothDom_40%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(BothDom_40_diff)

BothDom_40_diff_g= BothDom_40_diff %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot40=ggplot(BothDom_40_diff_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations (40%)") + scale_fill_brewer(palette = "Dark2")
```




```{r}
BothDom_30_diff=BothDom_30%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(BothDom_30_diff)

BothDom_30_diff_g= BothDom_30_diff %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot30=ggplot(BothDom_30_diff_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations (30%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_20_diff=BothDom_20%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(BothDom_20_diff)

BothDom_20_diff_g= BothDom_20_diff %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot20=ggplot(BothDom_20_diff_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations (20%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_10_diff=BothDom_10%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559", ChimpLoc!=HumanLoc)
nrow(BothDom_10_diff)

BothDom_10_diff_g= BothDom_10_diff %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot10=ggplot(BothDom_10_diff_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Dominant PAS in different locations (10%)") + scale_fill_brewer(palette = "Dark2")
```


```{r}
plot_grid(plot50, plot40, plot30, plot20, plot10,plotnone)
```


###Filters in same loc ok  

```{r}
BothDom_50_same=BothDom_50%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(BothDom_50_same)

BothDom_50_same_g= BothDom_50_same %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot50s=ggplot(BothDom_50_same_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant (50%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_40_same=BothDom_40%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(BothDom_40_same)
BothDom_40_same_g= BothDom_40_same %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot40s=ggplot(BothDom_40_same_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant (40%)") + scale_fill_brewer(palette = "Dark2")
```




```{r}
BothDom_30_same=BothDom_30%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(BothDom_30_same)
BothDom_30_same_g= BothDom_30_same %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot30s=ggplot(BothDom_30_same_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant (30%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_20_same=BothDom_20%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(BothDom_20_same)
BothDom_20_same_g= BothDom_20_same %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot20s=ggplot(BothDom_20_same_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant (20%)") + scale_fill_brewer(palette = "Dark2")
```

```{r}
BothDom_10_same=BothDom_10%>% filter(ChimpPAS!=HumanPAS,HumanLoc!="008559")
nrow(BothDom_10_same)
BothDom_10_same_g= BothDom_10_same %>% select(gene, ChimpLoc, HumanLoc) %>% gather("Species", "Location", -gene)


plot10s=ggplot(BothDom_10_same_g,aes(by=Species, x=Location, fill=Species))+ geom_bar(stat="count",position = "dodge") + labs(x="Location", y="Number of Genes", title="Different Dominant (10%)") + scale_fill_brewer(palette = "Dark2")
```


```{r}
plot_grid(plot50s, plot40s, plot30s, plot20s, plot10s,plotnones)
```

Plot numbers:  

```{r}
cutoffs=c("0.5",'0.4','0.3','0.2','0.1', "0")
#nPAS=c(nrow(BothDom_50),nrow(BothDom_40),nrow(BothDom_30),nrow(BothDom_20),nrow(BothDom_10), nrow(BothDom))
nPASDiff=c(nrow(BothDom_50_diff),nrow(BothDom_40_diff),nrow(BothDom_30_diff),nrow(BothDom_20_diff),nrow(BothDom_10_diff), nrow(DiffDomDiffLoc))
nPASSameLoc=c(nrow(BothDom_50_same),nrow(BothDom_40_same),nrow(BothDom_30_same),nrow(BothDom_20_same),nrow(BothDom_10_same), nrow(DiffDom))

NumberDF= as.data.frame(cbind(cutoffs, Genes=nPAS, DifferentPASandLoc=nPASDiff, DifferentPAS=nPASSameLoc)) %>% gather("Set", "count",-cutoffs )
NumberDF$count=as.numeric(NumberDF$count)

ggplot(NumberDF,aes(x=cutoffs, y=count, by=Set, fill=Set)) + geom_bar(stat="identity",position = "dodge" ) +geom_text(stat='identity', aes(label=count), vjust=0,position = position_dodge(width = 1)) + labs(x="Cutoff for PAS to be dominant (Both Species)", y="Number of genes")+ scale_fill_brewer(palette = "Dark2", labels=c("Different PAS Same Location", "Different PAS and Location", "All Genes"),name="") + theme(legend.position = "top")

```


##Compare to differencially used  

```{r}


PASMeta=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% select(PAS, gene, chr, start,end,loc)

DiffIso=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt",header = T, stringsAsFactors = F) %>% inner_join(PASMeta, by=c("chr",'start','end',"gene"))

DiffIso_sig= DiffIso %>% filter(SigPAU2=="Yes") 


DiffIsowDom=DiffIso %>% mutate(ChimpDom=ifelse(PAS %in% Chimp_Dom2$PAS, "Yes", "No"),HumanDom=ifelse(PAS %in% Human_Dom2$PAS, "Yes", "No") ) 


ggplot(DiffIsowDom,aes(x=loc,fill=ChimpDom)) + geom_bar(stat = "count") + facet_grid(~SigPAU2) +labs(x="", y="Number of PAS", title="Differentially used PAS colored by Dominant Status in Chimp")+ scale_fill_brewer(palette = "Dark2",name="Is PAS Dominant in Chimp") + theme(legend.position = "top")


ggplot(DiffIsowDom,aes(x=loc,fill=HumanDom)) + geom_bar(stat = "count") + facet_grid(~SigPAU2) +labs(x="", y="Number of PAS", title="Differentially used PAS colored by Dominant Status in Human")+ scale_fill_brewer(palette = "Dark2",name="Is PAS Dominant in Chimp") + theme(legend.position = "top")

```

Of the 3076 how many are in the dominant set.  

I need to use the set that includes ties here:  
```{r}
SiginChimp_Dom2= Chimp_Dom2 %>% ungroup %>% select(PAS) %>% inner_join(DiffIso_sig, by="PAS")
nrow(SiginChimp_Dom2)


SiginHuman_Dom2= Human_Dom2 %>% ungroup %>% select(PAS) %>% inner_join(DiffIso_sig, by="PAS")
nrow(SiginHuman_Dom2)



```

Enrichemnt:   
  

```{r}
EnrichChimp=c()
PvalueChimp=c()

for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(DiffIsowDom %>% filter(ChimpDom=="Yes", SigPAU2=="Yes", loc==i))
  m=nrow(DiffIsowDom %>% filter(SigPAU2=="Yes", loc==i))
  n=nrow(DiffIsowDom %>% filter(SigPAU2=="No", loc==i))
  k=nrow(DiffIsowDom %>% filter(loc==i,ChimpDom=="Yes"))
  N=nrow(DiffIsowDom %>% filter(loc==i))
  PvalueChimp=c(PvalueChimp, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichChimp=c(EnrichChimp, round(enrich,2))
}


EnrichHuman=c()
PvalueHuman=c()

for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(DiffIsowDom %>% filter(HumanDom=="Yes", SigPAU2=="Yes", loc==i))
  m=nrow(DiffIsowDom %>% filter(SigPAU2=="Yes", loc==i))
  n=nrow(DiffIsowDom %>% filter(SigPAU2=="No", loc==i))
  k=nrow(DiffIsowDom %>% filter(loc==i,HumanDom=="Yes"))
  N=nrow(DiffIsowDom %>% filter(loc==i))
  PvalueHuman=c(PvalueHuman, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichHuman=c(EnrichHuman, round(enrich,2))
}

locations=c('cds', 'end', 'intron', 'utr3', 'utr5')


EnrichDF=as.data.frame(cbind(locations, EnrichChimp,EnrichHuman ))
colnames(EnrichDF)=c("GenicLoc", "Chimp", "Human")

EnrichDF
```

Compare this result to a set of PAS by usage:  

For this I will use dominant at 30%, I will then use, used at 30%

```{r}
SiginChimp_Dom2_30= Chimp_Dom2 %>% ungroup%>% filter(Chimp>=.3)  %>% select(PAS) %>% inner_join(DiffIso_sig, by="PAS")
nrow(SiginChimp_Dom2)


SiginHuman_Dom2_30= Human_Dom2 %>% ungroup %>% filter(Human>=.3) %>% select(PAS) %>% inner_join(DiffIso_sig, by="PAS")
nrow(SiginHuman_Dom2)

HumanUsed30= allPAS %>% filter(Human>=.3)
nrow(HumanUsed30)
ChimpUsed30= allPAS %>% filter(Chimp>=.3)
nrow(ChimpUsed30)


allPAS_anno= allPAS %>% select(PAS,loc) %>% mutate(SigDPAU=ifelse(PAS %in% DiffIso_sig$PAS, "Yes","No"), HumanUse30=ifelse(PAS %in% HumanUsed30$PAS, "Yes", "No"), ChimpUse30=ifelse(PAS %in% ChimpUsed30$PAS, "Yes","No"), DomHuman=ifelse(PAS %in% SiginHuman_Dom2_30$PAS, "Yes", "No"), DomChimp=ifelse(PAS %in%SiginChimp_Dom2_30$PAS, "Yes", "No" ))



```


Enrichment of dominant in diff used human: 


```{r}
x=nrow(allPAS_anno %>% filter(DomHuman=="Yes", SigDPAU=="Yes"))
m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes"))
n=nrow(allPAS_anno %>% filter(SigDPAU=="No"))
k=nrow(allPAS_anno %>% filter(DomHuman=="Yes"))
N=nrow(allPAS_anno)
phyper(x,m,n,k,lower.tail=F)
enrich=(x/k)/(m/N)
enrich
```

Compare to human  just used at 30%
```{r}
x=nrow(allPAS_anno %>% filter(HumanUse30=="Yes", SigDPAU=="Yes"))
m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes"))
n=nrow(allPAS_anno %>% filter(SigDPAU=="No"))
k=nrow(allPAS_anno %>% filter(HumanUse30=="Yes"))
N=nrow(allPAS_anno)
phyper(x,m,n,k,lower.tail=F)
enrich=(x/k)/(m/N)
enrich
```

Enrichment of dominant in diff used chimp:



```{r}
x=nrow(allPAS_anno %>% filter(DomChimp=="Yes", SigDPAU=="Yes"))
m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes"))
n=nrow(allPAS_anno %>% filter(SigDPAU=="No"))
k=nrow(allPAS_anno %>% filter(DomChimp=="Yes"))
N=nrow(allPAS_anno)
phyper(x,m,n,k,lower.tail=F)
enrich=(x/k)/(m/N)
enrich
```

Enrichment of used at 30%  

```{r}
x=nrow(allPAS_anno %>% filter(ChimpUse30=="Yes", SigDPAU=="Yes"))
m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes"))
n=nrow(allPAS_anno %>% filter(SigDPAU=="No"))
k=nrow(allPAS_anno %>% filter(ChimpUse30=="Yes"))
N=nrow(allPAS_anno)
phyper(x,m,n,k,lower.tail=F)
enrich=(x/k)/(m/N)
enrich
```

At the 30% cutoff in the domiant PAS has 13.8X enrichement and 30% used has 2.49X enrichment. Is this based on the dominant at 30% always being in the other set?  

Do by location:  


```{r}
EnrichChimpDom=c()
PvalueChimpDom=c()

for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(allPAS_anno %>% filter(DomChimp=="Yes", SigDPAU=="Yes", loc==i))
  m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes", loc==i))
  n=nrow(allPAS_anno %>% filter(SigDPAU=="No", loc==i))
  k=nrow(allPAS_anno %>% filter(loc==i,DomChimp=="Yes"))
  N=nrow(allPAS_anno %>% filter(loc==i))
  PvalueChimpDom=c(PvalueChimpDom, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichChimpDom=c(EnrichChimpDom, round(enrich,2))
}

EnrichChimp30=c()
PvalueChimp30=c()
for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(allPAS_anno %>% filter(ChimpUse30=="Yes", SigDPAU=="Yes", loc==i))
  m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes", loc==i))
  n=nrow(allPAS_anno %>% filter(SigDPAU=="No", loc==i))
  k=nrow(allPAS_anno %>% filter(loc==i,ChimpUse30=="Yes"))
  N=nrow(allPAS_anno %>% filter(loc==i))
  PvalueChimp30=c(PvalueChimp30, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichChimp30=c(EnrichChimp30, round(enrich,2))
}
```

```{r}
EnrichHumanDom=c()
PvalueHumanDom=c()

for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(allPAS_anno %>% filter(DomHuman=="Yes", SigDPAU=="Yes", loc==i))
  m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes", loc==i))
  n=nrow(allPAS_anno %>% filter(SigDPAU=="No", loc==i))
  k=nrow(allPAS_anno %>% filter(loc==i,DomHuman=="Yes"))
  N=nrow(allPAS_anno %>% filter(loc==i))
  PvalueHumanDom=c(PvalueHumanDom, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichHumanDom=c(EnrichHumanDom, round(enrich,2))
}

EnrichHuman30=c()
PvalueHuman30=c()
for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(allPAS_anno %>% filter(HumanUse30=="Yes", SigDPAU=="Yes", loc==i))
  m=nrow(allPAS_anno %>% filter(SigDPAU=="Yes", loc==i))
  n=nrow(allPAS_anno %>% filter(SigDPAU=="No", loc==i))
  k=nrow(allPAS_anno %>% filter(loc==i,HumanUse30=="Yes"))
  N=nrow(allPAS_anno %>% filter(loc==i))
  PvalueHuman30=c(PvalueHuman30, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichHuman30=c(EnrichHuman30, round(enrich,2))
}
```

```{r}
EnrichDataDFHuman=as.data.frame(cbind(loc=c('cds', 'end', 'intron', 'utr3', 'utr5'), Used30= EnrichHuman30,Dominant=EnrichHumanDom)) %>% gather("Set", "Enrichment", -loc)%>% mutate(species="Human")

EnrichDataDFChimp=as.data.frame(cbind(loc=c('cds', 'end', 'intron', 'utr3', 'utr5'), Used30= EnrichChimp30,Dominant=EnrichChimpDom)) %>% gather("Set", "Enrichment", -loc) %>% mutate(species="Chimp")

EnrichDataDFBoth= EnrichDataDFHuman %>% bind_rows(EnrichDataDFChimp)


EnrichDataDFBoth$Enrichment=as.numeric(EnrichDataDFBoth$Enrichment)
EnrichDataDFBoth$Set=as.factor(EnrichDataDFBoth$Set)




ggplot(EnrichDataDFBoth, aes(x=loc, y=Enrichment, by=Set, fill=Set)) + geom_bar(stat="identity", position="dodge") + geom_hline(yintercept = 1) + facet_grid(~species)+ scale_fill_brewer(palette = "Dark2",name="") + theme(legend.position = "bottom") + labs(x="Genic Location", title="Dominant PAS are more enriched for dPAS \nthan all PAS used at same level") 
```

test opposite

```{r}

EnrichChimpDomOpp=c()
for (i in c('cds', 'end', 'intron', 'utr3', 'utr5')){
  x=nrow(allPAS_anno %>% filter(DomChimp=="Yes", SigDPAU=="Yes", loc==i))
  m=nrow(allPAS_anno %>% filter(DomChimp=="Yes", loc==i))
  n=nrow(allPAS_anno %>% filter(DomChimp=="No", loc==i))
  k=nrow(allPAS_anno %>% filter(loc==i,SigDPAU=="Yes"))
  N=nrow(allPAS_anno %>% filter(loc==i))
  enrich=(x/k)/(m/N)
  EnrichChimpDomOpp=c(EnrichChimpDomOpp, round(enrich,2))
}

EnrichChimpDomOpp
EnrichChimpDom
```
ok test. its the same.  

Make sure this isnt driven by counts:  

I can do this with raw counts.  

```{r}
#allPAS_anno= allPAS %>% select(PAS,loc) %>% mutate(SigDPAU=ifelse(PAS %in% DiffIso_sig$PAS, "Yes","No"), HumanUse30=ifelse(PAS %in% HumanUsed30$PAS, "Yes", "No"), ChimpUse30=ifelse(PAS %in% ChimpUsed30$PAS, "Yes","No"), DomHuman=ifelse(PAS %in% SiginHuman_Dom2_30$PAS, "Yes", "No"), DomChimp=ifelse(PAS %in%SiginChimp_Dom2_30$PAS, "Yes", "No" ))
PASMetaloc= PASMeta %>% select(PAS,loc)


HumanCounts=read.table("../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc", header = T)%>% separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")  %>% select(PAS, contains("_N"))  %>% gather("ind", "count", -PAS) %>% group_by(PAS) %>% summarise(meanCount=mean(count))

HumanCounts_use30= HumanCounts %>% filter(PAS %in% HumanUsed30$PAS) %>% mutate(Set="Use30", species="Human") %>% inner_join(PASMetaloc, by="PAS")
HumanCounts_Dom= HumanCounts %>% filter(PAS %in% SiginHuman_Dom2_30$PAS)  %>% mutate(Set="Dominant", species="Human") %>% inner_join(PASMetaloc, by="PAS")

ChimpCounts=read.table("../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc", header = T)%>% separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")  %>% select(PAS, contains("_N"))  %>% gather("ind", "count", -PAS) %>% group_by(PAS) %>% summarise(meanCount=mean(count))


ChimpCounts_use30= ChimpCounts %>% filter(PAS %in%ChimpUsed30$PAS) %>% mutate(Set="Use30", species="Chimp") %>% inner_join(PASMetaloc, by="PAS")
ChimpCounts_Dom= ChimpCounts %>% filter(PAS %in% SiginChimp_Dom2_30$PAS)  %>% mutate(Set="Dominant", species="Chimp") %>% inner_join(PASMetaloc, by="PAS")


AllCountsDF= HumanCounts_use30 %>% bind_rows(HumanCounts_Dom) %>% bind_rows(ChimpCounts_use30) %>% bind_rows(ChimpCounts_Dom)

```



```{r}
ggplot(AllCountsDF, aes(y=log10(meanCount), x=Set, fill=Set)) +geom_boxplot() +facet_grid(~species) +scale_fill_brewer(palette = "Dark2",name="") + theme(legend.position = "none") + labs(title="Counts for Dominant PAS at 30% vs Usage at 30%", y="log10(Mean Counts)") 

ggplot(AllCountsDF, aes(y=log10(meanCount), by=Set, fill=Set, x=loc)) +geom_boxplot() +facet_grid(~species) +scale_fill_brewer(palette = "Dark2",name="") + theme(legend.position = "none") + labs(title="Counts for Dominant PAS at 30% vs Usage at 30%", y="log10(Mean Counts)") 
```


##diff dominance and discovery  
DiffDom
```{r}

SamDomLoc=SameDom %>% ungroup() %>% select(ChimpPAS) %>% rename("PAS"=ChimpPAS) %>% inner_join(allPAS, by="PAS")
ggplot(SamDomLoc, aes(x=disc, fill=disc)) +geom_bar(stat="count") + labs(title="Same dominant")

DiffDomChimp= DiffDom %>% ungroup() %>% select(ChimpPAS) %>% rename("PAS"=ChimpPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomChimp, aes(x=disc, fill=disc)) +geom_bar(stat="count") + labs(title="Different dominant, Chimp PAS")



DiffDomHuman= DiffDom %>% ungroup() %>% select(HumanPAS) %>% rename("PAS"=HumanPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomHuman, aes(x=disc, fill=disc)) +geom_bar(stat="count") +  labs(title="Different dominant, Human PAS")
```

```{r}
#DiffDomDiffLoc

DiffDomDiffLocChimp= DiffDomDiffLoc %>% ungroup() %>% select(ChimpPAS) %>% rename("PAS"=ChimpPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomDiffLocChimp, aes(x=disc, fill=disc)) +geom_bar(stat="count") +  labs(title="Different dominant, different Loc, Chimp PAS")



DiffDomDiffLocHuman= DiffDomDiffLoc %>% ungroup() %>% select(HumanPAS) %>% rename("PAS"=HumanPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomDiffLocHuman, aes(x=disc, fill=disc)) +geom_bar(stat="count") +  labs(title="Different dominant, different Loc, Human PAS")
```

Does this go away if we say has to be 50%  
```{r}

DiffDomChimp50= BothDom_50_same %>% ungroup() %>% select(ChimpPAS) %>% rename("PAS"=ChimpPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomChimp50, aes(x=disc, fill=disc)) +geom_bar(stat="count") + labs(title="Different dominant, Chimp PAS")



DiffDomHuman50= BothDom_50_same %>% ungroup() %>% select(HumanPAS) %>% rename("PAS"=HumanPAS) %>% inner_join(allPAS, by="PAS")

ggplot(DiffDomHuman50, aes(x=disc, fill=disc)) +geom_bar(stat="count") +  labs(title="Different dominant, Human PAS")
```

Problems:

- snorna (need to be removed)
- mispriming? 
- why not disovered in the right species 

