---
title: "Explore Relationship between APA and expression"
author: "Briana Mittleman"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



In this analysis I will do as much as I can do compare APA and expression. I will look at the genes we have data for both. 
```{r}
library(tidyverse)
library("gplots")
library("scales")
library("RColorBrewer")
library(ggpubr)
```

I will start with a full correlation matrix using the counts for the top used PAS (mean human and chimp).    


```{r}
TopUsedPAS=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header=T,stringsAsFactors = F) %>% mutate(MeanUsage=(Chimp+Human)/2) %>% group_by(gene) %>% arrange(desc(MeanUsage)) %>% slice(1) %>% ungroup() %>% dplyr::select(PAS, chr, start, end) %>% rename("Chr"= chr,"Start"= start, "End"= end)
```

Filter these in the counts file:  

```{r}

#Both:human74:chr1:944201:944401:+:NOC2L_utr3


HumanCounts=read.table("../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc",stringsAsFactors = F,header = T) %>% inner_join(TopUsedPAS,by=c("Chr", "Start", "End")) %>% separate(Geneid, into=c("d", "pas", "chr","start","end","strand", "geneID"), sep=":") %>% separate(geneID,into=c("gene","loc"),sep="_") %>% dplyr::select(PAS, gene, contains("_N"))

ChimpCounts=read.table("../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc",stringsAsFactors = F,header = T) %>% separate(Geneid, into=c("d", "PAS", "chr","start","end","strand", "geneID"), sep=":") %>% separate(geneID,into=c("gene","loc"),sep="_") %>% inner_join(TopUsedPAS,by=c("PAS")) %>% dplyr::select(PAS, gene, contains("_N"))


AllCounts=HumanCounts %>% inner_join(ChimpCounts, by=c("gene", "PAS")) %>% dplyr::select(-PAS)


colnames(AllCounts)=paste("APA" ,colnames(AllCounts), sep="_")


AllCounts=AllCounts %>% rename("gene"= APA_gene)
```


Expression counts:  

```{r}

nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F)
HumanRNA=read.table("../Human/data/RNAseq/ExonCounts/RNAseqOrthoExon.fixed.fc", header = T, stringsAsFactors = F) %>% dplyr::select(-Chr,-Start,-End,-Strand, -Length)

ChimpRNA=read.table("../Chimp/data/RNAseq/ExonCounts/RNAseqOrthoExon.fixed.fc", header = T, stringsAsFactors = F) %>% dplyr::select(-Chr,-Start,-End,-Strand, -Length)


RNACounts=HumanRNA %>% inner_join(ChimpRNA,by="Geneid") %>% rename("Gene_stable_ID"=Geneid) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(-Gene_stable_ID, -Source_of_gene_name) %>% rename("gene"= Gene.name) %>% group_by(gene) %>% slice(1) %>% ungroup()


colnames(RNACounts)=paste("RNA" ,colnames(RNACounts), sep="_")

RNACounts=RNACounts %>% rename("gene"= RNA_gene)
```


Add all together:  

```{r}
BothPhenoCount=  RNACounts %>% inner_join(AllCounts, by="gene")%>% column_to_rownames(var="gene")

Names=colnames(BothPhenoCount)
pal <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), brewer.pal(12, "Set3"))

Species=c(rep("Human",6), rep("Chimp",6),rep("Human",6),rep("Chimp",6))
Meta=as.data.frame(cbind(Names,Species )) %>% mutate(Pheno=ifelse(grepl("RNA",Names),"RNA", "APA" )) 

cors <- cor(BothPhenoCount, method="spearman", use="pairwise.complete.obs")
colors <- colorRampPalette(c(brewer.pal(9, "Blues")[1],brewer.pal(9, "Blues")[9]))(100)
heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none',ColSideColors=pal[as.integer(as.factor(Meta$Species))], RowSideColors=pal[as.integer(as.factor(Meta$Pheno))+9], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```


Pvalues:  

```{r}
DE= read.table("../data/DiffExpression/DEtested_allres.txt",header=F, stringsAsFactors = F,col.names = c('Gene_stable_ID', 'logFC' ,'AveExpr', 't', 'P.Value', 'adj.P.Val', 'B')) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(-Gene_stable_ID, -Source_of_gene_name) %>% rename("gene"=Gene.name)

APA=read.table(c("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt"),header = T,stringsAsFactors = F) 
```


```{r}
GenesWithapa=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt", header = T, stringsAsFactors = F)

DEandAPA= DE %>% inner_join(APA, by="gene") %>% mutate(HasDPAU=ifelse(gene %in% GenesWithapa$gene, "Yes","No"))
```

Plot:

```{r}
ggplot(DEandAPA,aes(x=HasDPAU, y=abs(logFC)))+ geom_boxplot() + stat_compare_means() + labs(x="Gene has significant dPAS", y="Absolute value of DE log effect size", title="DE effect size by dAPA gene")
```

```{r}
ggplot(DEandAPA,aes(x=HasDPAU, y=adj.P.Val))+ geom_boxplot() + stat_compare_means() + labs(x="Gene has significant dPAS", y="Adjusted DE Pvalue", title="DE pvalue dAPA gene")
```


Correlation between effect sizes:  

```{r}
ggplot(DEandAPA,aes(x=logFC, y=logef,col=HasDPAU)) + geom_point() + labs(x="DE log Effect size", y="APA effect size",title="Relationship between effect sizes")  + geom_density2d(col="black")

ggplot(DEandAPA,aes(x=abs(logFC), y=abs(logef))) + geom_point(aes(col=HasDPAU) )+ labs(x="abs(DE log Effect size)", y="abs(APA effect size)",title="Relationship between absolute values of effect sizes") + geom_smooth(method="lm") +annotate("text", label="Beta= 0.158339 , R2= 0.01",y=40,x=5)

summary(lm(abs(DEandAPA$logef)~abs(DEandAPA$logFC)))
```

Subset the PAS in the top used:  


```{r}
TopUsedPASlow= TopUsedPAS %>% rename("chr"=Chr, "start"=Start, "end"=End)
DEandAPA_filt= DEandAPA %>% semi_join(TopUsedPASlow,by=c("chr","start","end"))



```
Plot:

```{r}
ggplot(DEandAPA_filt,aes(x=HasDPAU, y=abs(logFC)))+ geom_boxplot() + stat_compare_means() + labs(x="Gene has significant dPAS", y="Absolute value of DE log effect size", title="DE effect size by dAPA gene\n most used PAS")
```

```{r}
ggplot(DEandAPA_filt,aes(x=HasDPAU, y=adj.P.Val))+ geom_boxplot() + stat_compare_means() + labs(x="Gene has significant dPAS", y="Adjusted DE Pvalue", title="DE pvalue dAPA gene \n most used PAS")
```

Correlation between effect sizes:  

```{r}
ggplot(DEandAPA_filt,aes(x=logFC, y=logef,col=HasDPAU)) + geom_point() + labs(x="DE log Effect size", y="APA effect size", title="Relationship between effect sizes \n most used PAS") 

ggplot(DEandAPA_filt,aes(x=abs(logFC), y=abs(logef))) + geom_point(aes(col=HasDPAU) )+ labs(x="abs(DE log Effect size)", y="abs(APA effect size)",title="Relationship between absolute values of effect sizes \n most used PAS") + geom_smooth(method="lm") +annotate("text", label="Beta= 0.14957 , R2= 0.01",y=30,x=5)

summary(lm(abs(DEandAPA_filt$logef)~abs(DEandAPA_filt$logFC)))
```


Number of significant PAS and de  

```{r}
DEandAPA_sig= DEandAPA %>% filter(SigPAU2 == "Yes") %>% group_by(gene) %>% summarise(NsigPAS=n())

DEandAPA_sigWithDE= DE %>% inner_join(DEandAPA_sig,by="gene" )


summary(lm(data=DEandAPA_sigWithDE, abs(logFC)~NsigPAS))
ggplot(DEandAPA_sigWithDE, aes(x=NsigPAS, y=abs(logFC))) +geom_point() + geom_smooth(method = "lm") + annotate("text",label="beta=.15, r2=0.014, pvalue-2.93e-7",x=4, y=6) + labs(x="Number of differentially used PAS 20%",title="Relationship between DE effect size and dAPA PAS")

cor( DEandAPA_sigWithDE$NsigPAS, DEandAPA_sigWithDE$logFC, method = "spearman")
cor(DEandAPA_sigWithDE$B, DEandAPA_sigWithDE$NsigPAS, method = "spearman")

```


Maybe use the variance in effect sizes for dAPA:  

```{r}
APA_var= APA %>% group_by(gene) %>% summarise(VarLogef=var(logef), VarDeltaPAU=var(deltaPAU))

APA_varDE= APA_var %>% inner_join(DE, by="gene")
```


```{r}

cor(APA_varDE$VarDeltaPAU, APA_varDE$B, method = "spearman")

cor(APA_varDE$VarDeltaPAU, APA_varDE$logFC, method = "spearman")

cor(APA_varDE$VarLogef, APA_varDE$B, method = "spearman")

cor(APA_varDE$VarLogef, APA_varDE$logFC, method = "spearman")

```

Im stuck here. I will get suggestiions/keep thinking 
