---
title: "Information Content"
author: "Briana Mittleman"
date: "4/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggpubr)
library(workflowr)
library(cowplot)
```

Packages/functions for this:

vegan: diversity, can calculate shannon or simpson

I will probably do this in python because I can go gene by gene easier: 

scipy stats example

This is good because I will be able to change the base and see how it effects the measurements 

https://kite.com/python/docs/scipy.stats.entropy

default base is e

```{python, eval=F}
from scipy.stats import entropy
import numpy as np
from math import log, e
entropy([1/2, 1/2], base=2)
  
#shannon 
Shannon2 = -np.sum(pA*np.log2(pA))
```


I most likely want to use a uniform prior. for this. I could get more complicated in the future by weighting differences by utr and intron. this would help find "more surpising" results. 

simpson- squares the probability 


```{bas,eval=F}
from ecopy import diversity 

diversity(x, medod="simpson")


#x- side x species matrix, sites are rows, columns are species - ie column counts, row == pas
```


```{r}
library(vegan)

data(BCI)
dim(BCI)
H <- diversity(BCI)
length(H)

diversity(c(.5,.5,.5))
diversity(c(.25,.75,.25))
#more peak= lower 


diversity(c(.5,.5,.5), "simpson")
diversity(c(.25,.75,.25),"simpson")
#more peak= lower


diversity(c(.5,.5,.5), "inv")
diversity(c(.25,.75,.25),"inv")


```


Seem like it is most simple to use the mean usages for this.  
##Shannon  

First test:

use entropy in python with different bases. 
-base 2 is the classic shannon  and it uses the - when probabilities are given (ie uniform prior)

the python code will work with my meta file for now and take species as an input. 


$H=-\sum^{s}_{i=1}p_{i}log_{2}p_{i}$

$H=-\sum^{s}_{i=1}p_{i}lnp_{i}$

```{bash,eval=F}
mkdir ../data/InfoContent

python InfoContentShannon.py Human
python InfoContentShannon.py Chimp



```

Results: 

```{r}
HumanResInfo= read.table("../data/InfoContent/Human_InfoContent.txt", header = T,stringsAsFactors = F) %>% rename(Human_Base2=base2, Human_basee= basee)
ChimpResInfo= read.table("../data/InfoContent/Chimp_InfoContent.txt", header = T,stringsAsFactors = F) %>% rename(Chimp_Base2=base2, Chimp_basee= basee)

BothResInfo= HumanResInfo %>% inner_join(ChimpResInfo, by=c("gene", "numPAS")) %>% filter(numPAS > 1)
```

First plot the distributions:

```{r}
BothResInfo_2= BothResInfo %>% select(gene, contains("Base2")) %>% gather("species", "base2", -gene)

ggplot(BothResInfo_2, aes(x=base2, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Set1")+ labs(title="Shannon Information Content")


wilcox.test(BothResInfo$Human_Base2, BothResInfo$Chimp_Base2, alternative = "greater")
```

Human shift higher, ie less density:  


```{r}
BothResInfo_e= BothResInfo %>% select(gene, contains("basee")) %>% gather("species", "basee", -gene)

ggplot(BothResInfo_e, aes(x=basee, fill=species)) + geom_density(alpha=.3)
```

I want to look at this by dominance: 

```{r}
ggplot(BothResInfo_2,aes(x=base2, fill=species)) + geom_histogram() + facet_grid(~species)
```



Plot human vs chimp: 

```{r}
ggplot(BothResInfo,aes(x=Human_Base2,y= Chimp_Base2 )) + geom_point() + geom_abline(slope=1, intercept = 0) + stat_cor(col="blue") + geom_density_2d(col="blue")
```
```{r}
ggplot(BothResInfo,aes(x=Human_Base2,y= Chimp_Base2 ,col=numPAS)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0) +labs(title="Shannon Index Colored by number of PAS")
```

Does number explain: 

```{r}
summary(lm(BothResInfo$Human_Base2 ~BothResInfo$numPAS))


summary(lm(BothResInfo$Chimp_Base2 ~BothResInfo$numPAS ))
```
So this is working but the number of PAS explains most of the variation. Maybe I can normalize this out and look at residuals:  


```{r}
BothResInfoRes= BothResInfo %>% mutate(HumanNorm=residuals(BothResInfo$Human_Base2~BothResInfo$numPAS),ChimpNorm=residuals(BothResInfo$Chimp_Base2~BothResInfo$numPAS))
```




pull in dominance:  

```{r}
HumanRes=read.table("../data/DomDefGreaterX/Human_AllGenes_DiffTop.txt", col.names = c("Human_PAS", "gene","Human_DiffDom"),stringsAsFactors = F)

ChimpRes=read.table("../data/DomDefGreaterX/Chimp_AllGenes_DiffTop.txt", col.names = c("Chimp_PAS", "gene","Chimp_DiffDom"),stringsAsFactors = F)

BothRes=HumanRes %>% inner_join(ChimpRes,by="gene")

BothRes_10=BothRes %>% filter(Chimp_DiffDom >=0.1 | Human_DiffDom>=0.1) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=10) 
BothRes_20=BothRes %>% filter(Chimp_DiffDom >=0.2 | Human_DiffDom>=0.2) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=20)
BothRes_30=BothRes %>% filter(Chimp_DiffDom >=0.3 | Human_DiffDom>=0.3) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=30)
BothRes_40=BothRes %>% filter(Chimp_DiffDom >=0.4 | Human_DiffDom>=0.4) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=40)
BothRes_50=BothRes %>% filter(Chimp_DiffDom >=0.5 | Human_DiffDom>=0.5) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=50)
BothRes_60=BothRes %>% filter(Chimp_DiffDom >=0.6 | Human_DiffDom>=0.6) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=60)
BothRes_70=BothRes %>% filter(Chimp_DiffDom >=0.7 | Human_DiffDom>=0.7) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=70)
BothRes_80=BothRes %>% filter(Chimp_DiffDom >=0.8 | Human_DiffDom>=0.8) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=80)
BothRes_90=BothRes %>% filter(Chimp_DiffDom >=0.9 | Human_DiffDom>=0.9) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=90)

BothResAll=BothRes_10 %>% bind_rows(BothRes_20) %>% bind_rows(BothRes_30) %>% bind_rows(BothRes_40) %>% bind_rows(BothRes_50) %>% bind_rows(BothRes_60) %>% bind_rows(BothRes_70) %>% bind_rows(BothRes_80) %>% bind_rows(BothRes_90)

```
I want dominance in 1 or both at .4.  

```{r}
BothRes_40_each= BothRes_40 %>% mutate(Dom=ifelse(Human_DiffDom>=.4, ifelse(Chimp_DiffDom >=.4, "Both", "Human"), "Chimp"))

BothRes_40_each %>% group_by(Dom) %>% summarise(n())

BothRes_40_each %>% group_by(Set,Dom) %>% summarise(n())
```

```{r}
BothRes_40_eachsm= BothRes_40_each %>% select(gene, Set, Dom)


BothResInfoDom= BothResInfo %>% full_join(BothRes_40_eachsm, by="gene", fill="None") %>%  mutate(Set= replace_na(Set, "None"),Dom= replace_na(Dom, "None"))


ggplot(BothResInfoDom,aes(x=Human_Base2,y= Chimp_Base2, col=Dom )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") + labs(x="Human Information", y="Chimp Information", title="Shannon Information Index colored by whether gene has a dominant PAS")


ggplot(BothResInfoDom,aes(x=Human_Base2,y= Chimp_Base2, col=Set )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") +geom_density2d()+ labs(x="Human Information", y="Chimp Information", title="Shannon Information Index colored by Dominance Structure ")
```

```{r}

BothResInfoDom$numPAS=as.factor(BothResInfoDom$numPAS)
ggplot(BothResInfoDom,aes(x=Human_Base2,y= Chimp_Base2, col=numPAS )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + labs(x="Human Information", y="Chimp Information", title="Shannon Information Index colored by number of PAS") + facet_grid(~Dom)

#+ scale_color_brewer(palette = "Spectral")
```

Dominance and number of PAS:

```{r}
BothResInfoDom$numPAS=as.numeric(as.character(BothResInfoDom$numPAS))
ggplot(BothResInfoDom,aes(x=Dom, y=numPAS)) +geom_boxplot() +stat_compare_means() +  labs(x="Dominance Structure",y="Number of PAS", title="Number of PAS differ by dominance structure")


ggplot(BothResInfoDom,aes(x=Set, y=numPAS)) +geom_boxplot() +stat_compare_means() +  labs(x="Dominance Structure",y="Number of PAS", title="Number of PAS differ by dominance structure")
```



Ratio problem!!!! 

but the confounder is biological- number of PAS.

##Simpson

Try the simpson index. 

skit-bio: http://scikit-bio.org/docs/0.1.3/math.diversity.alpha.html

$D=\sum_{i=1}^{R}p_{i}^{2}$

and  

$D=1-\sum_{i=1}^{R}p_{i}^{2}$



```{bash,eval=F}
python infoContentSimpson.py Human
python infoContentSimpson.py Chimp

```

```{r}
SimpHuman=read.table("../data/InfoContent/Human_SimpsonInfoContent.txt", header = T, stringsAsFactors = F) %>% rename(simpson_Human=simpson) %>% mutate(simpOpp_Human=1-simpson_Human)
SimpChimp=read.table("../data/InfoContent/Chimp_SimpsonInfoContent.txt", header = T, stringsAsFactors = F)%>% rename(simpson_Chimp=simpson)%>% mutate(simpOpp_Chimp=1-simpson_Chimp)

BothSimp= SimpHuman %>% inner_join(SimpChimp, by=c("gene", "numPAS")) %>% filter(numPAS > 1)
```

Gather and plot:

```{r}
BothSimp_g= BothSimp %>% select(-contains("Opp")) %>% gather("species", "Simpson", -gene, -numPAS)
ggplot(BothSimp_g, aes(x=Simpson, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Set1")+labs(title="Simpson Index")
```

```{r}
BothOppSimp_g= BothSimp %>% select(-contains("simpson")) %>% gather("species", "SimpsonOpp", -gene, -numPAS)

ggplot(BothOppSimp_g, aes(x=SimpsonOpp, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Set1")+labs(title="Simpson Index (1-opp)")


wilcox.test(BothSimp$simpOpp_Human, BothSimp$simpOpp_Chimp, alternative = "greater")
```


Histogram: 

```{r}
ggplot(BothSimp_g,aes(x=Simpson, fill=species)) + geom_histogram() + facet_grid(~species)+scale_fill_brewer(palette = "Set1")

ggplot(BothOppSimp_g, aes(x=SimpsonOpp, fill=species)) + geom_histogram() + facet_grid(~species)+scale_fill_brewer(palette = "Set1")

```

here, higher index is lower diversity= more dominance (opposite of shannon)  

the opposite one is 1- sum. this is more dominance at lower values like shannon. I will go with this so the sign is the same.  


```{r}
BothInfoTypes=BothSimp %>% inner_join(BothResInfoRes, by=c("gene", "numPAS"))

BothInfoTypes_h=BothInfoTypes %>% select(gene,numPAS, simpOpp_Human, Human_Base2) %>% mutate(species="Human") %>% rename(Simpson= simpOpp_Human, Shannon=Human_Base2)
BothInfoTypes_c=BothInfoTypes %>% select(gene,numPAS, simpOpp_Chimp, Chimp_Base2) %>% mutate(species="Chimp")%>% rename(Simpson= simpOpp_Chimp, Shannon=Chimp_Base2)

BothInfoTypes_both=BothInfoTypes_h %>% bind_rows(BothInfoTypes_c)

ggplot(BothInfoTypes_both,aes(x=Simpson, y=Shannon, by=species, col=species)) +geom_point(alpha=.4) +geom_density2d(col="black") +  stat_cor(label.x=0) + geom_smooth(col="black",method = "lm") + facet_grid(~species) + labs(title="Correlation between Indicies") +theme(legend.position = "none")+scale_color_brewer(palette = "Set1")

```

There is more variation at the low end here. 

Compare human and chimp simpson by PAS number:  

```{r}
ggplot(BothInfoTypes,aes(x=simpOpp_Human,y= simpOpp_Chimp)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0)+labs(title="Simpson Index") + stat_cor(col="blue")+ geom_density_2d(col="blue")

ggplot(BothInfoTypes,aes(x=simpOpp_Human,y= simpOpp_Chimp ,col=numPAS)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0)+labs(title="Simpson Index Colored by number of PAS")



```

```{r}
summary(lm(BothInfoTypes$simpOpp_Human ~BothResInfo$numPAS))

cor.test(BothInfoTypes$simpOpp_Human,BothResInfo$numPAS)

summary(lm(BothInfoTypes$simpOpp_Chimp ~BothResInfo$numPAS ))

cor.test(BothInfoTypes$simpOpp_Chimp,BothResInfo$numPAS)
```

Number of PAS is less correlated with this index. 

Add in the dominanace structure to compare to simpson: 

```{r}

BothResBothInfoDom= BothInfoTypes %>% full_join(BothRes_40_eachsm, by="gene", fill="None") %>%  mutate(Set= replace_na(Set, "None"),Dom= replace_na(Dom, "None"))


ggplot(BothResBothInfoDom,aes(x=simpOpp_Human,y= simpOpp_Chimp, col=Dom )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") + labs(x="Human Simpson", y="Chimp Simpson", title="Simpson Information Index colored by whether gene has a dominant PAS")


ggplot(BothResBothInfoDom,aes(x=simpOpp_Human,y= simpOpp_Chimp, col=Set )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") +geom_density2d()+ labs(x="Human Simpson", y="Chimp Simpson", title="Simpson Information Index colored by Dominance Structure ")
```

```{r}
BothResBothInfoDom$numPAS=as.factor(BothResBothInfoDom$numPAS)
ggplot(BothResBothInfoDom,aes(x=simpOpp_Human,y= simpOpp_Chimp, col=numPAS )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + labs(x="Human Simpson", y="Chimp Simpson", title="Simpson Information Index colored by number of PAS ") + facet_grid(~Dom)


#+ scale_color_brewer(palette = "Spectral")
```
##Shannon Equitability
Equitability. Shannon diversity divided by the logarithm of number of taxa. This measures the evenness with which individuals are divided among the taxa present.

Shannon’s equitability (EH) measures the evenness of a community and can be easily calculated by diving the value of H with H_max, which equals to lnS(S=number of species encountered). Its value ranges between 0 and 1, with being complete evenness. (0-1)

$E_{h}=H/log2(NumPAS)$

```{r}
BothResBothInfoDomEH=BothResBothInfoDom %>% mutate(human_EH=Human_Base2/log2(as.numeric(as.character(numPAS))), chimp_EH=Chimp_Base2/log2(as.numeric(as.character(numPAS))))


BothEH= BothResBothInfoDomEH %>% select(gene, numPAS, human_EH,chimp_EH) %>% gather("species", "ShannonEH", -gene, -numPAS)

ggplot(BothEH, aes(x=ShannonEH, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Set1")+labs(title="Shannon equitability", x="Shannon equitability")


wilcox.test(BothResBothInfoDomEH$human_EH, BothResBothInfoDomEH$chimp_EH)
```

```{r}
BothResBothInfoDomEH$numPAS=as.numeric(as.character(BothResBothInfoDomEH$numPAS))
ggplot(BothResBothInfoDomEH,aes(x=human_EH,y= chimp_EH )) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0) +labs(title="Shannon equitability") + stat_cor(col="blue")+ geom_density_2d(col="blue")



BothResBothInfoDomEH$numPAS=as.numeric(as.character(BothResBothInfoDomEH$numPAS))
ggplot(BothResBothInfoDomEH,aes(x=human_EH,y= chimp_EH ,col=numPAS)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0) +labs(title="Shannon equitability Colored by number of PAS")

```
```{r}
summary(lm(BothResBothInfoDomEH$human_EH ~BothResBothInfoDomEH$numPAS))


summary(lm(BothResBothInfoDomEH$chimp_EH ~BothResBothInfoDomEH$numPAS ))
```
This normalizes the number of PAS.  

Correlation between values: 

```{r}

BothInfoTypes_eh_h=BothResBothInfoDomEH %>% select(gene,numPAS, simpOpp_Human, human_EH) %>% mutate(species="Human") %>% rename(Simpson= simpOpp_Human, ShannonEH=human_EH)
BothInfoTypes_eh_c=BothResBothInfoDomEH %>% select(gene,numPAS, simpOpp_Chimp, chimp_EH) %>% mutate(species="Chimp")%>% rename(Simpson= simpOpp_Chimp, ShannonEH=chimp_EH)

BothInfoTypes_bothEH=BothInfoTypes_eh_h %>% bind_rows(BothInfoTypes_eh_c)

ggplot(BothInfoTypes_bothEH,aes(x=Simpson, y=ShannonEH, by=species, col=species)) +geom_point(alpha=.4) +geom_density2d(col="black") +  stat_cor(label.x=0) + geom_smooth(col="black",method = "lm") + facet_grid(~species) + labs(title="Correlation between Indicies") +theme(legend.position = "none")+scale_color_brewer(palette = "Set1")
```


Look at it with dominance: 

```{r}
ggplot(BothResBothInfoDomEH,aes(x=human_EH,y= chimp_EH, col=Dom )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") + labs(x="Human equitability", y="Chimp equitability", title="Shannon equitability colored by whether gene has a dominant PAS")


ggplot(BothResBothInfoDomEH,aes(x=human_EH,y= chimp_EH, col=Set )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + scale_color_brewer(palette = "Set2") +geom_density2d()+ labs(x="Human equitability", y="Chimp equitability", title="Shannon equitability colored by Dominance Structure")
```

```{r}
BothResBothInfoDomEH$numPAS=as.factor(BothResBothInfoDomEH$numPAS)
ggplot(BothResBothInfoDomEH,aes(x=human_EH,y= chimp_EH, col=numPAS )) + geom_point(alpha=.3) + geom_abline(slope=1, intercept = 0) + labs(x="Human equitability", y="Chimp equitability", title="Shannon Equitability colored by number of PAS ") + facet_grid(~Dom)


```


plot simpson h/c colors:  

```{r}
ggplot(BothOppSimp_g, aes(x=SimpsonOpp, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Dark2",labels=c("Chimp", "Human"))+labs(title="Simpson Index", x="Simpson Index")+ theme_classic2()
```





Plot only 1 color to demonstrate:  

```{r}
ggplot(BothOppSimp_g, aes(x=SimpsonOpp )) + geom_density(fill="grey") +labs(title="Simpson Index", x="Simpson Index")+ theme_classic2()
```


Plot number of PAS and info content to use:  

```{r}
ggplot(BothInfoTypes,aes(x=simpOpp_Human,y= simpOpp_Chimp ,col=numPAS)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0)+labs(title="Simpson Index Colored by number of PAS", x="Human", y="Chimp") + theme_classic()



```

Plot shannon with HC colors 
```{r}
ggplot(BothResInfo_2, aes(x=base2, fill=species)) + geom_density(alpha=.5) + scale_fill_brewer(palette = "Dark2",labels=c("Chimp", "Human"))+labs(title="Shannon Index", x="Simpson Index")+ theme_classic2()
```


```{r}
ggplot(BothInfoTypes,aes(x=Human_Base2,y= Chimp_Base2 ,col=numPAS)) + geom_point(alpha=.4) + geom_abline(slope=1, intercept = 0)+labs(title="Shannon Index Colored by number of PAS", x="Human", y="Chimp") + theme_classic()


```


Plot number of PAS by index:

```{r}
ggplot(BothInfoTypes,aes(x=numPAS,y= Chimp_Base2 )) + geom_point(alpha=.4) + stat_cor()+ theme_classic()

ggplot(BothInfoTypes,aes(x=numPAS,y= Human_Base2 )) + geom_point(alpha=.4) + stat_cor()+ theme_classic()
```
```{r}
BothInfoTypesShanG= BothInfoTypes %>% select(gene, numPAS,Human_Base2,Chimp_Base2 )  %>% rename(Human=Human_Base2, Chimp=Chimp_Base2) %>% gather("Species", "value", -gene, -numPAS) 

shanoNum=ggplot(BothInfoTypesShanG,aes(x=numPAS,y= value ,col=Species)) + geom_point(alpha=.4) + stat_cor(col="black",label.y.npc="bottom")+ theme_classic() + facet_grid(~Species) + scale_color_brewer(palette = "Dark2") + labs(y="Shannon Information Content", title="Shannon Information Content and PAS number", x= "number of PAS in gene") 

shanoNum
```
```{r}
BothInfoTypesSimpG= BothInfoTypes %>% select(gene, numPAS,simpOpp_Human,simpOpp_Chimp )  %>% rename(Human=simpOpp_Human, Chimp=simpOpp_Chimp) %>% gather("Species", "value", -gene, -numPAS) 

simpnum=ggplot(BothInfoTypesSimpG,aes(x=numPAS,y= value ,col=Species)) + geom_point(alpha=.4) + stat_cor(col="black",label.y.npc="bottom")+ theme_classic() + facet_grid(~Species) + scale_color_brewer(palette = "Dark2")+ labs(y="Simpson Diversity", title="Simpson Diversity and PAS number", x= "number of PAS in gene") 

simpnum
```


```{r}
plot_grid(shanoNum,simpnum, nrow=2)
```

