---
title: "Information Content and dAPA"
author: "Briana Mittleman"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to look at the simpon information content results for genes with and without dAPA.  

```{r}
library(tidyverse)
library(ggpubr)
library(cowplot)
library(workflowr)
```

```{r}
SimpHuman=read.table("../data/InfoContent/Human_SimpsonInfoContent.txt", header = T, stringsAsFactors = F) %>% rename(simpson_Human=simpson) %>% mutate(simpOpp_Human=1-simpson_Human)
SimpChimp=read.table("../data/InfoContent/Chimp_SimpsonInfoContent.txt", header = T, stringsAsFactors = F)%>% rename(simpson_Chimp=simpson)%>% mutate(simpOpp_Chimp=1-simpson_Chimp)

BothSimp= SimpHuman %>% inner_join(SimpChimp, by=c("gene", "numPAS")) %>% filter(numPAS > 1)

dAPAGenes=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt", header=T,stringsAsFactors=F)
DiffIso=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header=T,stringsAsFactors = F) %>% select(gene) %>% unique() %>% mutate(dAPA=ifelse(gene %in% dAPAGenes$gene, "Yes", "No"))
```


Inner join the simpson results and the diff iso results. I want the results to be based on genes I could test in both.  

```{r}
BothSimp_dAPA=BothSimp %>% inner_join(DiffIso, by="gene")

```
 More dominance is the lower index.  

```{r}
BothSimp_dAPA_g= BothSimp_dAPA %>% select(gene, simpOpp_Human, simpOpp_Chimp, dAPA) %>% gather("speciesS", "simpson", -dAPA, -gene) %>% mutate(species=ifelse(grepl("Human",speciesS), "Human", "Chimp"))


ggplot(BothSimp_dAPA_g,aes(x=simpson, by=species, fill=species)) + geom_density(alpha=.4) +facet_grid(~dAPA)+ scale_fill_brewer(palette = "Set1")+labs(title="Simpson Index by dAPA")
```


```{r}
ggplot(BothSimp_dAPA_g,aes(x=simpson, by=dAPA, fill=dAPA)) + geom_density(alpha=.4) +facet_grid(~species)+ scale_fill_brewer(palette = "Set1")+labs(title="Simpson Index by dAPA in each species")
```


Are the species correlations different.  

```{r}
BothSimp_dAPAYes= BothSimp_dAPA %>% filter(dAPA=="Yes")
nrow(BothSimp_dAPAYes)
summary(lm(BothSimp_dAPAYes$simpOpp_Human~BothSimp_dAPAYes$simpOpp_Chimp))


ggplot(BothSimp_dAPAYes, aes(x=simpOpp_Human, y=simpOpp_Chimp)) +geom_point() + stat_cor(col="blue")+ geom_density_2d(col="blue")
```

```{r}

BothSimp_dAPANo= BothSimp_dAPA %>% filter(dAPA=="No")
nrow(BothSimp_dAPANo)

summary(lm(BothSimp_dAPANo$simpOpp_Human~BothSimp_dAPANo$simpOpp_Chimp))


ggplot(BothSimp_dAPANo, aes(x=simpOpp_Human, y=simpOpp_Chimp)) +geom_point() + stat_cor(col="blue")+ geom_density_2d(col="blue")
```

Ok this makes sense. The correlation between the species is higher when there is no dAPA.  



```{r}
simp=ggplot(BothSimp_dAPA,aes(x=simpOpp_Human, y=simpOpp_Chimp,by=dAPA, col=dAPA)) +geom_point(alpha=.5) +scale_color_brewer(palette = "Set1")+stat_cor() + labs(x="Human", y="Chimp", title="Simpson index by dAPA")+ theme(legend.position = "bottom")
simp
```

Do this for the other values:  

```{r}
HumanResInfo= read.table("../data/InfoContent/Human_InfoContent.txt", header = T,stringsAsFactors = F) %>% rename(Human_Base2=base2, Human_basee= basee)
ChimpResInfo= read.table("../data/InfoContent/Chimp_InfoContent.txt", header = T,stringsAsFactors = F) %>% rename(Chimp_Base2=base2, Chimp_basee= basee)

BothResInfo= HumanResInfo %>% inner_join(ChimpResInfo, by=c("gene", "numPAS")) %>% filter(numPAS > 1)%>% inner_join(DiffIso, by="gene")


shan=ggplot(BothResInfo,aes(x=Human_Base2, y=Chimp_Base2,by=dAPA, col=dAPA)) +geom_point(alpha=.5) +scale_color_brewer(palette = "Set1")+stat_cor() + labs(x="Human", y="Chimp", title="Shannon index by dAPA")+ theme(legend.position = "bottom")
shan
```


```{r}
BothResBothInfoDomEH=BothResInfo %>% mutate(human_EH=Human_Base2/log2(as.numeric(as.character(numPAS))), chimp_EH=Chimp_Base2/log2(as.numeric(as.character(numPAS)))) 


eh=ggplot(BothResBothInfoDomEH,aes(x=human_EH, y=chimp_EH,by=dAPA, col=dAPA)) +geom_point(alpha=.5) +scale_color_brewer(palette = "Set1")+stat_cor() + labs(x="Human", y="Chimp", title="Shannon equitability by dAPA") + theme(legend.position = "bottom")

eh
```

```{r}
plot_grid(shan,eh,simp, nrow=1)
```

