---
title: "Liftover PAS"
author: "Briana Mittleman"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)

```

Next step will be to prepare these for liftover. I will get the PAS as the furthest downstream base. I will liftover the regions 100bp upstream and 100bp downstream of the PAS.


As of now the PAS are still on the opposite strand. 

Pos strand: pas is the end
  - start_new= end-100
  - end_new=end + 100
  
neg strand: pas is the start
  - start_new= start-100
  - end_new= start + 100

human: /project2/gilad/briana/Comparative_APA/Human/data/cleanPeaks/human_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed
chimp: /project2/gilad/briana/Comparative_APA/Chimp/data/cleanPeaks/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed

Output:

```{bash,eval=F}
mkdir ../data/cleanPeaks_byspecies/

python preparePAS4lift.py ../Human/data/cleanPeaks/human_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed ../data/cleanPeaks_byspecies/human_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed

python preparePAS4lift.py ../Chimp/data/cleanPeaks/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed ../data/cleanPeaks_byspecies/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed



```


Chain files from:
http://hgdownload.soe.ucsc.edu/downloads.html#chimp dowload to ../data/chainFiles/


Liftover pipeline:

start with human- lift to chimp and back
start with chimp- lift to human and back 

```{bash,eval=F}
mkdir ../data/primaryLift
mkdir ../data/reverseLift

sbatch primaryLift.sh
sbatch reverseLift.sh
```


Results from primary lift:  

```{r}
unliftedH=read.table("../data/primaryLift/human_APApeaks_primarylift2Chimp_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()
unliftedC=read.table("../data/primaryLift/chimp_APApeaks_primarylift2Human_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()

liftedH=read.table("../data/primaryLift/human_APApeaks_primarylift2Chimp.bed",stringsAsFactors = F) %>% nrow()
liftedC=read.table("../data/primaryLift/chimp_APApeaks_primarylift2Human.bed",stringsAsFactors = F) %>% nrow()

primaryUnC=c("Chimp","Unlifted", unliftedC)
primaryUnH=c("Human","Unlifted", unliftedH)

primaryLH=c("Human","Lifted", liftedH)
primaryLC=c("Chimp","Lifted", liftedC)

header=c("species", "liftStat", "PAS")
primaryDF= as.data.frame(rbind(primaryLH,primaryLC, primaryUnH,primaryUnC)) 
colnames(primaryDF)=header


primaryDF$PAS=as.numeric(as.character(primaryDF$PAS)) 



primaryDF= primaryDF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)

```



```{r}
ggplot(primaryDF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results")
```

```{r}
ggplot(primaryDF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results")
```


Reverse lift:  

```{r}
re_unliftedH=read.table("../data/reverseLift/human_APApeaks_primarylift2Human_rev2Human_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()
re_unliftedC=read.table("../data/reverseLift/chimp_APApeaks_primarylift2Human_rev2Chimp_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()

re_liftedH=read.table("../data/reverseLift/human_APApeaks_primarylift2Chimp_rev2Human.bed",stringsAsFactors = F) %>% nrow()
re_liftedC=read.table("../data/reverseLift/chimp_APApeaks_primarylift2Human_rev2Chimp.bed",stringsAsFactors = F) %>% nrow()

re_UnC=c("Chimp","Unlifted", re_unliftedC)
re_UnH=c("Human","Unlifted", re_unliftedH)

re_LH=c("Human","Lifted", re_liftedH)
re_LC=c("Chimp","Lifted", re_liftedC)

header=c("species", "liftStat", "PAS")
re_DF= as.data.frame(rbind(re_LH,re_LC, re_UnH,re_UnC)) 
colnames(re_DF)=header


re_DF$PAS=as.numeric(as.character(re_DF$PAS)) 



re_DF= re_DF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)

```


```{r}
ggplot(re_DF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Reverse Liftover Results")
```

```{r}
ggplot(re_DF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2")+ labs(title="Reverse Liftover Results")
```

I need to now make sure they lifted to the same location. To do this I will overlap the reciprocal lifted PAS with the original files.  

I subset the original files by the pas that have an exact match in the reverse map. I can do this by pas name- start:end  

```{bash,eval=F}
mkdir ../data/cleanPeaks_lifted
python filterPostLift.py ../data/cleanPeaks_byspecies/human_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed ../data/reverseLift/human_APApeaks_primarylift2Chimp_rev2Human.bed ../data/cleanPeaks_lifted/Human_PASregions.bed

python filterPostLift.py ../data/cleanPeaks_byspecies/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed  ../data/reverseLift/chimp_APApeaks_primarylift2Human_rev2Chimp.bed ../data/cleanPeaks_lifted/Chimp_PASregions.bed
```


Results: 

```{r}
Human_recLift=read.table("../data/cleanPeaks_lifted/Human_PASregions.bed",stringsAsFactors = F, col.names = c("chr", "start","end", "name", "score", "strand")) 
Chimp_recLift=read.table("../data/cleanPeaks_lifted/Chimp_PASregions.bed",stringsAsFactors = F,col.names = c("chr", "start","end", "name", "score", "strand"))

originalH=unliftedH + liftedH
originalC=unliftedC + liftedC

#human
nrow(Human_recLift)/originalH


#chimp 
nrow(Chimp_recLift)/originalC
```


96% reciprocally lifted over

lift the chimp ones back to human

```{bash,eval=F}
sbatch recLiftchim2human.sh
```


Join the results: If they are discovered in both say so. 

I need to intersect these with bedtools to know when a PAS is 


```{bash,eval=F}
sort -k1,1 -k2,2n ../data/cleanPeaks_lifted/Chimp_PASregions_humanCoord.bed > ../data/cleanPeaks_lifted/Chimp_PASregions_humanCoord.sort.bed


sort -k1,1 -k2,2n ../data/cleanPeaks_lifted/Human_PASregions.bed > ../data/cleanPeaks_lifted/Human_PASregions.sort.bed
```

I used 75% overlap for the observed in both. I had it report number of basepair overlap so i can evaluate this. If it is bimodal i will change this.  (i chose 75% because its 50bp or within 1 read.)

```{bash,eval=F}
 sbatch intersectLiftedPAS.sh
```

```{r}
OverlapBoth_Test=read.table("../data/cleanPeaks_lifted/PASregions_identifiedbothTEST.txt", col.names = c("Hchr", "Hstart","Hend", "Hname", "Hscore", "Hstrand", "Cchr", "Cstart", "Cend", "Cname", "Cscore", "Cstrand", "overlap"),stringsAsFactors = F)

ggplot(OverlapBoth_Test, aes(x=overlap)) + geom_histogram(bins=30) + scale_y_log10() + geom_vline(xintercept = 125)

125/200
```
I will go with 62.5% overlap. I can prepare the files to make a full set. 

```{r}


OverlapBoth=read.table("../data/cleanPeaks_lifted/PASregions_identifiedboth.txt", col.names = c("Hchr", "Hstart","Hend", "Hname", "Hscore", "Hstrand", "Cchr", "Cstart", "Cend", "Cname", "Cscore", "Cstrand", "overlap"),stringsAsFactors = F) %>% mutate(meanScore=(Hscore+Cscore)/2, name=paste("Both", Hname, sep=":"), Bothname=paste(Hname, Cname, sep=":"))

overlap2= OverlapBoth %>% group_by(name) %>% filter(n()>1) %>%  mutate(id = row_number()) %>% filter(id==2)

OverlapBoth_format=read.table("../data/cleanPeaks_lifted/PASregions_identifiedboth.txt", col.names = c("Hchr", "Hstart","Hend", "Hname", "Hscore", "Hstrand", "Cchr", "Cstart", "Cend", "Cname", "Cscore", "Cstrand", "overlap"),stringsAsFactors = F) %>% mutate(meanScore=(Hscore+Cscore)/2, name=paste("Both", Hname, sep=":"),Bothname=paste(Hname, Cname, sep=":")) %>% filter(!Bothname %in% overlap2$Bothname) %>% dplyr::select(Hchr,Hstart,Hend,name,meanScore,Hstrand) 



HumanSpec=read.table("../data/cleanPeaks_lifted/PASregions_identifiedHuman.txt", col.names = c("Hchr", "Hstart","Hend", "Hname", "meanScore", 'Hstrand'),stringsAsFactors = F) %>% mutate(name=paste("Human", Hname, sep=":")) %>% dplyr::select(Hchr, Hstart,Hend, name, meanScore, Hstrand) 


ChimpSpec=read.table("../data/cleanPeaks_lifted/PASregions_identifiedChimp.txt", col.names = c("Hchr", "Hstart","Hend", "Cname", "meanScore", 'Hstrand'),stringsAsFactors = F) %>% mutate(name=paste("Chimp", Cname, sep=":")) %>% dplyr::select(Hchr, Hstart,Hend, name, meanScore, Hstrand)



```
Join all of these and plot characteristics 
```{r}
AllPAS=as.data.frame(rbind(OverlapBoth_format,HumanSpec,ChimpSpec)) %>% separate(name, into=c("discovery", "PAS"))


ggplot(AllPAS, aes(x=discovery, fill=discovery)) + geom_bar(stat="count")+ scale_fill_brewer(palette = "Dark2")

ggplot(AllPAS, aes(x=discovery, fill=discovery)) +  geom_bar(aes(y = (..count..)/sum(..count..)))+ scale_fill_brewer(palette = "Dark2")
```

Plot the difference in scores for both and discovered in 1 or other:

```{r}
ggplot(AllPAS, aes(x=discovery, fill=discovery, y=meanScore)) + geom_boxplot() + scale_y_log10()+ scale_fill_brewer(palette = "Dark2")
```

This is expected. Those found in both will be used more often. I expect many of those only discovered in 1 will drop out at the 5% cutoff.  

```{r}
AllPAS_use=as.data.frame(rbind(OverlapBoth_format,HumanSpec,ChimpSpec))

write.table(AllPAS_use,  "../data/cleanPeaks_lifted/AllPAS_postLift.bed", col.names = F, row.names = F, quote = F, sep = "\t") 



```
sort these: 
```{bash, eval=F}
sort -k1,1 -k2,2n ../data/cleanPeaks_lifted/AllPAS_postLift.bed > ../data/cleanPeaks_lifted/AllPAS_postLift.sort.bed
```

