---
title: "Liftover PAS"
author: "Briana Mittleman"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)

```

Next step will be to prepare these for liftover. I will get the PAS as the furthest downstream base. I will liftover the regions 100bp upstream and 100bp downstream of the PAS.


As of now the PAS are still on the opposite strand. 

Pos strand: pas is the end
  - start_new= end-100
  - end_new=end + 100
  
neg strand: pas is the start
  - start_new= start-100
  - end_new= start + 100

human: /project2/gilad/briana/Comparative_APA/Human/data/cleanPeaks/human_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed
chimp: /project2/gilad/briana/Comparative_APA/Chimp/data/cleanPeaks/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed

Output:

```{bash,eval=F}
mkdir ../data/cleanPeaks_byspecies/

python preparePAS4lift.py ../Human/data/cleanPeaks/human_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed ../data/cleanPeaks_byspecies/human_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed

python preparePAS4lift.py ../Chimp/data/cleanPeaks/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned.bed ../data/cleanPeaks_byspecies/chimp_APApeaks.ALLChrom.Filtered.Named.Cleaned_100bpreg.bed



```


Chain files from:
http://hgdownload.soe.ucsc.edu/downloads.html#chimp dowload to ../data/chainFiles/


Liftover pipeline:

start with human- lift to chimp and back
start with chimp- lift to human and back 

```{bash,eval=F}
mkdir ../data/primaryLift
mkdir ../data/reverseLift

sbatch primaryLift.sh
```


Results from primary lift:  

```{r}
unliftedH=read.table("../data/primaryLift/human_APApeaks_primarylift2Chimp_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()
unliftedC=read.table("../data/primaryLift/chimp_APApeaks_primarylift2Human_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()

liftedH=read.table("../data/primaryLift/human_APApeaks_primarylift2Chimp.bed",stringsAsFactors = F) %>% nrow()
liftedC=read.table("../data/primaryLift/chimp_APApeaks_primarylift2Human.bed",stringsAsFactors = F) %>% nrow()

primaryUnC=c("Chimp","Unlifted", unliftedC)
primaryUnH=c("Human","Unlifted", unliftedH)

primaryLH=c("Human","Lifted", liftedH)
primaryLC=c("Chimp","Lifted", liftedC)

header=c("species", "liftStat", "PAS")
primaryDF= as.data.frame(rbind(primaryLH,primaryLC, primaryUnH,primaryUnC)) 
colnames(primaryDF)=header


primaryDF$PAS=as.numeric(as.character(primaryDF$PAS)) 



primaryDF= primaryDF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)

```



```{r}
ggplot(primaryDF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results")
```

```{r}
ggplot(primaryDF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Primary Liftover Results")
```


Reverse lift:  

```{r}
re_unliftedH=read.table("../data/reverseLift/human_APApeaks_primarylift2Human_rev2Human_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()
re_unliftedC=read.table("../data/reverseLift/chimp_APApeaks_primarylift2Human_rev2Chimp_UNLIFTED.bed",stringsAsFactors = F) %>% nrow()

re_liftedH=read.table("../data/reverseLift/human_APApeaks_primarylift2Chimp_rev2Human.bed",stringsAsFactors = F) %>% nrow()
re_liftedC=read.table("../data/reverseLift/chimp_APApeaks_primarylift2Human_rev2Chimp.bed",stringsAsFactors = F) %>% nrow()

re_UnC=c("Chimp","Unlifted", re_unliftedC)
re_UnH=c("Human","Unlifted", re_unliftedH)

re_LH=c("Human","Lifted", re_liftedH)
re_LC=c("Chimp","Lifted", re_liftedC)

header=c("species", "liftStat", "PAS")
re_DF= as.data.frame(rbind(re_LH,re_LC, re_UnH,re_UnC)) 
colnames(re_DF)=header


re_DF$PAS=as.numeric(as.character(re_DF$PAS)) 



re_DF= re_DF %>% group_by(species) %>% mutate(nPAS=sum(PAS)) %>% ungroup() %>% mutate(proportion=PAS/nPAS)

```


```{r}
ggplot(re_DF,aes(x=species, y=PAS, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2") + labs(title="Reverse Liftover Results")
```

```{r}
ggplot(re_DF,aes(x=species, y=proportion, fill=liftStat)) + geom_bar(stat="identity",position = "dodge") + scale_fill_brewer(palette = "Dark2")+ labs(title="Reverse Liftover Results")
```

I need to now make sure they lifted to the same location. To do this I will overlap the reciprocal lifted PAS with the original files.  


