---
title: "Differential usage by mean usage"
author: "Briana Mittleman"
date: "6/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Answer question: likelihood of being differential based on usage proportion?

Start with plotting cumulative distribution of usage by differential usage, both in human and chimp. 

I need the mean usage in both species and if the site is differentially used: 

```{r}
library(workflowr)
library(tidyverse)
library(ggpubr)
library(cowplot)
```


```{r}
PASmeta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)

PASmetaSmall=PASmeta %>% select(PAS, chr, start, end, gene)
```


Diff used sites:  

```{r}
DiffIso= read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", header = T,stringsAsFactors = F) %>% inner_join(PASmetaSmall, by=c("chr", 'start','end', "gene")) %>% select(PAS,SigPAU2)

```


Usage: 

```{r}
PASMetaDiff= PASmeta %>% select(PAS,Chimp,Human)  %>% inner_join(DiffIso, by="PAS") %>% gather(Species, Usage, -PAS, -SigPAU2)
```


```{r}
ggplot(PASMetaDiff, aes(x=Usage, col=SigPAU2)) + stat_ecdf()+ scale_color_brewer(palette = "Set1") + facet_grid(~Species)
```


Look at usage by which PAS for those in ortho UTR  

1252 PA

```{r}
OrthoUTR=read.table("../data/orthoUTR/OrthoPASwithLocandSig.txt", header =T, stringsAsFactors = F)


```

First look at usage vs first, middle, last:  

```{r}
OrthoUTRspecusage=OrthoUTR %>% select(PAS,Chimp,Human, WhichSite,SigPAU2) %>% gather("Species","Usage", -WhichSite, -SigPAU2, -PAS)
```

```{r}
useCOl <- c("#d73027", "#4575b4","#fee090")
ggplot(OrthoUTRspecusage, aes(x=Usage, col=WhichSite)) + stat_ecdf()+ scale_color_manual(values = useCOl) + facet_grid(~Species)
```
This shows that last actually has highest usage  

Now for just these look at usage
```{r}
ggplot(OrthoUTRspecusage, aes(x=Usage, col=SigPAU2)) + stat_ecdf()+ scale_color_brewer(palette = "Set1") + facet_grid(~Species)
```

Still an effect but smaller.  

Now plot all of theses together: 

```{r}
ggplot(OrthoUTRspecusage, aes(x=Usage, col=WhichSite)) + stat_ecdf()+ scale_color_manual(values = useCOl) + facet_grid(SigPAU2~Species)
```
NUmber in sets: 

```{r}
OrthoUTRspecusage %>% group_by(Species,SigPAU2, WhichSite) %>% summarise(n=n()) %>% filter(Species=="Human") %>% ungroup()%>% select(SigPAU2, WhichSite, n)
```

With so few it is difficult to say much. But it does look like that the last site i usually higher used. The differentially used PAS are also the higher used. This means it is even more striking that the first sites are more often different.  
