---
title: "Map to hg19"
author: "Briana Mittleman"
date: "11/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


I have high levels of multi mapping in these samples. I want to test if this is the case in hg19 as well. I can try with 1 or 2 samples.

```{bash,eval=F}
mkdir ../Human/data/RNAseq/bam_hg19
sbatch mapRNAseqhg19.sh  
```

18504
hg19: multi map= 15.24 (83% map) 
hg38:  multi map= 21.39% (77.52% map)


18510-
hg19: multimap = 14.46% (83.97% map)
hg38: multimap= 20.49% (78.37%% map)


redo this with the new multimap criteria:  

```{bash,eval=F}
sbatch mapRNAseqhg19_newPipeline.sh  
```


New pipeline: 

18504
hg19: too many loci= 0.37%, multiple loci- 14.93% (83.19%) 
hg38: too many loci= 0.65%, multiple loci- 20.75% (77.50%)


18510-
hg19: too many loci= 0.34%, multiple loci- 14.17% (83.93%) 
hg38: too many loci= 0.60%, multiple loci- 19.89% (78.35%)


With the hg19 samples make the multimapping graphs.  

```{bash,eval=F}
sbatch hg19MapStats.sh

python numMultimap.py /project2/gilad/briana/Comparative_APA/data/MapStats/Human_hg19_RNAseq_mapstats.txt /project2/gilad/briana/Comparative_APA/data/MapStats/Humanhg19_groupedMultstat.txt

```


```{r}
HumanHG19= read.table("../data/MapStats/Humanhg19_groupedMultstat.txt", header = F, col.names = c("HI", "Hg19Count"), stringsAsFactors = F) %>% separate(HI, into=c("N", "i", "num"),sep=":") 
HumanHG19$num=as.integer(as.character(HumanHG19$num))
ggplot(HumanHG19, aes(x=num, y=Hg19Count)) + geom_bar(stat="identity") + labs(x="Number of loacations read maps to", y="Reads", title="Multimapping statistics for hg19 RNA seq") 
```

Add the hg38 human chimp.  


```{r}
HumanMM= read.table("../data/MapStats/Human_groupedMultstat.txt", header = F, col.names = c("HI", "Human"), stringsAsFactors = F)
ChimpMM=read.table("../data/MapStats/Chimp_groupedMultstat.txt",header = F, col.names = c("HI", "Chimp"),stringsAsFactors = F)
HumanHG19_2= read.table("../data/MapStats/Humanhg19_groupedMultstat.txt", header = F, col.names = c("HI", "Hg19"), stringsAsFactors = F)


allMM=HumanMM %>% inner_join(ChimpMM ,by="HI") %>% inner_join(HumanHG19_2, by="HI")%>% gather("species", "Count", Human:Hg19) %>% separate(HI, into=c("N", "i", "num"),sep=":")  %>% group_by(species) %>% mutate(fullCount=sum(Count), prop=Count/fullCount)

allMM$num=as.integer(as.character(allMM$num))
ggplot(allMM, aes(x=num, y=prop, fill=species,by=species)) + geom_bar(position = "dodge" ,stat="identity") + labs(x="Number of loacations read maps to", y="Reads", title="Multimapping statistics for RNA seq") + scale_fill_brewer(palette = "Dark2")
```



