---
title: "Mediation in protein"
author: "Briana Mittleman"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(workflowr)
library(qvalue)
library(edgeR)
library(tidyverse)
library(limma)
library(MASS)
```


In this analysis, I will ask if differrecnes in APA  are causal for differences in protein.  


To deal with multiple phenotypes per gene I will take the highest absolute effect size PAS.

```{bash,eval=F}
mkdir ../data/mediation_prot
```

```{r}
source("../code/mediation_test.R") #Obtain necessary functions
```

I cant do this because we dont have the protein effect sizes. I need to remodel  

```{r}
ProtComp=read.csv("../data/Khan_prot/Khan_TableS4.csv", stringsAsFactors = F, header = T) %>% dplyr::select(gene.symbol, human.GM18505.protein,human.GM18507.protein,human.GM18516.protein,human.GM19193.protein,human.GM19204.protein,chimp.18358.protein,chimp.18359.protein ,chimp.3659.protein ,chimp.4973.protein,chimp.Pt91.protein ) %>% column_to_rownames(var="gene.symbol") %>% drop_na()
```
These are normalized values and filtered values. I will create a design matrix and test for differential protein.  


```{r}
ProtMeta=as.data.frame(cbind(indiv=colnames(ProtComp), Species=c(rep("H",5), rep("C",5))))

Species <- factor(ProtMeta$Species)
design <- model.matrix(~ 0 + Species)
head(design)

colnames(design) <- gsub("Species", "", dput(colnames(design)))
```

Make a object 

```{r}
labels <- paste(ProtMeta$Species,ProtMeta$indiv, sep=" ")


```
Use classic linear model:  

```{r}
gene1_lm= summary(lm(as.numeric(ProtComp[1,])~Species))
gene1_lm$coefficients[2,1]
gene1_lm$coefficients[2,4]
```
human is 1.  


Effect size is  -0.007561, pvalue is .9405  


Run for all genes:  

```{r}
Ppvals <- c()
Peffect <- c()
for (i in 1:dim(ProtComp)[1]) {
model <- summary(lm(as.numeric(ProtComp[i,]) ~ Species))
pval <-model$coefficients[2,4]
effect= model$coefficients[2,1]
Ppvals <- c(Ppvals, pval)
Peffect <- c(Peffect,effect)
}

```


Join results: 

```{r}
Pres=as.data.frame(cbind(genes=rownames(ProtComp), P=Ppvals, Ef=Peffect))
Pres$P=as.numeric(as.character(Pres$P))

Pres$Ef=as.numeric(as.character(Pres$Ef))

adjP=qvalue(Pres$P)
PresF= Pres %>% mutate(adjPval=adjP$pvalues)

SigP=PresF %>% filter(adjPval <=.05)
nrow(SigP)
NotSigP=PresF %>% filter(adjPval >.05)
nrow(NotSigP)


```

I can use this for the mediation.  - effect size in down in human. 

The individuals are not the same. I cannot do this analysis.  



