---
title: "Mediation analysis"
author: "Briana Mittleman"
date: "1/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(limma)
```

In this analysis, I will ask if differrecnes in APA or splicing are causal for differences in expression. I will use midiation analysis as implemented in (https://lauren-blake.github.io/Regulatory_Evol/analysis/Final_effect_size_human_chimp.html) using Joyces package mediation. 

Notes:  

* need one value for gene, sample, expression values, mol pheno values (normalized and non normalized)  

* uses ashR,vashr,medinome package

* df= sample -2 (4)

* run for DE and non DE genes seperatly 


Ittai's version of the analysis: https://ittaieres.github.io/HiCiPSC/gene_expression.html#now,_get_the_appropriate_data_and_actually_run_the_mediation_analysis

For this version I do not need the package. I only need limma. I will try this first.  

To deal with multiple phenotypes per gene I will take the highest absolute effect size cluster or PAS for splicing and APA respectively.  

```{r}
source("../code/mediation_test.R") #Obtain necessary functions
```

Now I need to pull in the expression, apa, and splicing data.  
- expression= log2RPKM, and adju.P.val
- normalized nuclear apa phenotype  
- splicing normalized clusters


###Expression and APA  
```{r}
#expression
nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F) %>% dplyr::select(Gene_stable_ID,Gene.name)
ExpRes=read.table("../data/DiffExpression/DEtested_allres.txt", header = F, stringsAsFactors = F, col.names = c("Gene_stable_ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")) %>% inner_join(nameID,by="Gene_stable_ID") %>% dplyr::select(Gene.name,adj.P.Val )
ExpNorm=read.table("../data/DiffExpression/NormalizedExpressionPassCutoff.txt", header = T, stringsAsFactors = F) %>% inner_join(nameID,by="Gene_stable_ID") %>% dplyr::select(-Gene_stable_ID)
#join indiv with adjust pval
Exp=ExpRes %>% inner_join(ExpNorm, by="Gene.name") %>% rename("gene"= Gene.name) %>% select(gene, adj.P.Val, NA18498, NA18499, NA18502, NA18504, NA18510,NA18523,NA18358, NA3622,NA3659,NA4973,NAPT30,NAPT91)

#Chr	start	end	ID	NA18498_N	NA18499_N	NA18502_N	NA18504_N	NA18510_N	NA18523_N	NA18358_N	NA3622_N	NA3659_N	NA4973_N	NApt30_N	NApt91_N
#apa:  
apaN= read.table("../data/Pheno_5perc_nuclear/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc_Nuclear.txt.gz.phen_AllChrom", col.names = c("chr", "start", "end", "id","NA18498_APA", "NA18499_APA", "NA18502_APA", "NA18504_APA", "NA18510_APA","NA18523_APA", "NA18358_APA", "NA3622_APA","NA3659_APA", "NA4973_APA","NAPT30_APA","NAPT91_APA")) %>% separate(id, into=c("ch", "st", "en","id2"),sep=":") %>% separate(id2, into=c("gene", "strand","id3"),sep="_") %>% separate(id3, into=c("loc", "disc", "PAS"), sep="-") %>% select(gene, PAS,contains("NA"))
PASMeta=read.table("../data/PAS/PAS_5perc_either_HumanCoord_BothUsage_meta.txt", header = T, stringsAsFactors = F) %>% dplyr::select(PAS, chr, start,end, gene)
apaRes= read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") %>% separate(cluster, into=c("chr","gene"),sep=":")
apaPASres=read.table("../data/DiffIso_Nuclear/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('intron',  'logef' ,'Human', 'Chimp','deltaPAU')) %>% filter(intron != "intron") %>% separate(intron, into=c("chr","start", "end","gene"), sep=":") 
apaPASres$start=as.integer(apaPASres$start)
apaPASres$end=as.integer(apaPASres$end)
apaPASres$deltaPAU=as.numeric(apaPASres$deltaPAU)
apaPASres=apaPASres%>% inner_join(PASMeta,by=c("chr", "start", "end", "gene"))
#problem if there are 2 pas then the are opposite but same value - do with all one direction for 
apaPASres_topPos= apaPASres %>% group_by(gene) %>% top_n(1,abs(deltaPAU)) %>% top_n(1,deltaPAU)
apaPASres_topNeg= apaPASres %>% group_by(gene) %>% top_n(1,abs(deltaPAU)) %>% top_n(-1,deltaPAU)
#join with pvalue: 
apaPASres_topPos_pval=apaPASres_topPos %>%  inner_join(apaRes, by=c("gene"))  %>% dplyr::select( PAS, p.adjust) 
apaPASres_topNeg_pval=apaPASres_topNeg %>%  inner_join(apaRes, by=c("gene")) %>% dplyr::select( PAS, p.adjust)
#looks like the naming convention is not consistent...
apaPASres_topPos_pval_normAPA=apaPASres_topPos_pval %>% inner_join(apaN,by=c("PAS","gene")) %>% dplyr::select(-PAS,-p.adjust)
apaPASres_topNeg_pval_normAPA=apaPASres_topNeg_pval %>% inner_join(apaN,by=c("PAS","gene"))%>% dplyr::select(-PAS,-p.adjust)


#create a dataframe with expression and apa
APA_posandExp=Exp %>% inner_join(apaPASres_topPos_pval_normAPA, by="gene")
APA_negandExp=Exp %>% inner_join(apaPASres_topNeg_pval_normAPA,by="gene")
```

###First do the positive:  


Seperate De and not De  

```{r}
is_de <- which(APA_posandExp$adj.P.Val < .05)
isnot_de <- which(APA_posandExp$adj.P.Val >= .05)

gvec <- APA_posandExp$genes
Exp_pos <- APA_posandExp %>% select(-gene,-adj.P.Val, -contains("APA"))
APA_pos <- APA_posandExp %>% select(contains("APA"))

# metadata label
species <- factor(c("H","H","H","H","H","H","C","C","C","C","C","C"))
batch <- factor(c("A", "A", "B", "A", "B", "B", "B", "B","A", "A","A","B"))

metadata <- data.frame(sample=names(Exp_pos),
                       species=species,
                       batch=batch)
```

Compute indirect effects:  

```{r}
fit_de_pos <- test_mediation(exprs = Exp_pos[is_de,], 
                         fixed_covariates = list(species=metadata$species,
                                                 batch=metadata$batch),
                         varying_covariate = APA_pos[is_de,])

fit_node_pos <- test_mediation(exprs = Exp_pos[isnot_de,], 
                           fixed_covariates = list(species=metadata$species,
                                                   batch=metadata$batch),
                           varying_covariate = APA_pos[isnot_de,])
```


```{r}
DEdat <- data.frame(bf=fit_de_pos$tau, af=fit_de_pos$tau_prime)
#DEdat$color <- ifelse(DEdat$significance==TRUE, "red", "black")
plot(x=DEdat$bf, y=DEdat$af, ylab="Effect Size After Controlling for APA", xlab="Effect Size Before Controlling for APA", main="Effect of APA on Expression Divergence in DE genes", pch=16, cex=0.6, xlim=c(-10, 10), ylim=c(-10,10))

abline(0, 1)
abline(h=0)
abline(v=0)
```
Not DE:

```{r}
#Visualization for non DE genes:
noDEdat <- data.frame(bf=fit_node_pos$tau, af=fit_node_pos$tau_prime)
#noDEdat$color <- ifelse(noDEdat$significance==TRUE, "red", "black")
plot(x=noDEdat$bf, y=noDEdat$af, ylab="Effect Size After Controlling for APA", xlab="Effect Size Before Controlling for APA", main="Effect of APA on Expression Divergence in non-DE genes", pch=16, cex=0.6, xlim=c(-10, 10), ylim=c(-10,10), adj=0.6)
abline(0, 1)
abline(h=0)
abline(v=0)
```



###Negative:  
Seperate De and not De  

```{r}
is_de_neg <- which(APA_negandExp$adj.P.Val < .05)
isnot_de_gen <- which(APA_negandExp$adj.P.Val >= .05)

gvec_neg <- APA_negandExp$genes
Exp_neg <- APA_negandExp %>% select(-gene,-adj.P.Val, -contains("APA"))
APA_neg <- APA_negandExp %>% select(contains("APA"))


```

Compute indirect effects:  

```{r}
fit_de_neg<- test_mediation(exprs = Exp_neg[is_de_neg,], 
                         fixed_covariates = list(species=metadata$species,
                                                 batch=metadata$batch),
                         varying_covariate = APA_neg[is_de_neg,])

fit_node_neg <- test_mediation(exprs = Exp_neg[isnot_de_gen,], 
                           fixed_covariates = list(species=metadata$species,
                                                   batch=metadata$batch),
                           varying_covariate = APA_neg[isnot_de_gen,])
```


```{r}
DEdat_neg <- data.frame(bf=fit_de_neg$tau, af=fit_de_neg$tau_prime)
#DEdat$color <- ifelse(DEdat$significance==TRUE, "red", "black")
plot(x=DEdat_neg$bf, y=DEdat_neg$af, ylab="Effect Size After Controlling for APA", xlab="Effect Size Before Controlling for APA", main="Effect of APA on Expression Divergence in DE genes", pch=16, cex=0.6, xlim=c(-10, 10), ylim=c(-10,10))

abline(0, 1)
abline(h=0)
abline(v=0)
```

Not DE:

```{r}
#Visualization for non DE genes:
noDEdat_neg <- data.frame(bf=fit_node_neg$tau, af=fit_node_neg$tau_prime)
#noDEdat$color <- ifelse(noDEdat$significance==TRUE, "red", "black")
plot(x=noDEdat_neg$bf, y=noDEdat_neg$af, ylab="Effect Size After Controlling for APA", xlab="Effect Size Before Controlling for APA", main="Effect of APA on Expression Divergence in non-DE genes", pch=16, cex=0.6, xlim=c(-10, 10), ylim=c(-10,10), adj=0.6)
abline(0, 1)
abline(h=0)
abline(v=0)
```
