---
title: "Test Methods to Merge Within Species"
author: "Briana Mittleman"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(scales)
library(workflowr)
library(tidyverse)
library(cowplot)
```
##Exploration  
I noticed there are overlapping PAS within species. I deal with overlapping PAS between species by merging the PAS into one when 125 of the bases overlap. 

For the within species I will look at a merge on the PAS that pass [liftover](liftoverPAS.html). 

../data/cleanPeaks_lifted/Chimp_PASregions_humanCoord.sort.bed
../data/cleanPeaks_lifted/Chimp_PASregions.bed
../data/cleanPeaks_lifted/Human_PASregions.sort.bed

I will test the human coordinates and the chimp coordinates for the chimp pas. 

I will use bedtools merge 

```{bash, eval=F}
mkdir ../data/TestWithinMergePAS
bedtools merge -s -c 4,4 -o distinct,count -i ../data/cleanPeaks_lifted/Chimp_PASregions_humanCoord.sort.bed > ../data/TestWithinMergePAS/Chimp_PASregions_humanCoord.merge.bed

sort -k1,1 -k2,2n  ../data/cleanPeaks_lifted/Chimp_PASregions.bed >  ../data/cleanPeaks_lifted/Chimp_PASregions.sort.bed

bedtools merge -s -c 4,4 -o distinct,count  -i ../data/cleanPeaks_lifted/Chimp_PASregions.sort.bed > ../data/TestWithinMergePAS/Chimp_PASregions_chimpCoord.merge.bed

bedtools merge -s -c 4,4 -o distinct,count  -i ../data/cleanPeaks_lifted/Human_PASregions.sort.bed > ../data/TestWithinMergePAS/Human_PASregions.merge.bed


```


###Chimp  
Compare 2 chimp conditions:  

```{r}
ChimpHumanCoord=read.table("../data/TestWithinMergePAS/Chimp_PASregions_humanCoord.merge.bed",col.names = c("chr",'start','end','PAS', 'nPAS'),stringsAsFactors = F)


ChimpPAS=read.table("../data/TestWithinMergePAS/Chimp_PASregions_chimpCoord.merge.bed",col.names = c("chr",'start','end','PAS', 'nPAS'),stringsAsFactors = F)
```


```{r}
ChimpHumanCoord %>% group_by(nPAS) %>% summarise(n())
ggplot(ChimpHumanCoord,aes(x=nPAS)) + geom_bar(stat="count")
```
```{r}
ChimpPAS %>% group_by(nPAS) %>% summarise(n())
ggplot(ChimpPAS,aes(x=nPAS)) + geom_bar(stat="count")
```
Numbers are the same. Hopfully they are the same overlaps.  

```{r}
length(intersect(ChimpHumanCoord$PAS,ChimpPAS$PAS))
```
Ok these are the same. I can do it all in human coordinate.  

Explore the size of these bins.  

```{r}
ChimpHumanCoord_details= ChimpHumanCoord %>% mutate(size=end-start)
ChimpHumanCoord_details$nPAS=as.factor(ChimpHumanCoord_details$nPAS)
```


```{r}
hue_pal()(7)

```

```{r}
chimpsize=ggplot(ChimpHumanCoord_details, aes(x=nPAS, y=size,fill=nPAS))+ geom_boxplot() + geom_hline(yintercept = 200,col="#F8766D")+  geom_hline(yintercept = 400,col="#C49A00") + geom_hline(yintercept = 600,col="#53B400") + geom_hline(yintercept = 800,col="#00C094") + geom_hline(yintercept = 1000,col="#00B6EB") + geom_hline(yintercept = 1200,col="#A58AFF")+ geom_hline(yintercept = 1400,col="#FB61D7")+ labs(title="Size of chimp overlap PAS, colored by expected if overlap = 1 base")
```

Look at some of the top examples in the data. Is there evidence that high level groups should be filtered out?  I can also decide whether I take the middle 200 or use the full region.  

```{r}
ChimpHumanCoord_details %>% arrange(desc(nPAS)) %>% head()
```
This is before I do any combination with the human PAS. 


###Human  

```{r}
HumanPAS=read.table("../data/TestWithinMergePAS/Human_PASregions.merge.bed",col.names = c("chr",'start','end','PAS', 'nPAS'),stringsAsFactors = F)
```


```{r}
HumanPAS %>% group_by(nPAS) %>% summarise(n())
ggplot(HumanPAS,aes(x=nPAS)) + geom_bar(stat="count")
```

There are more overlaps in the human set.  

```{r}

HumanPAS_details= HumanPAS %>% mutate(size=end-start)
HumanPAS_details$nPAS=as.factor(HumanPAS_details$nPAS)


humansize=ggplot(HumanPAS_details, aes(x=nPAS, y=size,fill=nPAS))+ geom_boxplot() + geom_hline(yintercept = 200,col="#F8766D")+  geom_hline(yintercept = 400,col="#C49A00") + geom_hline(yintercept = 600,col="#53B400") + geom_hline(yintercept = 800,col="#00C094") + geom_hline(yintercept = 1000,col="#00B6EB") + geom_hline(yintercept = 1200,col="#A58AFF")+ geom_hline(yintercept = 1400,col="#FB61D7")+ labs(title="Size of human overlap PAS, colored by expected if overlap = 1 base")
```


```{r}
plot_grid(chimpsize,humansize)
```
This shows that the human overlaps at the higher numbers are closer together.  
```{r}
HumanPAS_details %>% arrange(desc(nPAS)) %>% head()

```

###Passing filters 
Focus on examples that make it in the final set.  

```{r}
overlapres=read.table("../data/OverlappingPAS/MergedPAS_HumanCords.txt", col.names = c("chr", "start", "end", "name", "numb")) %>% mutate(length=end-start) 

overlapres %>% arrange(desc(numb)) %>% head()
```

After looking at examples, it does not seem like merging is the right way to handle this. It would lead to inflatted numbers in regions and I do not want that. I will leave this for now.  
