---
title: "Multmap Experiment"
author: "Briana Mittleman"
date: "3/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

standard featureCounts (no multimapped reads)
featureCounts with -M flag (includes all alignments of multimapped reads)
featureCounts with --primary flag (includes multimapped reads, but only their primary alignment)
Stringent samtools quality filter followed by featureCounts with -M flag
Permissive samtools quality filter followed by featureCounts with -M flag  

```{r}
library(workflowr)
library(tidyverse)
```

```{bash,eval=F}
mkdir ../data/PAS_SAF
mkdir ../data/Test_FC_methods
```


PAS  
```{r}
Meta=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt",header = T)
ChimpMeta=read.table("../data/PAS_doubleFilter/PAS_doublefilter_either_ChimpCoordChimpUsage.sort.bed", col.names = c("chr","start","end", "PAS", "score", "strand"),stringsAsFactors = F)

#GeneID\tChr\tStart\tEnd\tStrand\n"
#peak_chr10_53894_53985_+	10	53894	53985	+

HumanSAF= Meta %>% rename("GeneID"=PAS, "Chr"=chr, "Start"=start, "End"=end,"Strand"=strandFix) %>% select(GeneID, Chr, Start, End, Strand)
write.table(HumanSAF,"../data/PAS_SAF/Human_PASdoublefilter.SAF", col.names = T,row.names = F, quote = F, sep="\t" )

ChimpSAF= ChimpMeta%>% rename("GeneID"=PAS, "Chr"=chr, "Start"=start, "End"=end,"Strand"=strand) %>% select(GeneID, Chr, Start, End, Strand)
write.table(ChimpSAF,"../data/PAS_SAF/Chimp_PASdoublefilter.SAF", col.names = T,row.names = F, quote = F, sep="\t" )

```


Fearture counts 5 ways:  


standard featureCounts (no multimapped reads)
featureCounts with -M flag (includes all alignments of multimapped reads)
featureCounts with --primary flag (includes multimapped reads, but only their primary alignment)
Stringent samtools quality filter followed by featureCounts with -M flag (-M, -Q 5)
medium samtools quality filter followed by featureCounts with -M flag (-M, -Q 15)
Permissive samtools quality filter followed by featureCounts with -M flag (-M, -Q 20)


```{bash,eval=F}
sbatch TestFC.sh
```
 

