---
title: "Dominant PAS and DE"
author: "Briana Mittleman"
date: "4/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(workflowr)
library(tidyverse)
library(cowplot)
library(RColorBrewer)
```

I will ask if there if dominance and DE are related. First I can ask if genes with dominant PAS are enriched in the DE genes.  

```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)
MetaCol=colnames(PAS)

nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F)
DE= read.table("../data/DiffExpression/DEtested_allres.txt",header=F, stringsAsFactors = F,col.names = c('Gene_stable_ID', 'logFC' ,'AveExpr', 't', 'P.Value', 'adj.P.Val', 'B')) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(-Gene_stable_ID, -Source_of_gene_name) %>% rename("gene"=Gene.name) %>% mutate(DE=ifelse(adj.P.Val<=.05, "Yes","No")) %>% select(gene,DE)
DE_yes= DE %>% filter(DE=="Yes")
```

Domiance 

```{r}
#9
HumanDom9=read.table("../data/DomDefGreaterX/Human_.9_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human9") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom9=read.table("../data/DomDefGreaterX/Chimp_.9_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp9") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#8 
HumanDom8=read.table("../data/DomDefGreaterX/Human_.8_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human8")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom8=read.table("../data/DomDefGreaterX/Chimp_.8_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp8") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#7 
HumanDom7=read.table("../data/DomDefGreaterX/Human_.7_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human7")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom7=read.table("../data/DomDefGreaterX/Chimp_.7_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp7")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))


#6 
HumanDom6=read.table("../data/DomDefGreaterX/Human_.6_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human6") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom6=read.table("../data/DomDefGreaterX/Chimp_.6_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp6") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#5  
HumanDom5=read.table("../data/DomDefGreaterX/Human_.5_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human5") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom5=read.table("../data/DomDefGreaterX/Chimp_.5_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp5")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#4 
HumanDom4=read.table("../data/DomDefGreaterX/Human_.4_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human4")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom4=read.table("../data/DomDefGreaterX/Chimp_.4_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp4")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#3 
HumanDom3=read.table("../data/DomDefGreaterX/Human_.3_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human3") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom3=read.table("../data/DomDefGreaterX/Chimp_.3_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp3")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#2
HumanDom2=read.table("../data/DomDefGreaterX/Human_.2_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human2")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom2=read.table("../data/DomDefGreaterX/Chimp_.2_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp2")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#1
HumanDom1=read.table("../data/DomDefGreaterX/Human_.1_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Human1")%>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))
ChimpDom1=read.table("../data/DomDefGreaterX/Chimp_.1_dominantPAS.txt", col.names = MetaCol,stringsAsFactors = F) %>% mutate(set="Chimp1") %>% mutate(DE=ifelse(gene %in% DE_yes$gene,"Yes","No"))

#all  
HumanDomAll= HumanDom1 %>% bind_rows(HumanDom2) %>% bind_rows(HumanDom3) %>% bind_rows(HumanDom4) %>% bind_rows(HumanDom5) %>% bind_rows(HumanDom6) %>% bind_rows(HumanDom7) %>% bind_rows(HumanDom8) %>% bind_rows(HumanDom9) 
ChimpDomAll= ChimpDom1 %>% bind_rows(ChimpDom2) %>% bind_rows(ChimpDom3) %>% bind_rows(ChimpDom4) %>% bind_rows(ChimpDom5) %>% bind_rows(ChimpDom6) %>% bind_rows(ChimpDom7) %>% bind_rows(ChimpDom8) %>% bind_rows(ChimpDom9) 

```



```{r}


ChimpSet=c('Chimp1','Chimp2', 'Chimp3', 'Chimp4', 'Chimp5', 'Chimp6', 'Chimp7', 'Chimp8','Chimp9')
EnrichChimp=c()
PvalueChimp=c()
for (i in ChimpSet){
  x=nrow(ChimpDomAll %>% filter(set==i, DE=="Yes"))
  m=nrow(DE_yes)
  n=nrow(DE) - nrow(DE_yes)
  k=nrow(ChimpDomAll %>% filter(set==i))
  N=nrow(DE)
  PvalueChimp=c(PvalueChimp, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichChimp=c(EnrichChimp, enrich)
}

PvalueChimp
EnrichChimp
```


```{r}
HumanSet=c('Human1','Human2', 'Human3', 'Human4', 'Human5', 'Human6', 'Human7', 'Human8','Human9')
EnrichHuman=c()
PvalueHuman=c()
for (i in HumanSet){
  x=nrow(HumanDomAll %>% filter(set==i, DE=="Yes"))
  m=nrow(DE_yes)
  n=nrow(DE) - nrow(DE_yes)
  k=nrow(HumanDomAll %>% filter(set==i))
  N=nrow(DE)
  PvalueHuman=c(PvalueHuman, phyper(x,m,n,k,lower.tail=F))
  enrich=(x/k)/(m/N)
  EnrichHuman=c(EnrichHuman, enrich)
}

PvalueHuman
EnrichHuman
```


No enrichment for these. The real question is if genes with different dominant PAS are DE. This requires chosing how to call different dominant. 

##Same vs different dominant:  

Are genes with different dominant at the cutoff .4 cutoff enriched for DE:  

```{r}
FourRes=read.table("../data/DomStructure_4/InclusiveDominantPASat4.txt", header = T,stringsAsFactors = F)
FourRes_diff= FourRes %>% filter(Set=="Different")
FourRes_same= FourRes %>% filter(Set=="Same")
```

```{r}
x=length(intersect(FourRes_diff$gene,DE_yes$gene))
m=nrow(DE_yes)
n=nrow(DE) - nrow(DE_yes)
k=nrow(FourRes %>% filter(Set=="Different"))
N=nrow(DE)
phyper(x,m,n,k,lower.tail=F)
(x/k)/(m/N)

```
```{r}
x=length(intersect(FourRes_same$gene,DE_yes$gene))
m=nrow(DE_yes)
n=nrow(DE) - nrow(DE_yes)
k=nrow(FourRes %>% filter(Set=="Same"))
N=nrow(DE)
phyper(x,m,n,k,lower.tail=F)
(x/k)/(m/N)

```


This is conditioned on the gene having a dominant PAS.  

do this based on a set tested both  

```{r}
All4= FourRes %>% select(gene,Set) %>% inner_join(DE, by="gene")

x=nrow(All4 %>% filter(Set=="Different", DE=="Yes"))
m=nrow(All4 %>% filter( DE=="Yes"))
n=nrow(All4 %>% filter( DE=="No"))
k=nrow(All4 %>% filter(Set=="Different"))
N=nrow(All4)
phyper(x,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

I am not sure what the set should be.  

```{r}
x=nrow(All4 %>% filter(Set=="Same", DE=="Yes"))
m=nrow(All4 %>% filter(DE=="Yes"))
n=nrow(All4 %>% filter(DE=="No"))
k=nrow(All4 %>% filter(Set=="Same"))
N=nrow(All4)
phyper(x,m,n,k,lower.tail=F)
(x/k)/(m/N)
```

##Robust to different cutoffs  

```{r}
HumanRes=read.table("../data/DomDefGreaterX/Human_AllGenes_DiffTop.txt", col.names = c("Human_PAS", "gene","Human_DiffDom"),stringsAsFactors = F)

ChimpRes=read.table("../data/DomDefGreaterX/Chimp_AllGenes_DiffTop.txt", col.names = c("Chimp_PAS", "gene","Chimp_DiffDom"),stringsAsFactors = F)

BothRes=HumanRes %>% inner_join(ChimpRes,by="gene")

BothRes_10=BothRes %>% filter(Chimp_DiffDom >=0.1 | Human_DiffDom>=0.1) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=10) 
BothRes_20=BothRes %>% filter(Chimp_DiffDom >=0.2 | Human_DiffDom>=0.2) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=20)
BothRes_30=BothRes %>% filter(Chimp_DiffDom >=0.3 | Human_DiffDom>=0.3) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=30)
BothRes_40=BothRes %>% filter(Chimp_DiffDom >=0.4 | Human_DiffDom>=0.4) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=40)
BothRes_50=BothRes %>% filter(Chimp_DiffDom >=0.5 | Human_DiffDom>=0.5) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=50)
BothRes_60=BothRes %>% filter(Chimp_DiffDom >=0.6 | Human_DiffDom>=0.6) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=60)
BothRes_70=BothRes %>% filter(Chimp_DiffDom >=0.7 | Human_DiffDom>=0.7) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=70)
BothRes_80=BothRes %>% filter(Chimp_DiffDom >=0.8 | Human_DiffDom>=0.8) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=80)
BothRes_90=BothRes %>% filter(Chimp_DiffDom >=0.9 | Human_DiffDom>=0.9) %>% mutate(Set= ifelse(Human_PAS==Chimp_PAS,"Same", "Different"),cut=90)

BothResAll=BothRes_10 %>% bind_rows(BothRes_20) %>% bind_rows(BothRes_30) %>% bind_rows(BothRes_40) %>% bind_rows(BothRes_50) %>% bind_rows(BothRes_60) %>% bind_rows(BothRes_70) %>% bind_rows(BothRes_80) %>% bind_rows(BothRes_90)


```

```{r}
Pval=c()
Enrich=c()
set=c(10,20,30,40,50,60,70,80,90)


All4= BothResAll  %>% select(gene,cut,Set) %>% inner_join(DE, by="gene")


for (i in set){
  x=nrow(All4 %>% filter(cut==i, Set=="Different", DE=="Yes"))
  m=nrow(All4 %>% filter(cut==i, DE=="Yes"))
  n=nrow(All4 %>% filter(cut==i, DE=="No"))
  k=nrow(All4 %>% filter(cut==i, Set=="Different"))
  N=nrow(All4 %>% filter(cut==i))
  val=phyper(x,m,n,k,lower.tail=F)
  Pval= c(Pval, val)
  en=(x/k)/(m/N)
  Enrich=c(Enrich, en)
}

ResDF=as.data.frame(cbind(set,Pval,Enrich))
ResDF$set=as.factor(ResDF$set)
ResDF$Pval=as.numeric(as.character(ResDF$Pval))
ResDF$Enrich=as.numeric(as.character(ResDF$Enrich))

ggplot(ResDF,aes(x=set, y=-log10(Pval))) + geom_bar(stat="identity") +labs(title="Enrichment pvalues for DE and different dominant \n condition on tested in both")
ggplot(ResDF,aes(x=set, y=Enrich)) + geom_bar(stat="identity") + geom_hline(yintercept = 1)+labs(title="Enrichment for DE and different dominant \n condition on tested in both")
```

```{r}
PvalSame=c()
EnrichSame=c()




for (i in set){
  x=nrow(All4 %>% filter(cut==i, Set=="Same", DE=="Yes"))
  m=nrow(All4 %>% filter(cut==i, DE=="Yes"))
  n=nrow(All4 %>% filter(cut==i, DE=="No"))
  k=nrow(All4 %>% filter(cut==i, Set=="Same"))
  N=nrow(All4 %>% filter(cut==i))
  val=phyper(x,m,n,k,lower.tail=F)
  PvalSame= c(PvalSame, val)
  en=(x/k)/(m/N)
  EnrichSame=c(EnrichSame, en)
}

ResDFSame=as.data.frame(cbind(set,PvalSame,EnrichSame))
ResDFSame$set=as.factor(ResDFSame$set)
ResDFSame$PvalSame=as.numeric(as.character(ResDFSame$PvalSame))
ResDFSame$EnrichSame=as.numeric(as.character(ResDFSame$EnrichSame))

ggplot(ResDFSame,aes(x=set, y=-log10(PvalSame))) + geom_bar(stat="identity") +labs(title="Enrichment pvalues for DE and same dominant \n condition on tested in both")
ggplot(ResDFSame,aes(x=set, y=EnrichSame)) + geom_bar(stat="identity") + geom_hline(yintercept = 1)+labs(title="Enrichment for DE and same dominant \n condition on tested in both")
```
