---
title: "Normalize Usage Phenotype"
author: "Briana Mittleman"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to normalize the phenotypes with the leafcutter scripts. This can be used to perform a PCA and assess the data quality. I will include, total nuclear human and chimp.  

```{r}
library(tidyverse)
library(reshape2)
library(ggpubr)
```

These are the inclusive phenotypes. I will need to subset of the 5% pas.  
../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt
../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt


The 5% pas are in ../data/Peaks_5perc/Peaks_5perc_either_bothUsage_noUnchr.txt

I will make a python script that will do this. I


```{bash,eval=F}
python filter5percPAS.py ../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt  ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Human_Pheno_5perc.txt

python filter5percPAS.py ../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt  ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Chimp_Pheno_5perc.txt
```


Join these to normalize the phenotypes together:  

```{r}
humanPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Human_Pheno_5perc.txt",stringsAsFactors = F, header = T)
chimpPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Chimp_Pheno_5perc.txt",stringsAsFactors = F, header = T)


allPheno=humanPheno %>% full_join(chimpPheno,by="chrom")


write.table(allPheno, "../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt", col.names = T, row.names = F, quote = F)
```










```{bash,eval=F}
gzip ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt

#conda deactivate 
conda deactivate 
conda deactivate 
#python 2
source ~/activate_anaconda_python2.sh 
#go to directory ../data/Pheno_5perc/
python ../../code/prepare_phenotype_table.py ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz
cat ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_chr* > ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_AllChrom

```

Use these normalized phenotypes for the PCA 

```{r}
metaData=read.table("../data/metadata_HCpanel.txt", header = T, stringsAsFactors = F)
```


```{r}
normPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_AllChrom", col.names = c('Chr',	'start',	'end',	'ID',	'18498_N',	'18498_T',	'18499_N',	'18499_T',	'18502_N',	'18502_T',	'18504_N',	'18504_T',	'18510_N',	'18510_T',	'18523_N',	'18523_T',	'18358_N',	'18358_T',	'3622_N',	'3622_T',	'3659_N',	'3659_T',	'4973_N',	'4973_T',	'pt30_N',	'pt30_T',	'pt91_N',	'pt91_T'))

normPheno_matrix=as.matrix(normPheno %>% select(-Chr, -start, -end, -ID))
```
Run PCA:  

```{r}
pca_Pheno=prcomp(normPheno_matrix, center=T,scale=T)
pca_df=as.data.frame(pca_Pheno$rotation) %>% rownames_to_column(var="ID")

```

```{r}
eigs <- pca_Pheno$sdev^2
proportion = eigs/sum(eigs)

plot(proportion)
```


Plot color as species and shape as total v nuclear 

```{r}
top5PC=pca_df %>% select(ID, PC1, PC2, PC3, PC4, PC5) %>%  inner_join(metaData, by="ID")
top5PC$Cycles=as.factor(top5PC$Cycles)

ggplot(top5PC,aes(x=PC1, y=PC2, col=Species, shape=Fraction)) + geom_point(size=3)
```



```{r}
ggplot(top5PC,aes(x=PC1, y=PC2, shape=Species,col=Cycles)) + geom_point(size=3) + geom_text(aes(label=Line), position = position_nudge(y = 0.03) )
```




PC1
```{r}
Fraction1=summary(lm(top5PC$PC1 ~ top5PC$Fraction))$adj.r.squared
Cycles1=summary(lm(top5PC$PC1 ~ top5PC$Cycles))$adj.r.squared
Species1=summary(lm(top5PC$PC1 ~ top5PC$Species))$adj.r.squared
Concentration1=summary(lm(top5PC$PC1 ~ top5PC$Concentration))$adj.r.squared
Library_concentration1=summary(lm(top5PC$PC1 ~ top5PC$Library_concentration))$adj.r.squared
PerMapClean1=summary(lm(top5PC$PC1 ~ top5PC$PerMapClean))$adj.r.squared
Ratio1=summary(lm(top5PC$PC1 ~ top5PC$Ratio))$adj.r.squared



PC1Fac=c(Fraction1, Species1, Concentration1, Ratio1,Cycles1,Library_concentration1, PerMapClean1)
```

PC2
```{r}
Fraction2=summary(lm(top5PC$PC2 ~ top5PC$Fraction))$adj.r.squared
Cycles2=summary(lm(top5PC$PC2 ~ top5PC$Cycles))$adj.r.squared
Species2=summary(lm(top5PC$PC2 ~ top5PC$Species))$adj.r.squared
Concentration2=summary(lm(top5PC$PC2 ~ top5PC$Concentration))$adj.r.squared
Library_concentration2=summary(lm(top5PC$PC2 ~ top5PC$Library_concentration))$adj.r.squared
PerMapClean2=summary(lm(top5PC$PC2 ~ top5PC$PerMapClean))$adj.r.squared
Ratio2=summary(lm(top5PC$PC2 ~ top5PC$Ratio))$adj.r.squared



PC2Fac=c(Fraction2, Species2, Concentration2, Ratio2,Cycles2,Library_concentration2, PerMapClean2)
```

PC3
```{r}
Fraction3=summary(lm(top5PC$PC3 ~ top5PC$Fraction))$adj.r.squared
Cycles3=summary(lm(top5PC$PC3 ~ top5PC$Cycles))$adj.r.squared
Species3=summary(lm(top5PC$PC3 ~ top5PC$Species))$adj.r.squared
Concentration3=summary(lm(top5PC$PC3 ~ top5PC$Concentration))$adj.r.squared
Library_concentration3=summary(lm(top5PC$PC3 ~ top5PC$Library_concentration))$adj.r.squared
PerMapClean3=summary(lm(top5PC$PC3 ~ top5PC$PerMapClean))$adj.r.squared
Ratio3=summary(lm(top5PC$PC3 ~ top5PC$Ratio))$adj.r.squared



PC3Fac=c(Fraction3, Species3, Concentration3, Ratio3,Cycles3,Library_concentration3, PerMapClean3)
```


PC4
```{r}
Fraction4=summary(lm(top5PC$PC4 ~ top5PC$Fraction))$adj.r.squared
Cycles4=summary(lm(top5PC$PC4 ~ top5PC$Cycles))$adj.r.squared
Species4=summary(lm(top5PC$PC4 ~ top5PC$Species))$adj.r.squared
Concentration4=summary(lm(top5PC$PC4 ~ top5PC$Concentration))$adj.r.squared
Library_concentration4=summary(lm(top5PC$PC4 ~ top5PC$Library_concentration))$adj.r.squared
PerMapClean4=summary(lm(top5PC$PC4 ~ top5PC$PerMapClean))$adj.r.squared
Ratio4=summary(lm(top5PC$PC4 ~ top5PC$Ratio))$adj.r.squared



PC4Fac=c(Fraction4,Species4, Concentration4, Ratio4,Cycles4,Library_concentration4, PerMapClean4)
```

PC5
```{r}
Fraction5=summary(lm(top5PC$PC5 ~ top5PC$Fraction))$adj.r.squared
Cycles5=summary(lm(top5PC$PC5 ~ top5PC$Cycles))$adj.r.squared
Species5=summary(lm(top5PC$PC5 ~ top5PC$Species))$adj.r.squared
Concentration5=summary(lm(top5PC$PC5 ~ top5PC$Concentration))$adj.r.squared
Library_concentration5=summary(lm(top5PC$PC5 ~ top5PC$Library_concentration))$adj.r.squared
PerMapClean5=summary(lm(top5PC$PC5 ~ top5PC$PerMapClean))$adj.r.squared
Ratio5=summary(lm(top5PC$PC5 ~ top5PC$Ratio))$adj.r.squared



PC5Fac=c(Fraction5,Species5, Concentration5, Ratio5,Cycles5,Library_concentration5, PerMapClean5)
```

Make DF and plot: 

```{r}
Exp=c("Fraction",'Species', 'Concentration', 'Ratio', 'Cycles','Library_concentration', 'PerMapClean')

pcandEx=as.data.frame(cbind(Experiment=Exp,pc1=PC1Fac, pc2=PC2Fac,  pc3=PC3Fac,  pc4=PC4Fac,  pc5=PC5Fac))

pcandExM=melt(pcandEx, id.var="Experiment",variable.name = "PC", value.name = "PVE")
pcandExM$PVE=as.numeric(pcandExM$PVE)

```

```{r}
ggplot(pcandExM, aes(x=PC, fill=PVE, y=Experiment))+ geom_tile() + labs(title="Proportion of variation explained in PCs by Experimental Faction", y="")+ scale_fill_distiller(palette = "Blues", direction=1)
```

