---
title: "Normalize Usage Phenotype"
author: "Briana Mittleman"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I want to normalize the phenotypes with the leafcutter scripts. This can be used to perform a PCA and assess the data quality. I will include, total nuclear human and chimp.  

```{r}
library(tidyverse)
library(reshape2)
library(ggpubr)
library("scales")
library("gplots")
library("RColorBrewer")
```

These are the inclusive phenotypes. I will need to subset of the 5% pas.  
../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt
../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt


The 5% pas are in ../data/Peaks_5perc/Peaks_5perc_either_bothUsage_noUnchr.txt

I will make a python script that will do this. I


```{bash,eval=F}
python filter5percPAS.py ../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt  ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Human_Pheno_5perc.txt

python filter5percPAS.py ../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt  ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Chimp_Pheno_5perc.txt
```


Join these to normalize the phenotypes together:  

```{r}
humanPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Human_Pheno_5perc.txt",stringsAsFactors = F, header = T)
chimpPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_Chimp_Pheno_5perc.txt",stringsAsFactors = F, header = T)


allPheno=humanPheno %>% full_join(chimpPheno,by="chrom")


write.table(allPheno, "../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt", col.names = T, row.names = F, quote = F)
```










```{bash,eval=F}
gzip ../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt

#conda deactivate 
conda deactivate 
conda deactivate 
#python 2
source ~/activate_anaconda_python2.sh 
#go to directory ../data/Pheno_5perc/
python ../../code/prepare_phenotype_table.py ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz

cat ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_chr* > ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_AllChrom

```

Use these normalized phenotypes for the PCA 

```{r}
metaData=read.table("../data/metadata_HCpanel.txt", header = T, stringsAsFactors = F)
```


```{r}
normPheno=read.table("../data/Pheno_5perc/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc.txt.gz.phen_AllChrom", col.names = c('Chr',	'start',	'end',	'ID',	'18498_N',	'18498_T',	'18499_N',	'18499_T',	'18502_N',	'18502_T',	'18504_N',	'18504_T',	'18510_N',	'18510_T',	'18523_N',	'18523_T',	'18358_N',	'18358_T',	'3622_N',	'3622_T',	'3659_N',	'3659_T',	'4973_N',	'4973_T',	'pt30_N',	'pt30_T',	'pt91_N',	'pt91_T'))

normPheno_matrix=as.matrix(normPheno %>% dplyr::select(-Chr, -start, -end, -ID))
```
Run PCA:
```{r}
# Load colors 

colors <- colorRampPalette(c(brewer.pal(9, "Blues")[1],brewer.pal(9, "Blues")[9]))(100)

pal <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), brewer.pal(12, "Set3"))
labels <- paste(metaData$Species,metaData$Line,metaData$Fraction, sep=" ")
```

```{r}

cors <- cor(normPheno_matrix, method="spearman", use="pairwise.complete.obs")


heatmap.2( cors, scale="none", col = colors, margins = c(12, 12), trace='none', denscol="white", labCol=labels, ColSideColors=pal[as.integer(as.factor(metaData$Species))], RowSideColors=pal[as.integer(as.factor(metaData$Fraction))+9], cexCol = 0.2 + 1/log10(15), cexRow = 0.2 + 1/log10(15))
```

```{r}
pca_Pheno=prcomp(t(normPheno_matrix), scale=F)
scores = pca_Pheno$x

```
Scores code
```{r}
#PCA function (original code from Julien Roux)
#Load in the plot_scores function
plot_scores <- function(pca, scores, n, m, cols, points=F, pchs =20, legend=T){
  xmin <- min(scores[,n]) - (max(scores[,n]) - min(scores[,n]))*0.05
  if (legend == T){ ## let some room (35%) for a legend                                                                                                                                                 
    xmax <- max(scores[,n]) + (max(scores[,n]) - min(scores[,n]))*0.50
  }
  else {
    xmax <- max(scores[,n]) + (max(scores[,n]) - min(scores[,n]))*0.05
  }
  ymin <- min(scores[,m]) - (max(scores[,m]) - min(scores[,m]))*0.05
  ymax <- max(scores[,m]) + (max(scores[,m]) - min(scores[,m]))*0.05
  plot(scores[,n], scores[,m], xlab=paste("PC", n, ": ", round(summary(pca)$importance[2,n],3)*100, "% variance explained", sep=""), ylab=paste("PC", m, ": ", round(summary(pca)$importance[2,m],3)*100, "% variance explained", sep=""), xlim=c(xmin, xmax), ylim=c(ymin, ymax), type="n")
  if (points == F){
    text(scores[,n],scores[,m], rownames(scores), col=cols, cex=1)
  }
  else {
    points(scores[,n],scores[,m], col=cols, pch=pchs, cex=1.3)
  }
}
```


```{r}
metaData$Species=as.factor(metaData$Species)
for (n in 1:1){
  col.v <- pal[as.integer(metaData$Species)]
  plot_scores(pca_Pheno, scores, n, n+1, col.v)
}
```

```{r}
eigs <- pca_Pheno$sdev^2
proportion = eigs/sum(eigs)

plot(proportion)
```


```{r}
x.pca <- pca_Pheno

tech_factors <- metaData
tech_factors_sum <- tech_factors[,c(2:15)] %>% dplyr::select(-Library,-Line)

p_comps <- 1:6
pc_cov_cor <- matrix(nrow = ncol(tech_factors_sum), ncol = length(p_comps),
                     dimnames = list(colnames(tech_factors_sum), colnames(x.pca$x)[p_comps]))
for (pc in p_comps) {
  for (covariate in 1:ncol(tech_factors_sum)) {
    lm_result <- lm(x.pca$x[, pc] ~ tech_factors_sum[, covariate])
    r2 <- summary(lm_result)$r.squared
    pc_cov_cor[covariate, pc] <- r2
  }
}

pc_cov_pval <- matrix(nrow = ncol(tech_factors_sum), ncol = length(p_comps),
                      dimnames = list(colnames(tech_factors_sum), colnames(x.pca$x)[p_comps]))

for (pc in p_comps) {
  for (covariate_2 in 1:ncol(tech_factors_sum)) {
    lm_result_2 <- lm(x.pca$x[, pc] ~ tech_factors_sum[, covariate_2])
    pval <- anova(lm_result_2)$'Pr(>F)'[1]
    pc_cov_pval[covariate_2, pc] <- pval
  }
}

PCs <- c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6")
Tech_fac <- colnames(tech_factors_sum)
#Tech_fac <- c("Species",   "Individual", "O2.",  "Condition" , "Sex", "RIN" , "CO2", "Purity_high", "Purity_med" ,
              #"Expt_Batch", "RNA_Batch", "Library_Batch", "Seq_pool", "Episomal_integration" )

heatmap.2(as.matrix(pc_cov_cor[Tech_fac,PCs]),col=brewer.pal(4, "Greens"), trace="none",
          Rowv=FALSE, Colv=FALSE, key=T, main="Cor. PCs & tech factors", dendrogram="none",
          key.title=NA, cexRow=0.9, cexCol=0.9)
```


```{r}
log10_pc_cov_pval <- -log(pc_cov_pval)
heatmap.2(as.matrix(log10_pc_cov_pval[Tech_fac,PCs]), col=brewer.pal(9, "Greens"), trace="none",
          Rowv=FALSE, Colv=FALSE, key=T, main="-log10 pval of cor. PCs & tech factors", dendrogram="none",
          key.title=NA, cexRow=0.9, cexCol=0.9)
```


