---
title: "Compare Nuclear Fraction PAS Double filter"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(reshape2)
library(tidyverse)

```


Compare nuclear fraction PAS between human and chimp.  I need to fiter the phenotypes fixed for leafcutter further to have only the double filtered PAS. 

I need a fc file with the human and chimp nuclear samples. I will make a group file with the identifier being human or chimp.  

These were created with prepareCleanLiftedFC_5perc4LC.py  (in human and chimp extra analysis TvN)  

../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc
../Human/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc


```{bash,eval=F}
mkdir ../data/NuclearHvC_DF
```



```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% mutate(chrom=paste(chr,start,end,gene, sep=":"))
```


hee feature counts are in  ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc. I need to subset these for those in the annotations.  Keep the PAS in this file: ../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed. I will do this with a python script.  

```{bash,eval=F}
mkdir ../Chimp/data/CleanLiftedPeaks4LC/
python prepareCleanLiftedFC_5perc4LC.py ../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc ../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed ../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc
```

```{bash,eval=F}
mkdir ../Human/data/CleanLiftedPeaks4LC/
python prepareCleanLiftedFC_5perc4LC.py ../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc ../data/Peaks_5perc/Peaks_5perc_either_HumanCoordHummanUsage.bed ../Human/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc
```

```{r}
human=read.table("../Human/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc", stringsAsFactors = F, header = T) %>% rownames_to_column(var="chrom") %>% filter(chrom %in% PAS$chrom)


chimp=read.table("../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc", stringsAsFactors = F, header = T)%>% rownames_to_column(var="chrom")%>% filter(chrom %in% PAS$chrom)
```


```{r}
Allsamps=human %>% full_join(chimp,by="chrom") 

AllNuclear=Allsamps %>% dplyr::select(chrom,contains("_N")) %>% column_to_rownames(var="chrom")

write.table(AllNuclear, "../data/NuclearHvC_DF/ALLPAS_postLift_LocParsed_HvC_Nuclear_fixed4LC_Doublefilt.fc",row.names = T, col.names = T, quote = F)

```


I will make the id file here.  

```{r}
Inds=colnames(AllNuclear) 
Species=c(rep("Human",5), rep("Chimp", 6))

idFileDF=as.data.frame(cbind(Inds,Species))

write.table(idFileDF, "../data/NuclearHvC_DF/sample_goups.txt",row.names = F, col.names = F, quote = F)

```


Split by chromosome.  



```{bash,eval=F}
mkdir ../data/DiffIso_Nuclear_DF/

python subset_diffisopheno_Nuclear_HvC_DF.py 1
python subset_diffisopheno_Nuclear_HvC_DF.py 2
python subset_diffisopheno_Nuclear_HvC_DF.py 3
python subset_diffisopheno_Nuclear_HvC_DF.py 4
python subset_diffisopheno_Nuclear_HvC_DF.py 5
python subset_diffisopheno_Nuclear_HvC_DF.py 6
python subset_diffisopheno_Nuclear_HvC_DF.py 7
python subset_diffisopheno_Nuclear_HvC_DF.py 8
python subset_diffisopheno_Nuclear_HvC_DF.py 9
python subset_diffisopheno_Nuclear_HvC_DF.py 10
python subset_diffisopheno_Nuclear_HvC_DF.py 11
python subset_diffisopheno_Nuclear_HvC_DF.py 12
python subset_diffisopheno_Nuclear_HvC_DF.py 13
python subset_diffisopheno_Nuclear_HvC_DF.py 14
python subset_diffisopheno_Nuclear_HvC_DF.py 16
python subset_diffisopheno_Nuclear_HvC_DF.py 18
python subset_diffisopheno_Nuclear_HvC_DF.py 19
python subset_diffisopheno_Nuclear_HvC_DF.py 20
python subset_diffisopheno_Nuclear_HvC_DF.py 21
python subset_diffisopheno_Nuclear_HvC_DF.py 22
```

Run leafcutter:

```{bash,eval=F}

sbatch runNuclearDiffIso_DF.sh
```

Concatinate results:  

```{bash,eval=F}
awk '{if(NR>1)print}' ../data/DiffIso_Nuclear_DF/TN_diff_isoform_chr*.txt_effect_sizes.txt > ../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt


awk '{if(NR>1)print}' ../data/DiffIso_Nuclear_DF/TN_diff_isoform_chr*.txt_cluster_significance.txt > ../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_significance.txt

```


Significant clusters:  

```{r}
sig=read.table("../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sig$p.adjust=as.numeric(as.character(sig$p.adjust))
```



```{r}
qqplot(-log10(runif(nrow(sig))), -log10(sig$p.adjust),ylab="-log10 Adjusted Leafcutter pvalue", xlab="-log 10 Uniform expectation", main="Leafcutter differencial isoform analysis between Species")
abline(0,1)
```


```{r}
tested_genes=nrow(sig)
tested_genes
```

```{r}
sig_genes=sig %>% filter(p.adjust<.05)
number_sig_genes=nrow(sig_genes)
number_sig_genes
```

Effect Sizes  

```{r}
effectsize=read.table("../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('intron',  'logef' ,'Chimp', 'Human','deltaPAU')) %>% filter(intron != "intron")

effectsize$deltaPAU=as.numeric(as.character(effectsize$deltaPAU))
effectsize$logef=as.numeric(as.character(effectsize$logef))
```


```{r}
plot(sort(effectsize$deltaPAU),main="Leafcutter delta PAU", ylab="Delta PAU", xlab="PAS Index")
```

Effect size >.2 color:  

```{r}
effectsize_col= effectsize %>% arrange(deltaPAU) %>%  mutate(diffuse=ifelse(abs(deltaPAU)>=.2,"Yes", "No"))
effectsize_col$color <- ifelse(effectsize_col$diffuse=="Yes", "red", "black")

plot(sort(effectsize_col$deltaPAU),main="Leafcutter delta PAU", ylab="Delta PAU", xlab="PAS Index",col=alpha(effectsize_col$color, 0.6))
legend("topleft", legend=c("abs(DPAU) > .2, (n=3103)", "Not differentially used (n=37673)"), col=c("red", "black"), pch=16:16, cex=0.8)
```


Are those discovered used more in chimp those discovered in chimp?

```{r}
PASinfo=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt",header = T, stringsAsFactors = F)
```

Join this with the effect sizes.  

```{r}
effectsize_sep=effectsize %>% separate(intron, into=c("chr", "start", "end", "gene"),sep=":")
effectsize_sep$start=as.integer(effectsize_sep$start)
effectsize_sep$end=as.integer(effectsize_sep$end)
effectsize_anno=effectsize_sep %>% inner_join(PASinfo, by=c("chr", "start", "end","gene"))

```
```{r}
ggplot(effectsize_anno, aes(x=disc, y=deltaPAU)) + geom_boxplot()
```

Volcano plot: 

I need the effect sizes and the significance. I need to plot only the top PAS per cluster.  

```{r}
sig_geneP=sig %>% separate(cluster,into = c("chr", "gene"), sep=":") %>% dplyr::select(gene, p.adjust)

effectsizeTop=effectsize_sep %>% group_by(gene) %>% summarise(Min=min(deltaPAU), Max=max(deltaPAU)) %>% mutate(TopdPAU=ifelse(abs(Min)>Max, Min, Max))

#exclude when the max=min 
effectsizeTopFilt=effectsizeTop %>% filter(abs(Min) != Max)

effectsize_wES=effectsizeTopFilt %>% inner_join(sig_geneP, by="gene") %>% mutate(Species=ifelse(TopdPAU > 0.2 & p.adjust<.05, "Human", ifelse(TopdPAU < -0.2 & p.adjust< .05, "Chimp", "Neither")))
```

This is the significance for the gene. 


```{r}
ggplot(effectsize_wES,aes(x=TopdPAU, y=-log10(p.adjust))) +geom_point(aes(col=Species),alpha=.5) + labs(title="Top PAS per gene \nExclude 2 PAS genes")+ geom_text(data=subset(effectsize_wES, -log10(p.adjust) >20 & abs(TopdPAU)>.2 ), aes(x=TopdPAU,y=-log10(p.adjust) +2,label=gene))
``` 
Not the best way to visualize this because every PAS per gene is assigned the same pvalue. 


Try this including the matching one. I will make 2 plots. One with human dominant, one with chimp dominant. 

```{r}
effectsizeTopHuman=effectsize_sep %>% group_by(gene) %>% summarise(Min=min(deltaPAU), Max=max(deltaPAU)) %>% mutate(TopdPAU=ifelse(abs(Min) > Max, Min, ifelse(abs(Min)==Max, Min, Max)),TwoPAS=ifelse(abs(Min)==Max, T, F))

effectsize_wES_human=effectsizeTopHuman %>% inner_join(sig_geneP, by="gene") %>% mutate(Species=ifelse(TopdPAU > 0.2 & p.adjust<.05, "Human", ifelse(TopdPAU < -0.2 & p.adjust< .05, "Chimp", "Neither")))

effectsizeTopChimp=effectsize_sep %>% group_by(gene) %>% summarise(Min=min(deltaPAU), Max=max(deltaPAU)) %>% mutate(TopdPAU=ifelse(abs(Max)>=abs(Min), Max, Min), TwoPAS=ifelse(abs(Min)==Max, T, F))

effectsize_wES_chimp=effectsizeTopChimp %>% inner_join(sig_geneP, by="gene") %>% mutate(Species=ifelse(TopdPAU > 0.2 & p.adjust<.05, "Human", ifelse(TopdPAU < -0.2 & p.adjust< .05, "Chimp", "Neither")))
```


```{r}
ggplot(effectsize_wES_human,aes(x=TopdPAU, y=-log10(p.adjust))) + geom_point(aes(col=Species, shape=TwoPAS),alpha=.5) + labs(title="Top PAS per gene \nHuman dominant for 2 PAS")+ geom_text(data=subset(effectsize_wES_human,-log10(p.adjust) >20 & abs(TopdPAU)>.2 ), aes(x=TopdPAU,y=-log10(p.adjust) +2,label=gene))

```
```{r}
ggplot(effectsize_wES_chimp,aes(x=TopdPAU, y=-log10(p.adjust))) + geom_point(aes(col=Species, shape=TwoPAS),alpha=.5)  + labs(title="Top PAS per gene \nChimp dominant for 2 PAS")+ geom_text(data=subset(effectsize_wES_chimp, -log10(p.adjust) >20 & abs(TopdPAU)>.2), aes(x=TopdPAU,y=-log10(p.adjust) +2,label=gene)) 

```

Write out the significant genes with >.2 difference.

```{r}
effectsize_sep_pval=effectsize_sep %>% full_join(sig_geneP, by="gene")
#significant > .2
effectsize_sep_pval_sig= effectsize_sep_pval %>% filter(p.adjust <= .05,abs(deltaPAU) >=0.2)
nrow(effectsize_sep_pval_sig)
#genes
effectsize_sep_pval_sig_genes=effectsize_sep_pval_sig %>% dplyr::select(gene) %>% unique()
nrow(effectsize_sep_pval_sig_genes)

effectsize_sep_pval_colforsig=effectsize_sep_pval %>% mutate(SigPAU2=ifelse(p.adjust <= .05 &abs(deltaPAU) >=0.2, "Yes","No"))
```



```{r}
effectsize_wES_chimpOnly= effectsize_wES %>% filter(Species=="Chimp")
effectsize_wES_HumanOnly= effectsize_wES %>% filter(Species=="Human")


write.table(effectsize_sep_pval_colforsig,"../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt", col.names = T, row.names = F, quote = F )
  
write.table(effectsize_wES_chimpOnly,"../data/DiffIso_Nuclear_DF/SignifianceChimpPAS_2_Nuclear.txt",col.names =T, row.names = F,quote = F)

write.table(effectsize_wES_HumanOnly,"../data/DiffIso_Nuclear_DF/SignifianceHumanPAS_2_Nuclear.txt",col.names =T, row.names = F,quote = F)

write.table(effectsize_sep_pval_sig,"../data/DiffIso_Nuclear_DF/SignifianceEitherPAS_2_Nuclear.txt",col.names =T, row.names = F,quote = F)

write.table(effectsize_sep_pval_sig_genes,"../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt",col.names =T, row.names = F,quote = F)
```





