---
title: "Establish Ortho 3' UTRs"
author: "Briana Mittleman"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

I want to get a list of ortho 3' UTRs. This is important because I want to look at the differentially used PAS and see if they make longer or shorter 3' UTRs. This could point to regulatory variation.  


I will start with the ortho exon file. Convert these to bed files to be able to merge 

```{bash,eval=F}
grep three_prime_utr /project2/gilad/kenneth/OrthoExonPartialMapping/human.noM.gtf > ../data/OrthoExonBed/Human3UTR.gtf
grep three_prime_utr /project2/gilad/kenneth/OrthoExonPartialMapping/chimp.noM.gtf > ../data/OrthoExonBed/Chimp3UTR.gtf

python SAF2Bed.py ../data/OrthoExonBed/Human3UTR.gtf ../data/OrthoExonBed/Human3UTR.bed
python SAF2Bed.py ../data/OrthoExonBed/Chimp3UTR.gtf ../data/OrthoExonBed/Chimp3UTR.bed


sort -k1,1 -k2,2n ../data/OrthoExonBed/Human3UTR.bed >  ../data/OrthoExonBed/Human3UTR.sort.bed
sort -k1,1 -k2,2n ../data/OrthoExonBed/Chimp3UTR.bed >  ../data/OrthoExonBed/Chimp3UTR.sort.bed

#mrge by strand and gene 

bedtools merge -s -i ../data/OrthoExonBed/Human3UTR.sort.bed -c 4,5,6 -o distinct,mean,distinct > ../data/OrthoExonBed/Human3UTR.merged.sort.bed

bedtools merge -s -i ../data/OrthoExonBed/Chimp3UTR.sort.bed -c 4,5,6 -o distinct,mean,distinct > ../data/OrthoExonBed/Chimp3UTR.merged.sort.bed
```

Look at the human one and the annotation in humans:  

```{bash,eval=F}
mkdir ../data/orthoUTR

grep utr3 ../../genome_anotation_data/hg38_refseq_anno/hg38_ncbiRefseq_Formatted_Allannotation_noSNO.Resort.bed > ../data/orthoUTR/g38_ncbiRefseq_Formatted_Allannotation_UTR3.bed
```

There are a lot of exons in this set that look like they are not utrs.  

For this analsis. I need one 3' UTR per gene. I can look at the most distal one first.  

I can do this here with if else statements:  

+ strand: max start  

- strand:

```{r}
humanMergeutr= read.table("../data/OrthoExonBed/Human3UTR.merged.sort.bed", col.names = c('chr','start','end','gene','score','strand'), stringsAsFactors = F) %>% group_by(gene)

humanMergeutrpos= humanMergeutr %>% filter(strand=="+") %>% group_by(gene) %>% top_n(1,start)

humanMergeutrneg= humanMergeutr %>% filter(strand=="-") %>% group_by(gene) %>% top_n(-1,start)

humanDistalboth=humanMergeutrpos %>% bind_rows(humanMergeutrneg)
write.table(humanDistalboth, "../data/orthoUTR/HumanDistal3UTR.bed", col.names = F, row.names = F, quote = F, sep="\t")
```

```{bash,eval=F}
sort -k1,1 -k2,2n ../data/orthoUTR/HumanDistal3UTR.bed > ../data/orthoUTR/HumanDistal3UTR.sort.bed
```

Looks like that worked.  

Do it with an intercect with the anno and see if the results are similar:  

```{bash,eval=F}
bedtools intersect -s -wa -a ../data/OrthoExonBed/Human3UTR.merged.sort.bed -b ../data/orthoUTR/g38_ncbiRefseq_Formatted_Allannotation_UTR3.bed > ../data/OrthoExonBed/Human3UTR.merged.sort.overlapHumanannno.bed
```
overlap has 
135340
17680


I will now work with the most distal PAS. I want to see how many of the PAS fall in these regions. 

```{bash,eval=F}
bedtools intersect -s -wa -a ../data/PAS_doubleFilter/PAS_doublefilter_either_HumanCoordHummanUsage.bed -b ../data/orthoUTR/HumanDistal3UTR.sort.bed > ../data/orthoUTR/PASOverlapinDistal3UTR.bed

bedtools intersect -s -wa -wb -a ../data/PAS_doubleFilter/PAS_doublefilter_either_HumanCoordHummanUsage.bed -b ../data/orthoUTR/HumanDistal3UTR.sort.bed > ../data/orthoUTR/PASOverlapinDistal3UTR_bothWritten.bed
```

```{r}
metaPAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", stringsAsFactors = F, header = T) %>% select(gene, PAS, loc)
Overlapping= read.table("../data/orthoUTR/PASOverlapinDistal3UTR_bothWritten.bed", col.names = c("chrpas", "startpas", "endpas","PAS", "humanusage", "strandpas", "chrutr", "startutr", "endutr","geneUTR", "score","strand"),stringsAsFactors = F) %>% select(-strandpas, -chrutr, -startutr, -endutr)  %>% inner_join(metaPAS, by=c("PAS"))

 Overlapping %>% filter(gene != geneUTR) %>% nrow()
 
 
 Overlapping %>% filter(loc!="utr3") %>% nrow()
 
 Overlappingfilt= Overlapping%>% filter(gene == geneUTR)
```

184 of these are in different genes. check these.  
405 not annotated as UTR3  


```{r}
NotUTRanno= Overlapping %>% filter(loc!="utr3") %>% filter(gene == geneUTR)
nrow(NotUTRanno)
ggplot(NotUTRanno,aes(x=loc, fill=loc) ) + geom_histogram(stat="count") + scale_fill_brewer(palette = "Dark2") + labs(title="Location of 400 PAS overlapping PAS but not annotated 3' UTR")
```
This is good. It looks like most of the time the problem is just that the ortho utr extends the original annotation. We call it an end or cds.  

```{r}
nrow(Overlapping)/nrow(metaPAS)

nrow(Overlapping)/nrow(metaPAS %>% filter(loc=="utr3"))
```
almost 70% of the 3' UTR pas are represented.

Look at how many genes are represented  
```{r}

OverlappingGenes= Overlapping %>% filter(gene == geneUTR)  %>% group_by(gene) %>% summarise(nPAS=n())
nrow(OverlappingGenes)
nrow(OverlappingGenes %>% filter(nPAS>1))

OverlappingGenes$nPAS= as.factor(OverlappingGenes$nPAS)

ggplot(OverlappingGenes, aes(x=nPAS)) + geom_histogram(stat="count") + labs(x="Number of PAS", y="Genes", title="Number of PAS per gene with PAS in ortho 3' UTR")
```

There is at least one pas in 6931 genes. Thats pretty good.  There are 3098 with at least 2.  


Look to see if these represent any of those differenctially used:  

```{r}
PASGene=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt",stringsAsFactors = F, header = T) %>% select(PAS, chr, start, end,loc)
DiffUsed=read.table("../data/DiffIso_Nuclear_DF/AllPAS_withGeneSig.txt",header = T,stringsAsFactors = F) %>% inner_join(PASGene, by=c("chr",'start','end')) %>% mutate(OrthoUTR=ifelse(PAS %in% Overlappingfilt$PAS, "Yes","No"))
DiffUsed_sig =DiffUsed %>% filter(SigPAU2=="Yes")
```


```{r}
ggplot(DiffUsed_sig,aes(x=OrthoUTR)) +geom_histogram(stat="count")


```

Proportion: 

```{r}
DiffUsed_sig %>% group_by(OrthoUTR) %>% summarise(nEach=n()) %>% mutate(prop=nEach/nrow(DiffUsed_sig))
```

37% of the differentially used are in the 3' UTR.  

Filter to utr  

```{r}
DiffUsed_sig %>% filter(loc=="utr3")%>%group_by(OrthoUTR) %>% summarise(nEach=n()) %>% mutate(prop=nEach/nrow(DiffUsed_sig%>% filter(loc=="utr3")))

```

834 sig are in the 3' UTR.


This seems like an ok way to do this. Write out the filtered overlap to use in the next analysis. THerea are about 7K genes in this set

```{r}
 write.table(Overlappingfilt, "../data/orthoUTR/FilteredPASOverlapOrthoUTR.text",col.names = T, quote = F, row.names = F)
```


