---
title: "Ortho Exon evaluation as an annotation"
author: "Briana Mittleman"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

I am still concerned about annotation. I want to see if using the ortho exon file as an annotation tool would be more appropriate. I want to look at it and see if there is a way to combine exons by gene and see if the spaces inbetween could be introns.  

First look at the human one.  

```{r}
HumanOrtho=read.table("/project2/gilad/kenneth/OrthoExonPartialMapping/human.noM.gtf", sep="\t")
```

Parse this into a bed file with a python script.  

```{bash,eval=F}
mkdir ../data/OrthoExonBed
python SAF2Bed.py /project2/gilad/kenneth/OrthoExonPartialMapping/human.noM.gtf ../data/OrthoExonBed/human.noM.bed

sort -k1,1 -k2,2n ../data/OrthoExonBed/human.noM.bed > ../data/OrthoExonBed/human.noM.sort.bed 
```

##All PAS

```{bash,eval=F}
sbatch overlapPASandOrthoexon.sh 
```

Results:  
```{r}
PASMeta=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt",header = T,stringsAsFactors = F) %>% select(PAS, gene,loc)
OverlapPAS=read.table("../data/OrthoExonBed/allPASinOrtho.bed", col.names = c("chr", "start", "end", "PAS", "score", "strand"),stringsAsFactors = F)%>% group_by(PAS) %>% summarize(n=n())%>%  mutate(OE="In") %>% inner_join(PASMeta,by="PAS")
nrow(OverlapPAS)


NotOverlapPAS=read.table("../data/OrthoExonBed/allPAS_NOT_inOrtho.bed", col.names = c("chr", "start", "end", "PAS", "score", "strand"),stringsAsFactors = F)%>% group_by(PAS) %>% summarize(n=n())%>%  mutate(OE="OUT") %>% inner_join(PASMeta,by="PAS")
nrow(NotOverlapPAS)


ALLPAS_ortho=OverlapPAS %>% bind_rows(NotOverlapPAS)

ggplot(ALLPAS_ortho, aes(x=OE, by=loc,fill=loc)) +geom_bar(stat="count",position = "dodge") + labs(x="Is PAS overlapping ortho exon", title="All PAS in OrthoExon",y="Number of PAS")+ scale_fill_brewer(palette = "Dark2")

```
Ok this is the expected distribution, We expect the coding and 3' UTRs to be in the ortho exon file. What is more interesting is genes with and without exons in the ortho exon.  



Take this to the gene level. 

```{r}
OrthoBed=read.table("../data/OrthoExonBed/human.noM.sort.bed", col.names = c("chr","start","end","gene", "nExon","strand"),stringsAsFactors = F) %>% group_by(gene) %>% summarise(nExon=n())
```

Now I look to see which of the PAS are in genes not in the ortho exon.  

```{r}
PASMeta_GeneOE= PASMeta %>% mutate(OE=ifelse(gene %in% OrthoBed$gene, "Yes", "No"))

PASMeta_GeneOEgene= PASMeta_GeneOE %>% group_by(gene, OE) %>% summarise()

ggplot(PASMeta_GeneOEgene, aes(x=OE, fill=OE))+ geom_histogram(stat="count") + labs(x="Is gene in Ortho Exon", y="Genes") + scale_fill_brewer(palette = "Dark2")
```
Look at the genes without:  

```{r}
PASMeta_GeneOE_NO= PASMeta_GeneOE %>% filter(OE=="No")
nrow(PASMeta_GeneOE_NO)

ggplot(PASMeta_GeneOE_NO,aes(x=loc, fill=OE)) + geom_bar(stat="count")


```
```{r}
Genesnotin=PASMeta_GeneOE_NO %>% group_by(gene) %>% summarise(n=n())
```

1000 genes not in ortho exons.  



```{r}
#sno
Genesnotin %>% filter(grepl("SNO",gene)) %>% nrow()
#linc
Genesnotin %>% filter(grepl("LINC",gene)) %>% nrow()
#loc
Genesnotin %>% filter(grepl("LOC",gene)) %>% nrow()
```

Are these the genes with different dominant PAS.  

```{r}
allPAS= read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T) 
ChimpPASwMean =allPAS %>% dplyr::select(-Human)
HumanPASwMean =allPAS %>% dplyr::select(-Chimp)

Chimp_Dom= ChimpPASwMean %>%
  group_by(gene) %>%
  top_n(1,Chimp) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,Chimp) %>% 
  rename(ChimpLoc=loc, ChimpPAS=PAS)

Human_Dom= HumanPASwMean %>% 
  group_by(gene) %>% 
  top_n(1, Human) %>% 
  mutate(nPer=n()) %>% 
  filter(nPer==1) %>% 
  dplyr::select(gene,loc,PAS,Human) %>% 
  rename(HumanLoc=loc, HumanPAS=PAS)


#merge

BothDom= Chimp_Dom %>% inner_join(Human_Dom,by="gene")
DifDom= BothDom %>% filter(ChimpPAS!=HumanPAS)


```
Plot this before I remove those genes.  
```{r}
DifDom_before=DifDom %>% select(gene, ChimpLoc, HumanLoc) %>% gather("species","loc",-gene)

ggplot(DifDom_before, aes(x=loc, by=species, fill=species))+geom_histogram(position = "dodge",stat = "count")+ labs(title="Location of PAS with different Dominant",y="PAS")+scale_fill_brewer(palette = "Dark2")
```
Remove the not ortho genes:  

```{r}
DifDom_after=DifDom_before %>% anti_join(Genesnotin,by="gene")
ggplot(DifDom_after, aes(x=loc, by=species, fill=species))+geom_histogram(position = "dodge",stat = "count")+ labs(title="Location of PAS with different Dominant\n after removing genes not in orthoexon",y="PAS")+scale_fill_brewer(palette = "Dark2")
```
That didnt help.  

Check the matching ones (are these in ortho exon)

```{r}
SameDom= BothDom %>% filter(ChimpPAS==HumanPAS)
nrow(SameDom)
SameDom_after=SameDom %>% anti_join(Genesnotin,by="gene")
nrow(SameDom_after)
```
Where did I lose genes.  

```{r}
6906/7638
3432/4028
```
Lose more in the different dominant than in the same dominant. 
Percent lost=

```{r}
(7638-6906)/7638
(4028-3432)/4028
```



##MultiMap  

I will check if some of the problem genes, I found in my pipeline are in this.  

```{r}
ChimpMM=read.table("../data/multimap/Chimp_Uniq_multimapPAS.txt", stringsAsFactors = F, header = T)
HumanMM=read.table("../data/multimap/Human_Uniq_multimapPAS.txt", stringsAsFactors = F, header = T)
BothMM=read.table("../data/multimap/Both_multimapPAS.txt",stringsAsFactors = F, header = T)


AllMM=ChimpMM %>% bind_rows(HumanMM) %>% bind_rows(BothMM)

```


I will overlap this with the ortho exon file. I will look at those that overlap and those that do not. I need a bed file    


```{r}
AllMM_bed=AllMM %>% mutate(Name=paste(gene, PAS,loc, MultiMap, sep=":")) %>% select(chr, start,end, Name, Human, strandFix)

write.table(AllMM_bed,"../data/multimap/allMM.bed",col.names = F, quote = F, row.names = F, sep="\t")
```


```{bash,eval=F}
sort -k1,1 -k2,2n ../data/multimap/allMM.bed > ../data/multimap/allMM.sort.bed
```


Overlap.  


```{bash,eval=F}
sbatch overlapMMandOrthoexon.sh 
```


Results:  
```{r}
InOrtho=read.table("../data/OrthoExonBed/allMMinOrtho.bed", col.names = c("chr", "start", "end", "name", "score", "strand")) %>% group_by(name) %>% summarize(n=n())%>% separate(name, into=c("gene", "PAS","loc", "MM"),sep=":") %>% mutate(OE="In")
NotInOrtho=read.table("../data/OrthoExonBed/allMM_NOT_inOrtho.bed", col.names = c("chr", "start", "end", "name", "score", "strand")) %>% group_by(name) %>% summarize(n=n()) %>% separate(name, into=c("gene", "PAS","loc", "MM"),sep=":")  %>% mutate(OE="Out")

AllOrthores=InOrtho %>% bind_rows(NotInOrtho)

ggplot(AllOrthores, aes(x=OE, by=loc,fill=loc)) +geom_bar(stat="count",position = "dodge") + labs(x="Is PAS overlapping ortho exon", title="PAS impacted by multimapping and ortho exon",y="Number of PAS")+ scale_fill_brewer(palette = "Dark2")

```

Proportion:  

```{r}
AllOrthores %>% group_by(OE) %>% summarise(n=n())

#look only at utr  

AllOrthores %>% filter(loc=="utr3") %>% group_by(OE) %>% summarise(n=n())
```

Look at the UTR sequences that are in:  

```{r}
AllOrthores %>% filter(OE=="In", loc=="utr3")
``` 

##distance to next annotated ortho exon boundry  

I can look at the intronic PAS in both species to see how far they are from an ortho exon.  

I need to merge exons.  

```{bash,eval=F}
sort -k1,1 -k2,2n ../data/OrthoExonBed/chimp.noM.bed >  ../data/OrthoExonBed/chimp.noM.sort.bed 
bedtools merge -s -c 4,5,6 -o distinct,count,distinct  -i  ../data/OrthoExonBed/chimp.noM.sort.bed >  ../data/OrthoExonBed/chimp.noM.sort.merged.bed 
bedtools merge -s -c 4,5,6 -o distinct,count,distinct  -i  ../data/OrthoExonBed/human.noM.sort.bed >  ../data/OrthoExonBed/human.noM.sort.merged.bed 

mkdir ../data/TestAnnoBiasOE
```

Now I can map all of the intronic PAS  

For intronic PAS in genes in the ortho exon, filter the bed files 
```{r}
PASMeta_GeneOE_intronYes= PASMeta_GeneOE %>% filter(loc=="intron", OE=="Yes")

HumanBed=read.table("../data/PAS_doubleFilter/PAS_doublefilter_either_HumanCoordHummanUsage.sort.bed", col.names = c("chr", 'start','end','PAS','usage','strand'),stringsAsFactors = F) %>% semi_join(PASMeta_GeneOE_intronYes, by="PAS")

write.table(HumanBed, "../data/TestAnnoBiasOE/HumanIntronicGeneinOE.bed",col.names = F, row.names = F, quote = F, sep="\t")

ChimpBed=read.table("../data/PAS_doubleFilter/PAS_doublefilter_either_ChimpCoordChimpUsage.sort.bed", col.names = c("chr", 'start','end','PAS','usage','strand'),stringsAsFactors = F) %>% semi_join(PASMeta_GeneOE_intronYes, by="PAS")

write.table(ChimpBed, "../data/TestAnnoBiasOE/ChimpIntronicGeneinOE.bed",col.names = F, row.names = F, quote = F, sep="\t")


```

Find the closest:  

same strand. look only upstream (-id) to say closest annotated 5' splice site.  then do only downstream to say closest annotated 3' splice site.  

```{bash,eval=F}
sbatch ClosestorthoEx.sh
```

Look at upstream  


If the gene is a different gene, make it 0

```{r}
HumanUpstream=read.table("../data/TestAnnoBiasOE/HumanUpstream.intronic.txt",stringsAsFactors = F, col.names = c("chrPAS", "startPAS", "endPAS", "PAS", "HumanUsage", "strandPAS", "chrExon","startExon","endExon", "Orthogene", "n","strandIntron", "UpstreamdistancePAS2Exon")) %>% inner_join(PASMeta,by="PAS") %>% mutate(Fixed=ifelse(gene==Orthogene, UpstreamdistancePAS2Exon,0))
HumanDownstream=read.table("../data/TestAnnoBiasOE/HumanDownstream.intronic.txt",stringsAsFactors = F, col.names = c("chrPAS", "startPAS", "endPAS", "PAS", "HumanUsage", "strandPAS", "chrExon","startExon","endExon", "Orthogene", "n","strandIntron", "DownstreamdistancePAS2Exon")) %>% inner_join(PASMeta,by="PAS") %>% mutate(Fixed=ifelse(gene==Orthogene, DownstreamdistancePAS2Exon,0))



HumanBoth=as.data.frame(cbind(HumanUpstream[,1:6],gene=HumanUpstream[,14], "UpstreamHuman"=HumanUpstream[,16],"DownstreamHuman"=HumanDownstream[,16])) %>% 
  mutate(DominanceHuman=ifelse(PAS %in% BothDom$HumanPAS, "yes","no"))
```

Chimp:

```{r}
ChimpUpstream=read.table("../data/TestAnnoBiasOE/ChimpUpstream.intronic.txt",stringsAsFactors = F, col.names = c("chrPAS", "startPAS", "endPAS", "PAS", "ChimpUsage", "strandPAS", "chrExon","startExon","endExon", "Orthogene", "n","strandIntron", "UpstreamdistancePAS2Exon"))  %>% inner_join(PASMeta,by="PAS") %>% mutate(Fixed=ifelse(gene==Orthogene, UpstreamdistancePAS2Exon,0))
ChimpDownstream=read.table("../data/TestAnnoBiasOE/ChimpDownstream.intronic.txt",stringsAsFactors = F, col.names = c("chrPAS", "startPAS", "endPAS", "PAS", "ChimpUsage", "strandPAS", "chrExon","startExon","endExon", "Orthogene", "n","strandIntron", "DownstreamdistancePAS2Exon"))  %>% inner_join(PASMeta,by="PAS") %>% mutate(Fixed=ifelse(gene==Orthogene, DownstreamdistancePAS2Exon,0))
ChimpBoth=as.data.frame(cbind(PAS=ChimpUpstream[,4], 'UpstreamChimp'=ChimpUpstream[,16],'DownstreamChimp'=ChimpDownstream[,16])) %>% 
  mutate(DominanceChimp=ifelse(PAS %in% BothDom$ChimpPAS, "yes","no"))
```


Join together:  

```{r}
BothSpeciesDistance= HumanBoth %>% inner_join(ChimpBoth, by="PAS")

BothSpeciesDistance$UpstreamChimp=as.numeric(as.character(BothSpeciesDistance$UpstreamChimp))
BothSpeciesDistance$DownstreamChimp=as.numeric(as.character(BothSpeciesDistance$DownstreamChimp))
```

Look at distance based on human dominance then chimp dominance  

```{r}
ggplot(BothSpeciesDistance, aes(x=UpstreamHuman, y=UpstreamChimp, col=DominanceHuman)) + geom_point(alpha=.3)+ geom_abline(aes(slope=1,intercept=0))

```

```{r}
ggplot(BothSpeciesDistance, aes(x=UpstreamHuman, y=UpstreamChimp, col=DominanceHuman)) + geom_point(alpha=.3) + ylim(c(-10000,0)) +xlim(c(-10000,0)) + geom_abline(aes(slope=1,intercept=0)) + scale_color_brewer(palette = "Dark2")

```


I want to color by dominant in human, chimp, both neither 

```{r}
BothSpeciesDistance_dom=BothSpeciesDistance %>% mutate(Dominance=ifelse(DominanceHuman=="yes", ifelse(DominanceChimp=="yes", "Both", "human"), ifelse(DominanceChimp=="yes", "chimp", "Neither")))
```

Upstream Plots:  

```{r}
ggplot(BothSpeciesDistance_dom, aes(x=UpstreamHuman, y=UpstreamChimp, col=Dominance)) + geom_point(alpha=.3)+ geom_abline(aes(slope=1,intercept=0)) + geom_smooth(method="lm", alpha=.1)+scale_color_brewer(palette = "Dark2") 
```
```{r}
ggplot(BothSpeciesDistance_dom, aes(x=UpstreamHuman, y=UpstreamChimp, col=Dominance)) + geom_point(alpha=.3)+ geom_abline(aes(slope=1,intercept=0))+ylim(c(-50000,0)) +xlim(c(-50000,0)) + scale_color_brewer(palette = "Dark2") + geom_smooth(method="lm", alpha=.1)
```

Downstream  

```{r}
ggplot(BothSpeciesDistance_dom, aes(x=DownstreamHuman, y=DownstreamChimp, col=Dominance)) + geom_point(alpha=.3)+ geom_abline(aes(slope=1,intercept=0)) + scale_color_brewer(palette = "Dark2") + geom_smooth(method="lm", alpha=.1) + geom_smooth(method="lm", alpha=.1)
```
There are a lot of PAS that are far from an exon on the human side. 
```{r}
ggplot(BothSpeciesDistance_dom, aes(x=DownstreamHuman, y=DownstreamChimp, col=Dominance)) + geom_point(alpha=.3)+ geom_abline(aes(slope=1,intercept=0))+ylim(c(0,50000)) +xlim(c(0,50000)) + scale_color_brewer(palette = "Dark2") + geom_smooth(method="lm", alpha=.01)
```

Look at the top examples:  

```{r}
BothSpeciesDistance_dom_TopDiffUp=BothSpeciesDistance_dom %>% mutate(Updiff=abs(UpstreamHuman-UpstreamChimp)) %>% select(PAS, gene, Updiff) %>% arrange(desc(Updiff))

head(BothSpeciesDistance_dom_TopDiffUp)
```

