---
title: "Number of PAS by Species"
author: "Briana Mittleman"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(reshape2)
library(workflowr)
```


###Both fraction
In this analysis I will look at thenumber of PAS per species at a gene level. I am only including PAS in chr1-22. These results use mean usage accross fraction.  

```{r}
PAS=read.table("../data/Peaks_5perc/Peaks_5perc_either_bothUsage_noUnchr.txt", stringsAsFactors = F, header = T)
```


I want to look at the number of PAS at 5% in each gene by human and chimp.  

```{r}
PAS_sm=PAS %>% select(gene, Chimp, Human) 
PAS_m= melt(PAS_sm, id.var="gene", variable.name="species", value.name="meanUsage") %>% filter(meanUsage >=0.05) %>% group_by(species, gene) %>% summarise(nPAS=n())

#pos = more human, neg = more chimp 
PAS_spread=PAS_m %>% spread(species, nPAS, fill=0) %>% mutate(DiffPAS=Human-Chimp)
```

```{r}
summary(PAS_spread$DiffPAS)

#more human
nrow(PAS_spread[PAS_spread$DiffPAS>0,])

#more chimp
nrow(PAS_spread[PAS_spread$DiffPAS<0,])

#same
nrow(PAS_spread[PAS_spread$DiffPAS==0,])

#all
nrow(PAS_spread)
```


```{r}
prop.test(x=c(3759, 2596), n=c(15711,15711), alternative ="greater")

```
```{r}
ggplot(PAS_spread, aes(x=DiffPAS)) + geom_bar(stat="count") +geom_vline(xintercept = mean(PAS_spread$DiffPAS),col="red") + labs(title="Difference in number of PAS at 5% Human vs Chimp", y="Genes", x="N Human PAS - N Chimp PAS")
```

Plot distribution of N pas by species:  

Wilcoxan test to see if there is a difference in this distribution.  

```{r}
ChimpNPAS=PAS_m %>% filter(species=="Chimp")
HumanNPAS=PAS_m %>% filter(species=="Human")


wilcox.test(HumanNPAS$nPAS,ChimpNPAS$nPAS ,alternative = "greater")
```
```{r}
ggplot(PAS_m,aes(x=nPAS, by=species, fill=species)) + geom_density(stat="count",alpha=.5) +  scale_fill_brewer(palette = "Dark2") + annotate(geom="text",x=7.5, y=4000, label="1 sided Wilcoxan Test p=2.55e-6 \n Human > Chimp") + labs(title="Distribution for number of PAS >= 5%", x="Number of PAS",y="Genes")
```


###Total fraction

```{r}
HumanAnno=read.table("../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndH=colnames(HumanAnno)[9:ncol(HumanAnno)]

HumanUsage=read.table("../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt", col.names = IndH) %>% select(contains("_T"))


HumanMeanTotal=as.data.frame(cbind(HumanAnno[,1:8], Human=rowMeans(HumanUsage))) 


```

```{r}
ChimpAnno=read.table("../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndC=colnames(ChimpAnno)[9:ncol(ChimpAnno)]

ChimpUsage=read.table("../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt", col.names = IndC) %>% select(contains("_T"))

ChimpMeanTotal=as.data.frame(cbind(ChimpAnno[,1:8], Chimp=rowMeans(ChimpUsage)))

```

Filter 5% and group by gene  
```{r}
BothMean_total=HumanMeanTotal %>% inner_join(ChimpMeanTotal,by=c("chr", "start", "end", "strand","loc", "disc", "PAS", "gene")) %>% filter(Chimp >=.05 | Human >=0.05) 

BothMean_total_M=melt(BothMean_total, id.vars = c("chr", "start", "end", "strand","loc", "disc", "PAS", "gene"), value.name = "Total_Usage", variable.name = "Species" )

BothMean_total_gene=BothMean_total_M %>% filter(Total_Usage>=0.05) %>% group_by(Species, gene) %>% summarise(nPASTotal=n()) 


#pos = more human, neg = more chimp 
PAS_Total_spread=BothMean_total_gene %>% spread(Species, nPASTotal, fill=0) %>% mutate(TotalDiffPAS=Human-Chimp)
```
```{r}
ggplot(PAS_Total_spread, aes(x=TotalDiffPAS)) + geom_bar(stat="count") +geom_vline(xintercept = mean(PAS_Total_spread$TotalDiffPAS),col="red") + labs(title="Difference in number of PAS at 5% Human vs Chimp \n Total Fraction", y="Genes", x="N Human PAS - N Chimp PAS")
```
```{r}
ggplot(BothMean_total_gene,aes(x=nPASTotal, by=Species, fill=Species)) + geom_density(stat="count",alpha=.5) +  scale_fill_brewer(palette = "Dark2")  + labs(title="Distribution for number of PAS >= 5%\n Total Fraction", x="Number of PAS",y="Genes")
```
###Nuclear fraction  


```{r}

HumanUsageNuclear=read.table("../Human/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt", col.names = IndH) %>% select(contains("_N"))


HumanMeanNuclear=as.data.frame(cbind(HumanAnno[,1:8], Human=rowMeans(HumanUsageNuclear))) 


```

```{r}


ChimpUsageNuclear=read.table("../Chimp/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt", col.names = IndC) %>% select(contains("_N"))

ChimpMeanNuclear=as.data.frame(cbind(ChimpAnno[,1:8], Chimp=rowMeans(ChimpUsageNuclear)))

```


Filter 5% and group by gene  
```{r}
BothMean_nuclear=HumanMeanNuclear %>% inner_join(ChimpMeanNuclear,by=c("chr", "start", "end", "strand","loc", "disc", "PAS", "gene")) %>% filter(Chimp >=.05 | Human >=0.05) 

BothMean_nuclear_M=melt(BothMean_nuclear, id.vars = c("chr", "start", "end", "strand","loc", "disc", "PAS", "gene"), value.name = "Nuclear_Usage", variable.name = "Species" )



BothMean_nuclear_gene=BothMean_nuclear_M %>% filter(Nuclear_Usage>=0.05) %>% group_by(Species, gene) %>% summarise(nPASNuclear=n()) 


#pos = more human, neg = more chimp 
PAS_Nuclear_spread=BothMean_nuclear_gene %>% spread(Species, nPASNuclear, fill=0) %>% mutate(NuclearDiffPAS=Human-Chimp)
```
```{r}
ggplot(PAS_Nuclear_spread, aes(x=NuclearDiffPAS)) + geom_bar(stat="count") +geom_vline(xintercept = mean(PAS_Nuclear_spread$NuclearDiffPAS),col="red") + labs(title="Difference in number of PAS at 5% Human vs Chimp \n Nuclear Fraction", y="Genes", x="N Human PAS - N Chimp PAS")
```


```{r}
ggplot(BothMean_nuclear_gene,aes(x=nPASNuclear, by=Species, fill=Species)) + geom_density(stat="count",alpha=.5) +  scale_fill_brewer(palette = "Dark2")  + labs(title="Distribution for number of PAS >= 5%\n Nuclear Fraction", x="Number of PAS",y="Genes")
```

###compare total and nuclear  
Compare total and nuclear. 

```{r}
mean(PAS_Total_spread$TotalDiffPAS)

mean(PAS_Nuclear_spread$NuclearDiffPAS)
```

Is the skew different in total and nuclear:

```{r}
t.test(PAS_Total_spread$TotalDiffPAS, PAS_Nuclear_spread$NuclearDiffPAS,alternative = "greater") 

```
This means skew toward more PAS in human that chimp is stronger in the total fraction.  
