---
title: "Species Specific PAS Double Filter"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this analysis I want to look at the PAS that are identified at at least 10% in one species but are not identified in the other species. I will work with avergage nuclear. I can then run the differential apa analysis with only the PAS identified in both.


```{r}
library(tidyverse)
library(ggpubr)
```


```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)

HumanSpec= PAS %>% filter(Chimp==0)
nrow(HumanSpec)
ChimpSpec= PAS %>% filter(Human==0)
nrow(ChimpSpec)
```
This is a lot better. This is down from 2500 and 4000 respecively.  

Look at the distribution of these accross the gene. 

```{r}
ggplot(HumanSpec,aes(x=loc,fill=loc)) + geom_bar(stat="count")  + labs(x="Genic location", y="Number of PAS", title="Location of Human Specific PAS") + scale_fill_brewer(palette = "Dark2")


ggplot(HumanSpec,aes(x=loc, y=Human, ,fill=loc)) + geom_boxplot()  + labs(x="Genic location", y="Human Average Usage", title="Human Specific PAS") + scale_fill_brewer(palette = "Dark2")

```



```{r}
ggplot(ChimpSpec,aes(x=loc,fill=loc)) + geom_bar(stat="count")  + labs(x="Genic location", y="Number of PAS", title="Location of Chimp Specific PAS")+ scale_fill_brewer(palette = "Dark2")


ggplot(ChimpSpec,aes(x=loc, y=Chimp, ,fill=loc)) + geom_boxplot()  + labs(x="Genic location", y="Chimp Average Usage", title="Chimp Specific PAS") + scale_fill_brewer(palette = "Dark2")
```

Next I will see if these are due to low expression. I will pull in the average normalized expression and rerun the filter.  

```{r}
nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F) %>% dplyr::select(Gene_stable_ID,Gene.name)
expr=read.table("../data/DiffExpression/NoramalizedExpression.txt",header = T,stringsAsFactors = F) %>% rename('Gene_stable_ID'=genes) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(Gene.name,Chimp, Human) %>% rename("ChimpExp"=Chimp, "HumanExp"=Human, "gene"=Gene.name)

PAS_exp=PAS %>% inner_join(expr,by="gene")
nrow(PAS_exp)
nrow(PAS_exp)/nrow(PAS)

HumanSpecExp= PAS_exp %>% filter(Chimp==0)
nrow(HumanSpecExp)
nrow(HumanSpecExp)/nrow(HumanSpec)
ChimpSpecExp= PAS_exp %>% filter(Human==0)
nrow(ChimpSpecExp)
nrow(ChimpSpecExp)/nrow(ChimpSpec)
```


We dont lose as many this way. This is evident the filter worked.  


```{r}
PAS_exp_spe=PAS_exp %>% mutate(HumanSpec=ifelse(gene %in%HumanSpecExp$gene, "yes", "no"), ChimpSpec=ifelse(gene %in% ChimpSpecExp$gene, "yes","no"))

ggplot(PAS_exp_spe,aes(x=HumanSpec,y=HumanExp)) + geom_boxplot() + stat_compare_means(method = "t.test") + labs(x="Presence of Human Specific PAS", y="Average Normalized Expression", title="Expression in Genes with Human Specific PAS")

ggplot(PAS_exp_spe,aes(x=ChimpSpec,y=ChimpExp)) + geom_boxplot() + stat_compare_means(method = "t.test") + labs(x="Presence of Chimp Specific PAS", y="Average Normalized Expression", title="Expression in Genes with Chimp Specific PAS")
```
Still a difference but not as strong.  

Ask if signal sites account for this:



```{r}
PASSS=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter_withSSTop2.txt", header = T, stringsAsFactors = F) %>% dplyr::select(PAS,HumanTopSS ,ChimpTopSS)

HumanSpec_SS=HumanSpec %>% inner_join(PASSS, by="PAS")  %>% rename("HumanSS"= HumanTopSS, "ChimpSS"=ChimpTopSS) %>% mutate(OnlyHuman=ifelse(HumanSS=="Yes" & ChimpSS=="No", "Yes", "No"), OnlyChimp=ifelse(HumanSS=="No" & ChimpSS=="Yes", "Yes", "No"))
ChimpSpec_SS=ChimpSpec %>% inner_join(PASSS, by="PAS")%>% rename("HumanSS"= HumanTopSS, "ChimpSS"=ChimpTopSS) %>%  mutate(OnlyHuman=ifelse(HumanSS=="Yes" & ChimpSS=="No", "Yes", "No"), OnlyChimp=ifelse(HumanSS =="No" & ChimpSS=="Yes", "Yes", "No"))



```
Summary:  

```{r}
HumanSpec_SS$OnlyHuman= as.factor(HumanSpec_SS$OnlyHuman)

summary(HumanSpec_SS$OnlyHuman)


HumanSpec_SS$OnlyChimp= as.factor(HumanSpec_SS$OnlyChimp)

summary(HumanSpec_SS$OnlyChimp)

HumanSpec_SS %>% filter(OnlyHuman=="Yes") 
```



Look at usage of these:

```{r}
HumanSpec_SSexp=HumanSpec_SS  %>% filter(OnlyHuman=="Yes") 

ggplot(HumanSpec_SSexp,aes(x=loc, y=Human,fill=loc))+ geom_boxplot() + scale_fill_brewer(palette = "Dark2")
```

13 sites where there is a gain of a signal site in the human specific.  I can look at them in IGV

13 of 233 


```{r}
ChimpSpec_SS$OnlyHuman= as.factor(ChimpSpec_SS$OnlyHuman)

summary(ChimpSpec_SS$OnlyHuman)


ChimpSpec_SS$OnlyChimp= as.factor(ChimpSpec_SS$OnlyChimp)

summary(ChimpSpec_SS$OnlyChimp)


ChimpSpec_SS %>% filter(OnlyChimp=="Yes")


```




11 sites where there is a gain of a signal site in the chimp specific.   


11 of 226



Look at usage of these:

```{r}
ChimpSpec_SSexp=ChimpSpec_SS  %>% filter(OnlyChimp=="Yes") 

ggplot(ChimpSpec_SSexp,aes(x=loc, y=Chimp,fill=loc))+ geom_boxplot() + scale_fill_brewer(palette = "Dark2") + labs(title="Chimp location for Chimp specific PAS with added Signal",x="Genic Location", y="Chimp Mean Usage")
```


About 5 percent in both.  


interesting chimp specific in MAN2b2, there is an AATAAA


Number with opposite relationship. 

```{r}
 HumanSpec_SS  %>% filter(OnlyChimp=="Yes")


ChimpSpec_SS %>% filter(OnlyHuman=="Yes")
```
2 in the opposite for human, 4 opposite for chimp.  
