---
title: "Species Specific PAS Double Filter"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this analysis I want to look at the PAS that are identified at at least 10% in one species but are not identified in the other species. I will work with avergage nuclear. I can then run the differential apa analysis with only the PAS identified in both.


```{r}
library(tidyverse)
library(ggpubr)
```


```{r}
PAS=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)

HumanSpec= PAS %>% filter(Chimp==0)
nrow(HumanSpec)
ChimpSpec= PAS %>% filter(Human==0)
nrow(ChimpSpec)
```
This is a lot better. This is down from 2500 and 4000 respecively.  

Look at the distribution of these accross the gene. 

```{r}
ggplot(HumanSpec,aes(x=loc,fill=loc)) + geom_bar(stat="count")  + labs(x="Genic location", y="Number of PAS", title="Location of Human Specific PAS")

ggplot(ChimpSpec,aes(x=loc,fill=loc)) + geom_bar(stat="count")  + labs(x="Genic location", y="Number of PAS", title="Location of Chimp Specific PAS")
```

Next I will see if these are due to low expression. I will pull in the average normalized expression and rerun the filter.  

```{r}
nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F) %>% dplyr::select(Gene_stable_ID,Gene.name)
expr=read.table("../data/DiffExpression/NoramalizedExpression.txt",header = T,stringsAsFactors = F) %>% rename('Gene_stable_ID'=genes) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(Gene.name,Chimp, Human) %>% rename("ChimpExp"=Chimp, "HumanExp"=Human, "gene"=Gene.name)

PAS_exp=PAS %>% inner_join(expr,by="gene")
nrow(PAS_exp)
nrow(PAS_exp)/nrow(PAS)

HumanSpecExp= PAS_exp %>% filter(Chimp==0)
nrow(HumanSpecExp)
nrow(HumanSpecExp)/nrow(HumanSpec)
ChimpSpecExp= PAS_exp %>% filter(Human==0)
nrow(ChimpSpecExp)
nrow(ChimpSpecExp)/nrow(ChimpSpec)
```


We dont lose as many this way. This is evident the filter worked.  


```{r}
PAS_exp_spe=PAS_exp %>% mutate(HumanSpec=ifelse(gene %in%HumanSpecExp$gene, "yes", "no"), ChimpSpec=ifelse(gene %in% ChimpSpecExp$gene, "yes","no"))

ggplot(PAS_exp_spe,aes(x=HumanSpec,y=HumanExp)) + geom_boxplot() + stat_compare_means(method = "t.test") + labs(x="Presence of Human Specific PAS", y="Average Normalized Expression", title="Expression in Genes with Human Specific PAS")

ggplot(PAS_exp_spe,aes(x=ChimpSpec,y=ChimpExp)) + geom_boxplot() + stat_compare_means(method = "t.test") + labs(x="Presence of Chimp Specific PAS", y="Average Normalized Expression", title="Expression in Genes with Chimp Specific PAS")
```
Still a difference but not as strong.  
