---
title: "Can signal site explain differential usage"
author: "Briana Mittleman"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis, I will ask if presence or absense of a signal site in either species can explain the diffentially used PAS. I can do overlaps and look at correlations.  

```{r}
library(workflowr)
library(RColorBrewer)
library(ggpubr)
library(tidyverse)
```

```{r}
topSS=c("AATAAA", "ATTAAA")
MetaPAS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter_withSS.txt", header = T, stringsAsFactors = F) %>% mutate(ChimpTopSS=ifelse(ChimpPAS %in% topSS, "Yes", "No"),HumanTopSS=ifelse(HumanPAS %in% topSS, "Yes", "No") )
```

Pull in the differentiall used PAS:  

```{r}
DiffUsed=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherPAS_2_Nuclear.txt", header = T, stringsAsFactors = F) %>% rename("Human_NormUse"=Human, "Chimp_NormUse"=Chimp)
```


Join this by chr, start,end, gene  

```{r}
DiffUsed_anno=DiffUsed %>% inner_join(MetaPAS,by=c("chr","start", "end","gene"))
```

Ask how many of the effect size either way are those with SS in one species  


+ upreg in human
- upred in chimp 
```{r}
DiffUsed_anno_humanup=DiffUsed_anno %>% filter(deltaPAU>0)
nrow(DiffUsed_anno_humanup)

nrow(DiffUsed_anno_humanup %>% filter(HumanTopSS=="Yes", ChimpTopSS=="No"))


DiffUsed_anno_chimpup=DiffUsed_anno %>% filter(deltaPAU<0)
nrow(DiffUsed_anno_chimpup)

nrow(DiffUsed_anno_chimpup %>% filter(HumanTopSS=="No", ChimpTopSS=="Yes"))
```


Significance:  


```{r}
humanOnlyPattern=MetaPAS %>%  filter(HumanTopSS=="Yes", ChimpTopSS=="No")

chimpOnlyPattern=MetaPAS %>%  filter(HumanTopSS=="No", ChimpTopSS=="Yes")

nrow(humanOnlyPattern)+nrow(chimpOnlyPattern)

```
Of the 44432 we see that 720 have the pattern of interest. We choose 1312 of them and 32 come out.  Look to see if it is more than expected by change 


Human up reg: 
phyper(success in sample, sucesss in possible, failure possible, sample size)

up with pattern
up reg general
no up used 
all with pattern
```{r}

#DiffUsed_anno_humanup %>% filter(HumanTopSS=="Yes", ChimpTopSS=="No")
x= nrow(DiffUsed_anno_humanup %>% filter(HumanTopSS=="Yes", ChimpTopSS=="No"))
m= nrow(DiffUsed_anno_humanup)
n=nrow(MetaPAS)-m
k=nrow(humanOnlyPattern)
N=nrow(MetaPAS)


#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)

x
```
This means that of those that have a top signal site only in human there is an enrichment for dAPA PAS.


Try  oppostie dir. 

```{r}

x= nrow(DiffUsed_anno_humanup %>% filter(HumanTopSS=="No", ChimpTopSS=="Yes"))
m= nrow(DiffUsed_anno_humanup)
n=nrow(MetaPAS)-m
k=nrow(chimpOnlyPattern)
N=nrow(MetaPAS)


#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)

x
```
So this is 32 vs 14.  

Do this for chimp.  

```{r}

x= nrow(DiffUsed_anno_chimpup %>% filter(HumanTopSS=="No", ChimpTopSS=="Yes"))
m= nrow(DiffUsed_anno_chimpup)
n=nrow(MetaPAS)-m
k=nrow(chimpOnlyPattern)
N=nrow(MetaPAS)


#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)
```
Strong enrichment here too:  

```{r}

x= nrow(DiffUsed_anno_chimpup %>% filter(HumanTopSS=="Yes", ChimpTopSS=="No"))
m= nrow(DiffUsed_anno_chimpup)
n=nrow(MetaPAS)-m
k=nrow(humanOnlyPattern)
N=nrow(MetaPAS)


#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)

x
```
No enrichment for this pattern. 

This direction has 32 vs 4.

This can be something I use. This suggests that if there are differences in signal site they are likely to lead to differential usage in the expected direction.  

Next question: is the pattern enriched in PAS called as dAPA: 
Are dAPA more likely than expected to have this pattern?

I can use a chisq test of independence.  
```{r}
DiffUsed_annotest= DiffUsed_anno %>% mutate(UpReg=ifelse(deltaPAU< 0, "Chimp", "Human"), Pattern=ifelse(ChimpTopSS=="Yes", ifelse(HumanTopSS=="Yes", "none", "Chimp"), ifelse(HumanTopSS=="Yes", "Human", "none")))  %>% select(PAS, UpReg,Pattern ) %>% filter(Pattern!="none")

DiffUsed_annotest %>% group_by(UpReg, Pattern) %>% summarise(nType=n())


toTest=DiffUsed_annotest %>% group_by(UpReg, Pattern) %>% summarise(nType=n()) %>% spread(Pattern, nType) %>% column_to_rownames("UpReg")

chisq.test(toTest)
```
This is significant 

```{r}
DiffUsed_annotest %>% group_by(UpReg, Pattern) %>% summarise(nType=n()) %>% spread(Pattern, nType)
```



Are these the strongest differences?  
```{r}
DiffUsed_annoPatternAssign= DiffUsed_anno %>% mutate(UpReg=ifelse(deltaPAU< 0, "Chimp", "Human"), Pattern=ifelse(ChimpTopSS=="Yes", ifelse(HumanTopSS=="Yes", "none", "Chimp"), ifelse(HumanTopSS=="Yes", "Human", "none")), ExpectedPattern=ifelse(Pattern!="none", "Yes", "No"))

DiffUsed_annoPatternAssign$ExpectedPattern= as.factor(DiffUsed_annoPatternAssign$ExpectedPattern)
ggplot(DiffUsed_annoPatternAssign, aes(x=ExpectedPattern, y=abs(deltaPAU), fill=ExpectedPattern))+ geom_boxplot()+ geom_jitter(alpha=.1)+stat_compare_means() + scale_fill_brewer(palette = "Set1") + labs(title="delta PAU by signal site in expected direction") + theme(legend.position = "none")

```




Plot dPAU by presence and absense:



```{r}
ggplot(DiffUsed_anno_humanup,aes(y=abs(deltaPAU), x=HumanTopSS))+ geom_boxplot()+ stat_compare_means() + labs(title="Human upregualted PAS by presence of Signal")


ggplot(DiffUsed_anno_humanup,aes(x=abs(deltaPAU), by=HumanTopSS, fill=HumanTopSS))+ geom_density(alpha=.5)+  labs(title="Human upregualted PAS by presence of Signal") + scale_fill_discrete(name="Human Signal Site Detected")


```



```{r}
ggplot(DiffUsed_anno_chimpup,aes(x=ChimpTopSS, y=abs(deltaPAU))) + geom_boxplot() + stat_compare_means() + labs(title="Chimp upregualted PAS by presence of Signal")


ggplot(DiffUsed_anno_chimpup,aes(x=abs(deltaPAU), by=ChimpTopSS, fill=ChimpTopSS))+ geom_density(alpha=.5)+  labs(title="Chimp upregualted PAS by presence of Signal") + scale_fill_discrete(name="Chimp Signal Site Detected") + annotate("text",label="Wilcoxon, p=0.035",x=.8,y=7.5)
```

It does not look like presence of a signal within the upregulated matters.  




I choose these is the original SS analysis. I used the chooseSignalSite.py it was a hierarchical model 
```{r}
MetaAllSS=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter_withSS.txt",stringsAsFactors = F,header = T)
MetaAllSSsmall=MetaAllSS %>% select(chr, start,end,PAS)
```

This has every PAS with a signal site. I will add information about if the PAS is differentially used.  

```{r}
DiffUsed_small= DiffUsed %>% select(chr,start,end, gene, deltaPAU, p.adjust) %>% inner_join(MetaAllSSsmall,by=c("chr",'start','end'))

DiffUsed_small_human= DiffUsed_small %>% filter(deltaPAU>0)
DiffUsed_small_chimp= DiffUsed_small %>% filter(deltaPAU<0)

MetaAllSS_diffUsage=MetaAllSS %>% mutate(DiffUsed=ifelse(PAS %in% DiffUsed_small$PAS, "yes","no"), HumanUp=ifelse(PAS %in% DiffUsed_small_human$PAS, "yes","no"), ChimpUp=ifelse(PAS %in% DiffUsed_small_chimp$PAS, "yes","no"))
```

Look at any vs none.  

```{r}
x= nrow(MetaAllSS_diffUsage %>% filter(HumanUp=="yes",HumanPAS!="None", ChimpPAS=="None"))
m=nrow(MetaAllSS_diffUsage %>% filter(HumanUp=="yes"))
n=nrow(MetaAllSS_diffUsage %>% filter(HumanUp!="yes"))
k= nrow(MetaAllSS_diffUsage %>% filter(HumanPAS!="None", ChimpPAS=="None"))

N=nrow(MetaAllSS_diffUsage)

x
#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)

```

Chimp:  


```{r}
x= nrow(MetaAllSS_diffUsage %>% filter(ChimpUp=="yes",HumanPAS=="None", ChimpPAS!="None"))
m=nrow(MetaAllSS_diffUsage %>% filter(ChimpUp=="yes"))
n=nrow(MetaAllSS_diffUsage %>% filter(ChimpUp!="yes"))
k= nrow(MetaAllSS_diffUsage %>% filter(HumanPAS=="None", ChimpPAS!="None"))

N=nrow(MetaAllSS_diffUsage)


#pval
phyper(x-1,m,n,k,lower.tail=F)

(x/k)/(m/N)


x
```



Examples:  

```{r}
DiffUsed_annoPatternAssign %>% filter(ExpectedPattern=="Yes") %>% filter(gene=="IRF5")
```

IRF5 is interesting because it is an apaQTL and associated with SLE 

```{r}
DiffUsed_annoPatternAssign %>% filter(ExpectedPattern=="Yes") 
```

