---
title: "Same and Different Dominant PAS with Signal Sites"
author: "Briana Mittleman"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(workflowr)
library(tidyverse)
```

I want to test if a difference in signal site distribution may explain the different dominant PAS. I am working with nuclear fraction results.  

```{r}
HumanIntronChimpUTR=read.table("../data/DominantPAS/Nuclear_HumanIntronicChimpUTR.txt",stringsAsFactors = F, header = T ) %>% rename("PAS"=HumanPAS)
SameDom=read.table("../data/DominantPAS/Nuclear_SameDom.txt", stringsAsFactors = F, header = T)
HumanUTRChimpIntron= read.table("../data/DominantPAS/Nuclear_HumanUTRChimpIntronic.txt", stringsAsFactors = F, header = T) %>% rename("PAS"=ChimpPAS)
```

Upload results for the PAS signal sites:  

```{r}
HumanSS=read.table("../data/SignalSites/HumanPresenceofSS.txt", col.names = c("PAS", "SS_human", "SS2_human", "anySS_human"), stringsAsFactors = F)
ChimpSS=read.table("../data/SignalSites/ChimpPresenceofSS.txt", stringsAsFactors = F,col.names = c("PAS", "SS_chimp", "SS2_chimp", "anySS_chimp"))
```

Join these:


I will also ask if human intronic PAS have a SS in human but not in chimp or vice versa. 

```{r}
HumanIntronChimpUTR_SS=HumanIntronChimpUTR %>% inner_join(HumanSS, by=c("PAS")) %>% inner_join(ChimpSS,by=c("PAS"))

SSinHumanOnly=HumanIntronChimpUTR_SS %>% filter(anySS_human=="yes", anySS_chimp=="no") 
nrow(SSinHumanOnly)
SSchimpOnly=HumanIntronChimpUTR_SS %>% filter(anySS_human=="no", anySS_chimp=="yes")
nrow(SSchimpOnly)
```

```{r}
prop.test(x=c(12,10), n=c(926,926), alternative="greater")
```

It does not look like this is the mechanism of action.  

Opposite direction.  I will ask if chimp intronic PAS have a SS in chimp but not in human or vice versa. 

```{r}
HumanUTRChimpIntron_SS=HumanUTRChimpIntron %>% inner_join(HumanSS, by=c("PAS")) %>% inner_join(ChimpSS,by=c("PAS"))

SSinHumanOnly_opp=HumanUTRChimpIntron_SS %>% filter(anySS_human=="yes", anySS_chimp=="no") 
nrow(SSinHumanOnly_opp)
SSchimpOnly_opp=HumanUTRChimpIntron_SS %>% filter(anySS_human=="no", anySS_chimp=="yes")
nrow(SSinHumanOnly_opp)
```

 This does not look like the mechanism this way either.  
 
** It may be interesting to look at the less frequently used signal sites.  **
 
 
I want to ask if the same dominant PAS have signal sites. It may be interesting if these intronic PAS are more likely to have a signal site than all of the intronic PAS.  
```{r}
HumanSS_2=read.table("../data/SignalSites/HumanPresenceofSS.txt", col.names = c("HumanPAS", "SS_human", "SS2_human", "anySS_human"), stringsAsFactors = F)
ChimpSS_2=read.table("../data/SignalSites/ChimpPresenceofSS.txt", stringsAsFactors = F,col.names = c("ChimpPAS", "SS_chimp", "SS2_chimp", "anySS_chimp"))

SameDom_SS=SameDom %>% inner_join(HumanSS_2, by="HumanPAS") %>% inner_join(ChimpSS_2, by="ChimpPAS") %>% mutate(bothSS=ifelse(anySS_human=="yes" & anySS_chimp =="yes", "yes", "no"), onehasSS=ifelse(anySS_human=="yes" | anySS_chimp =="yes", "yes", "no"))

#most have in both  

```
 
 
Plot

```{r}
SameDom_SS_byloc=SameDom_SS %>% group_by(HumanLoc,bothSS) %>% summarise(count=n()) %>% ungroup() %>% group_by(HumanLoc) %>% mutate(nGroup=sum(count), propSS=count/nGroup)

ggplot(SameDom_SS_byloc, aes(x=HumanLoc, y=count, by=bothSS, fill=bothSS)) +geom_bar(stat="identity",position="dodge")

ggplot(SameDom_SS_byloc, aes(x=HumanLoc, y=propSS, by=bothSS, fill=bothSS)) +geom_bar(stat="identity",position="dodge") 
```

Make the same plot with all of them. I need to pull in the metadata to do this because it has the locations. 



