---
title: "Total v Nuclear in Both species"
author: "Briana Mittleman"
date: "3/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(workflowr)
library(tidyverse)
```

I am interested in testing for PAS that are differentially used between fractions in one species but not the other. I will condition this analysis on the PAS that pass both filters in the nuclear fraction. I need to filter the fc files for PAS in the final set.  


##Prep and run LC
fixed for leafcutter are in:  

/project2/gilad/briana/Comparative_APA/Human/data/CleanLiftedPeaks4LC (filter with: /project2/gilad/briana/Comparative_APA/data/PAS_doubleFilter)
/project2/gilad/briana/Comparative_APA/Chimp/data/CleanLiftedPeaks4LC

```{r}
PAS_H=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% mutate(PASn=paste(chr,start,end,gene,sep=":") ) %>% dplyr::select(PASn, PAS) 



HumanAll=read.table("../Human/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc", header = T,stringsAsFactors = F) %>% rownames_to_column(var="PASn")  %>% inner_join(PAS_H,by="PASn")%>% dplyr::select(-PAS)

ChimpAll=read.table("../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc", header = T,stringsAsFactors = F)  %>% rownames_to_column(var="PASn") %>% inner_join(PAS_H,by="PASn") %>% dplyr::select(-PAS)

```

I need to write these out without column names. I will then add the original headers back.  


```{bash,eval=F}
mkdir ../Human/data/CleanLiftedPeaks4LC_DF
head -n 1 ../Human/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc > ../Human/data/CleanLiftedPeaks4LC_DF/Header.txt
mkdir ../Chimp/data/CleanLiftedPeaks4LC_DF
head -n 1 ../Chimp/data/CleanLiftedPeaks4LC/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc > ../Chimp/data/CleanLiftedPeaks4LC_DF/Header.txt
```

```{r}
write.table(HumanAll, "../Human/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc",quote = F, col.names = F, row.names = F)

write.table(ChimpAll, "../Chimp/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc",quote = F, col.names = F, row.names = F)
```



```{bash,eval=F}
cat ../Human/data/CleanLiftedPeaks4LC_DF/Header.txt ../Human/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Human_fixed4LC.fc > ../Human/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Human_fixed4LC_wInd.fc

cat ../Chimp/data/CleanLiftedPeaks4LC_DF/Header.txt ../Chimp/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Chimp_fixed4LC.fc > ../Chimp/data/CleanLiftedPeaks4LC_DF/ALLPAS_postLift_LocParsed_Chimp_fixed4LC_wInd.fc


mkdir ../Human/data/DiffIso_Human_DF/

python subset_diffisopheno_Huma_tvN_DF.py 1
python subset_diffisopheno_Huma_tvN_DF.py 2
python subset_diffisopheno_Huma_tvN_DF.py 3
python subset_diffisopheno_Huma_tvN_DF.py 4
python subset_diffisopheno_Huma_tvN_DF.py 5
python subset_diffisopheno_Huma_tvN_DF.py 6
python subset_diffisopheno_Huma_tvN_DF.py 7
python subset_diffisopheno_Huma_tvN_DF.py 8
python subset_diffisopheno_Huma_tvN_DF.py 9
python subset_diffisopheno_Huma_tvN_DF.py 10
python subset_diffisopheno_Huma_tvN_DF.py 11
python subset_diffisopheno_Huma_tvN_DF.py 12
python subset_diffisopheno_Huma_tvN_DF.py 13
python subset_diffisopheno_Huma_tvN_DF.py 14
python subset_diffisopheno_Huma_tvN_DF.py 15
python subset_diffisopheno_Huma_tvN_DF.py 16
python subset_diffisopheno_Huma_tvN_DF.py 18
python subset_diffisopheno_Huma_tvN_DF.py 19
python subset_diffisopheno_Huma_tvN_DF.py 20
python subset_diffisopheno_Huma_tvN_DF.py 21
python subset_diffisopheno_Huma_tvN_DF.py 22

mkdir ../Chimp/data/DiffIso_Chimp_DF/

#subset_diffisopheno_Chimp_tvN_DF.py

python subset_diffisopheno_Chimp_tvN_DF.py 1
python subset_diffisopheno_Chimp_tvN_DF.py 2
python subset_diffisopheno_Chimp_tvN_DF.py 3
python subset_diffisopheno_Chimp_tvN_DF.py 4
python subset_diffisopheno_Chimp_tvN_DF.py 5
python subset_diffisopheno_Chimp_tvN_DF.py 6
python subset_diffisopheno_Chimp_tvN_DF.py 7
python subset_diffisopheno_Chimp_tvN_DF.py 8
python subset_diffisopheno_Chimp_tvN_DF.py 9
python subset_diffisopheno_Chimp_tvN_DF.py 10
python subset_diffisopheno_Chimp_tvN_DF.py 11
python subset_diffisopheno_Chimp_tvN_DF.py 12
python subset_diffisopheno_Chimp_tvN_DF.py 13
python subset_diffisopheno_Chimp_tvN_DF.py 14
python subset_diffisopheno_Chimp_tvN_DF.py 15
python subset_diffisopheno_Chimp_tvN_DF.py 16
python subset_diffisopheno_Chimp_tvN_DF.py 18
python subset_diffisopheno_Chimp_tvN_DF.py 19
python subset_diffisopheno_Chimp_tvN_DF.py 20
python subset_diffisopheno_Chimp_tvN_DF.py 21
python subset_diffisopheno_Chimp_tvN_DF.py 22

#Samples groups:  
#../Human/data/DiffIso_Human/sample_groups.txt
#../Chimp/data/DiffIso_Chimp/sample_groups.txt 

sbatch runHumanDiffIsoDF.sh
sbatch runChimpDiffIsoDF.sh
```


##Human Results:

Results are in  ../Human/data/DiffIso_Human_DF/ 

```{bash,eval=F}
awk '{if(NR>1)print}' ../Human/data/DiffIso_Human_DF/TN_diff_isoform_chr*.txt_effect_sizes.txt > ../Human/data/DiffIso_Human_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt


awk '{if(NR>1)print}' ../Human/data/DiffIso_Human_DF/TN_diff_isoform_chr*.txt_cluster_significance.txt > ../Human/data/DiffIso_Human_DF/TN_diff_isoform_allChrom.txt_significance.txt
```

```{r}
sigH=read.table("../Human/data/DiffIso_Human_DF/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigH$p.adjust=as.numeric(as.character(sigH$p.adjust))
sigH_genes=sigH %>% filter(p.adjust<.05)
number_sigH_genes=nrow(sigH_genes)
number_sigH_genes

effectsizeH=read.table("../Human/data/DiffIso_Human_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('intron',  'logef' ,'Nuclear', 'Total','deltaPAU')) %>% filter(intron != "intron")

effectsizeH$deltaPAU=as.numeric(as.character(effectsizeH$deltaPAU))
effectsizeH$logef=as.numeric(as.character(effectsizeH$logef))

effectsizeH_deltaPAU= effectsizeH %>% mutate(SigDPAU2=ifelse(abs(deltaPAU) > .2, "Yes", "No"))  %>% separate(intron, into=c('chr','start','end','gene'), sep=":") %>% mutate(cluster=paste(chr,gene, sep=":")) %>% inner_join(sigH_genes,by="cluster")


nrow(effectsizeH_deltaPAU %>% filter(SigDPAU2=="Yes"))
```

3500 differentially used PAS.  

```{r}
plot(sort(effectsizeH$deltaPAU),main="Human Leafcutter delta PAU", ylab="Delta PAU", xlab="PAS Index")

effectsizeH_deltaPAU_col= effectsizeH_deltaPAU %>% arrange(deltaPAU) %>%  mutate(diffuse=ifelse(abs(deltaPAU)>=.2,"Yes", "No"))
effectsizeH_deltaPAU_col$color <- ifelse(effectsizeH_deltaPAU_col$diffuse=="Yes", "red", "black")

plot(sort(effectsizeH_deltaPAU_col$deltaPAU),main="Leafcutter delta PAU Human", ylab="Delta PAU", xlab="PAS Index",col=alpha(effectsizeH_deltaPAU_col$color, 0.6))
```
##Chimp results  

Results are in  ../Chimp/data/DiffIso_Chimp_DF/ 

```{bash,eval=F}
awk '{if(NR>1)print}' ../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_chr*.txt_effect_sizes.txt > ../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt


awk '{if(NR>1)print}' ../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_chr*.txt_cluster_significance.txt > ../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_allChrom.txt_significance.txt
```

```{r}
sigC=read.table("../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") 

sigC$p.adjust=as.numeric(as.character(sigC$p.adjust))
sigC_genes=sigC %>% filter(p.adjust<.05)
number_sigC_genes=nrow(sigC_genes)
number_sigC_genes

effectsizeC=read.table("../Chimp/data/DiffIso_Chimp_DF/TN_diff_isoform_allChrom.txt_effect_sizes.txt", stringsAsFactors = F, col.names=c('intron',  'logef' ,'Nuclear', 'Total','deltaPAU')) %>% filter(intron != "intron")

effectsizeC$deltaPAU=as.numeric(as.character(effectsizeC$deltaPAU))
effectsizeC$logef=as.numeric(as.character(effectsizeC$logef))

effectsizeC_deltaPAU= effectsizeC %>% mutate(SigDPAU2=ifelse(abs(deltaPAU) > .2, "Yes", "No"))  %>% separate(intron, into=c('chr','start','end','gene'), sep=":") %>% mutate(cluster=paste(chr,gene, sep=":")) %>% inner_join(sigC_genes,by="cluster")


nrow(effectsizeC_deltaPAU %>% filter(SigDPAU2=="Yes"))
``` 


```{r}
plot(sort(effectsizeC$deltaPAU),main="Chimp Leafcutter delta PAU", ylab="Delta PAU", xlab="PAS Index")
effectsizeC_deltaPAU_col= effectsizeC_deltaPAU %>% arrange(deltaPAU) %>%  mutate(diffuse=ifelse(abs(deltaPAU)>=.2,"Yes", "No"))
effectsizeC_deltaPAU_col$color <- ifelse(effectsizeC_deltaPAU_col$diffuse=="Yes", "red", "black")

plot(sort(effectsizeC_deltaPAU_col$deltaPAU),main="Leafcutter delta PAU Chimp", ylab="Delta PAU", xlab="PAS Index",col=alpha(effectsizeC_deltaPAU_col$color, 0.6))
```

1291 differentially used 


This doesnt seem right. Let me check it.  

```{r}

plot(sort(effectsizeH_deltaPAU_col$deltaPAU),main="Leafcutter delta PAU Human", ylab="Delta PAU", xlab="PAS Index",col=alpha(effectsizeH_deltaPAU_col$color, 0.6))
plot(sort(effectsizeC_deltaPAU_col$deltaPAU),main="Leafcutter delta PAU Chimp", ylab="Delta PAU", xlab="PAS Index",col=alpha(effectsizeC_deltaPAU_col$color, 0.6))

```

