---
title: "Try different quantification methods"
author: "Briana Mittleman"
date: "3/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggpubr)
library(VennDiagram)
library(Matrix)
```

I am worried about using featureCounts. It looks like there are reads that are not counted.  In this analysis, I will look at correlation between counts as well as the specific locations I have been worried about.  

First,  I will do this in human nuclear with the final PAS.  

This is the fixed strand. The 
```{r}
Meta=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt",header = T)
ChimpMeta=read.table("../data/PAS_doubleFilter/PAS_doublefilter_either_ChimpCoordChimpUsage.sort.bed", col.names = c("chr","start","end", "PAS", "score", "strand"),stringsAsFactors = F) %>% select(PAS)
HumanCountsFeatureCounts=read.table("../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc", header = T) %>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")  %>% select(-contains("_T")) %>% filter(PAS %in% Meta$PAS) 

ChimpCountsFeatureCounts=read.table("../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc", header = T) %>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")  %>% select(-contains("_T")) %>% right_join(ChimpMeta,by="PAS")
```
First use the bedtools method:  


```{bash,eval=F}
mkdir ../data/testQuant

sbatch humanMultiCov.sh

sbatch chimpMultiCov.sh
```

```{r}
HumanDFMulticov=read.table("../data/testQuant/Human_DF_PAS.txt")
```

Compare:  

```{r}
# cor(a,b)

HumanCountsFeatureCounts_mat= as.matrix(HumanCountsFeatureCounts %>% select(contains("_N")))

HumanDFMulticov_mat=as.matrix(HumanDFMulticov[,7:12])
#PAS correlation 
Human_DFcor= cor(HumanCountsFeatureCounts_mat,HumanDFMulticov_mat)
```
I care about the diagonal for the matrix.  

```{r}
Human_DFcordiag= diag(Human_DFcor)

Human_DFcordiag
```
There are the correlations for individual. 

```{r}
ChimpDFMulticov=read.table("../data/testQuant/Chimp_DF_PAS.txt")

ChimpCountsFeatureCounts_mat= as.matrix(ChimpCountsFeatureCounts %>% select(contains("_N")))

ChimpDFMulticov_mat=as.matrix(ChimpDFMulticov[,7:12])
#PAS correlation 
Chimp_DFcor= cor(ChimpCountsFeatureCounts_mat,ChimpDFMulticov_mat)

Chimp_DFcordiag= diag(Chimp_DFcor)

Chimp_DFcordiag

```


Compare species  
```{r}
DF_Multi=as.data.frame(cbind(Chimp=Chimp_DFcordiag,Human=Human_DFcordiag)) %>% gather("Species", "Correlation")


ggplot(DF_Multi,aes(x=Species, y=Correlation, fill=Species)) + geom_boxplot() + geom_jitter() + stat_compare_means() + labs(title="Double filtered Final PAS correlation \nbetween feature counts and multiCov")
```
Difference in the matricies:  

```{r}
HumanDiff= HumanCountsFeatureCounts_mat-HumanDFMulticov_mat
ChimpDiff=ChimpCountsFeatureCounts_mat-ChimpDFMulticov_mat
```
Extract the non
```{r}
nnzero(HumanDiff)
sum(HumanDiff)
HumanDiff_g=as.data.frame(HumanDiff) %>% gather("Ind", "count")

ggplot(HumanDiff_g, aes(x=Ind, y=count,col=Ind))+ geom_boxplot() + labs(y="FeatureCount - MultiCov", title="Human Difference in Counts for final PAS")
```

Look at the examples with the min values.  

```{r}
HumanDiff_rowS= rowSums(HumanDiff)
#min:  
minDiff=which(HumanDiff_rowS==min(HumanDiff_rowS))

#36964 (look at this row in the df)
HumanCountsFeatureCounts[minDiff,]
HumanDFMulticov[minDiff,]

```



```{r}
nnzero(ChimpDiff)
sum(ChimpDiff)
ChimpDiff_g=as.data.frame(ChimpDiff) %>% gather("Ind", "count")

ggplot(ChimpDiff_g, aes(x=Ind, y=count,col=Ind))+ geom_boxplot() + labs(y="FeatureCount - MultiCov", title="Chimp Difference in Counts for final PAS")
```
Look at the examples with the min values.  

```{r}
ChimpDiff_rowS= rowSums(ChimpDiff)
#min:  
minDiffChimp=which(ChimpDiff_rowS==min(ChimpDiff_rowS))

ChimpCountsFeatureCounts[minDiffChimp,]
ChimpDFMulticov[minDiffChimp,]

```


This shows that in general counts are lower in the features counts.  


Examples:  

human188040 
```{r}
Meta %>% filter(PAS=="human188040")

ChimpDFMulticov %>% filter(V4=="human188040")
ChimpCountsFeatureCounts %>% filter(PAS=="human188040")


HumanDFMulticov %>% filter(V4=="human188040")
HumanCountsFeatureCounts %>% filter(PAS=="human188040")
```
Same counts here. This is why it was not found in chimp. It seems it wouldnt pass the non zero filter.   
chimp171173
```{r}
Meta %>% filter(PAS=="chimp171173")


HumanDFMulticov %>% filter(V4=="chimp171173") 
HumanCountsFeatureCounts %>% filter(PAS=="chimp171173")
```
chimp171172

```{r}
Meta %>% filter(PAS=="chimp171172")


HumanDFMulticov %>% filter(V4=="chimp171172") 
HumanCountsFeatureCounts %>% filter(PAS=="chimp171172")
```
This is fixed. (ADI1)  

LAT  human124355  

```{r}
Meta %>% filter(PAS=="human124355")

ChimpDFMulticov %>% filter(V4=="human124355")
ChimpCountsFeatureCounts %>% filter(PAS=="human124355")


HumanDFMulticov %>% filter(V4=="human124355")
HumanCountsFeatureCounts %>% filter(PAS=="human124355")
```

Makes sense. Discovered only in human because on contig. Counting is ok.  

Compare the PAS discovered in human:  

```{r}
humanDisc=Meta %>% filter(disc=="Human")

ChimpFC_human= ChimpCountsFeatureCounts %>% filter(PAS %in% humanDisc$PAS)
ChimpMC_human= ChimpDFMulticov %>% filter(V4 %in% humanDisc$PAS)

#correlation  
ChimpFC_human_mat= as.matrix(ChimpFC_human[,13:18])
ChimpMC_human_mat=as.matrix(ChimpMC_human[,7:12])
#PAS correlation 
Chimp_DFcor_Humandisc= cor(ChimpFC_human_mat,ChimpMC_human_mat)

Chimp_DFcor_Humandiscdiag= diag(Chimp_DFcor_Humandisc)

Chimp_DFcor_Humandiscdiag
```

Is the multicov higher?  

```{r}
ChimpHumanDiscDiff= ChimpFC_human_mat-ChimpMC_human_mat

nnzero(ChimpHumanDiscDiff)

sum(ChimpHumanDiscDiff)

summary(ChimpHumanDiscDiff)


```

```{r}
ChimpHumanDiscDiff_rowS= rowSums(ChimpHumanDiscDiff)
#min:  
minDiffChimpinH=which(ChimpHumanDiscDiff_rowS==min(ChimpHumanDiscDiff_rowS))

ChimpFC_human[minDiffChimpinH,]
ChimpMC_human[minDiffChimpinH,]

```
chr 1 (.5X) RPS8 (snoRNAs near- could be problem )

Redo with different map quality  

-q 255 this will not count multimapping reads.  
```{bash,eval=F}
sbatch chimpMultiCov255.sh
sbatch humanMultiCov255.sh

```

```{r}
HumanDFMulticov255=read.table("../data/testQuant/Human_DF_PAS_255.txt")


HumanDFMulticov255_mat=as.matrix(HumanDFMulticov255[,7:12])


#compare this to the earlier one    

HumanDiffMulti=HumanDFMulticov_mat- HumanDFMulticov255_mat

HumanDiffMulti_g=as.data.frame(HumanDiffMulti) %>% gather("Ind", "Count")

ggplot(HumanDiffMulti_g,aes(x=Ind, y=Count, col=Ind)) + geom_boxplot() 


#compare to FC  

HumanDiffw225= HumanCountsFeatureCounts_mat - HumanDFMulticov255_mat
HumanDiffw225_g=as.data.frame(HumanDiffw225) %>% gather("Ind", "Count")
ggplot(HumanDiffw225_g,aes(x=Ind, y=Count, col=Ind)) + geom_boxplot() 

```
Check examples:  

```{r}
Meta %>% filter(PAS=="chimp171172")

HumanDFMulticov %>% filter(V4=="chimp171172") 
HumanDFMulticov255 %>% filter(V4=="chimp171172") 
HumanCountsFeatureCounts %>% filter(PAS=="chimp171172")
```

Ok it seems like maybe multimapping accounts for some of the differece.  


Try including the multi map reads in feature counts.  I will test the -M flag and the --primary flag to compare. The SAF file has all of the PAS.

```{bash,eval=F}
sbatch BothFCMM.sh 
```

Grep the results:  Chimp:chimp171172:chr2:3497907:3498106:+:ADI1_utr3	chr2	3497907	3498106	+	200	23	6	10	15	27	110

So this is the problem. The IGV. 

##Inclusive 
Test also on the original calls. To see if it can handle this one.  

I need to change bed to include both positive and negative. I can do this with python. 

```{bash,eval=F}
python bed2Bedbothstrand.py ../Human/data/inclusivePeaks/human_APApeaks.ALLChrom.bed ../data/testQuant/Human_ALLpeaks.Bothstrand.bed

python bed2Bedbothstrand.py ../Chimp/data/inclusivePeaks/chimp_APApeaks.ALLChrom.bed ../data/testQuant/Chimp_ALLpeaks.Bothstrand.bed

sbatch humanMultiCov_inclusive.sh
sbatch chimpMultiCovInclusive.sh


#look at difference with the -O flag  


sbatch BothFCnewOInclusive.sh
```
All counts are 0. Fix this. 



##Address multimaping problem  
it looks like the problem is multimapping. I want to get rid of PAS that multi map in one species only. To do this I want to compare the featurecounts with and without the -M flag.  

```{bash,eval=F}

sbatch BothFCMM.sh 


python fixFChead_bothfrac.py ../data/testQuant/ALLPAS_postLift_LocParsed_Human_BothFrac_Mflag ../data/testQuant/ALLPAS_postLift_LocParsed_Human_BothFrac_Mflag_fixed.fc  

```


Only look at the PAS in the double filtered for now. I want to know how many I would lose in both frac.  
```{r}
HumanCountsNoMM=read.table("../Human/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Human_fixed.fc", header = T) %>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")%>% filter(PAS %in% Meta$PAS) 

HC_noMM=rowSums(as.matrix(HumanCountsNoMM[,13:24]) )

HumanCountMM=read.table("../data/testQuant/ALLPAS_postLift_LocParsed_Human_BothFrac_Mflag", header = T)%>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":") %>% filter(PAS %in% Meta$PAS) 


HC_MM=rowSums(as.matrix(HumanCountMM[,13:24]) )

HC_diff=HC_MM-HC_noMM


```

```{r}
plot(HC_diff)
abline(120,0,col="red")
```
12 libraries. Lets say it needs to have at least 10 in each or 120
```{r}
sum(HC_diff>120)
```




```{r}

ChimpCountsNoMM=read.table("../Chimp/data/CleanLiftedPeaks_FC/ALLPAS_postLift_LocParsed_Chimp_fixed.fc", header = T)  %>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":") %>% right_join(ChimpMeta,by="PAS")

CC_noMM=rowSums(as.matrix(ChimpCountsNoMM[,13:24]) )

ChimpCountMM=read.table("../data/testQuant/ALLPAS_postLift_LocParsed_Chimp_BothFrac_Mflag", header = T) %>%  separate(Geneid, into=c("disc", "PAS", "chr2", "start2", "end2", 'strand', 'geneid'), sep=":")  %>% right_join(ChimpMeta,by="PAS")


CC_MM=rowSums(as.matrix(ChimpCountMM[,13:24]) )



CC_diff=CC_MM-CC_noMM


```


```{r}
plot(CC_diff)
abline(120,0,col="red")
```
```{r}

sum(HC_diff>120)



```

```{r}
humanMM=Human=HumanCountsNoMM[which(HC_diff>120),2]
chimpMM=ChimpCountsNoMM[which(CC_diff>120),2]

bothMM=intersect(humanMM,chimpMM)

overlap <- intersect(humanMM, chimpMM) 
H.unique <- setdiff(humanMM,chimpMM) # n 1st NOT 2nd
C.unique <- setdiff(chimpMM,humanMM)
full.list <- unique(c(humanMM,chimpMM))

```

```{r}
grid.newpage()
venn.plot <- draw.pairwise.venn(area1 = length(humanMM),
                           area2 = length(chimpMM),
                           cross.area = length(overlap),
                           c("Human", "Chimp"), scaled = TRUE,
                           fill = c("green", "blue"),
                           cex = 1.5,
                           cat.cex = 1.5,
                           cat.pos = c(320, 25),
                           cat.dist = .05) 
```

total remove if this filter: 

```{r}
length(full.list)
```

Test a range of filters:  

I will test how many PAS we would lose if we filtered at each value. I will see if there is natural cuttoff.  
```{r}
filterRange=seq(0,500000,by=100)
ChimpDiffN=c()
HumanDiffN=c()

for (i in filterRange){
  ChimpDiffN= c(ChimpDiffN, sum(CC_diff>i))
  HumanDiffN= c(HumanDiffN, sum(HC_diff>i))
}


LostPASdf=as.data.frame(cbind(Filter=filterRange, Chimp=ChimpDiffN, Human=HumanDiffN))
#%>% mutate(All=42318)
LostPASdf_g= LostPASdf %>% gather("Species","PAS_lost", -Filter)

```


Plot:  

```{r}
ggplot(LostPASdf_g, aes(x=Filter, y=log10(PAS_lost),col=Species)) + geom_point(alpha=.5) + labs(y="log10(PAS lost due to filter)", x="Filter for difference in read count due to multimapping", title="PAS lost due to multimap filter by different filter") + scale_color_brewer(palette = "Dark2")+geom_vline(xintercept = 20000) + annotate("text", label="Difference of 20,000 \ncounts across individuals", x=120000,y=3)

```

```{r}
ggplot(LostPASdf_g, aes(x=log10(Filter), y=log10(PAS_lost),col=Species)) + geom_point(alpha=.5) + labs(y="log10(PAS lost due to filter)", x="log10(Filter for difference in read count due to multimapping)", title="PAS lost due to multimap filter by different filter") + scale_color_brewer(palette = "Dark2")

```



Look at proportion.  


```{r}
LostPASdfPROP=LostPASdf_g%>% mutate(All=42318, propLeft=PAS_lost/All, PropRemain=1-propLeft)
```
```{r}
ggplot(LostPASdfPROP, aes(x=Filter, y=PropRemain,col=Species)) + geom_point(alpha=.3)
```

Zoom in.  Look at filter over 1000 
 
```{r}
LostPASdfPROP_filt=LostPASdfPROP %>% filter(Filter>1000)


ggplot(LostPASdfPROP_filt, aes(x=Filter, y=PropRemain,col=Species)) + geom_point(alpha=.3) + geom_vline(xintercept = 20000) +  annotate("text", label="Difference of 20,000 \ncounts across individuals", x=120000,y=.998) + scale_color_brewer(palette = "Dark2") + labs(x="Filter for difference in reads", y="Proportion of PAS remaining", title="Proportion of PAS remaining by species after multimap filter")

```

Use this filter to make the boxplot  



```{r}
humanMM2=Human=HumanCountsNoMM[which(HC_diff>20000),2]
chimpMM2=ChimpCountsNoMM[which(CC_diff>20000),2]

bothMM2=intersect(humanMM2,chimpMM2)
overlap2 <- intersect(humanMM2, chimpMM2) 
grid.newpage()
venn.plot <- draw.pairwise.venn(area1 = length(humanMM2),
                           area2 = length(chimpMM2),
                           cross.area = length(overlap2),
                           c("Human", "Chimp"), scaled = TRUE,
                           fill = c("orange", "green"),
                           cex = 1.5,
                           cat.cex = 1.5,
                           cat.pos = c(320, 25),
                           cat.dist = .05) 
```

expect to loose chimp171173 in human  

```{r}
HumanCountsNoMM[HumanCountsNoMM$PAS=="chimp171173",]
HC_diff[22206]

```

Should it be the proportion of reads lost?  

I need the sum for the reads with and without multimapping at each PAS.  Proportion of reads in the mulimap set 

```{r}
HumanPASMultiMap=as.data.frame(cbind(PAS=HumanCountMM$PAS,WithMultiMap=HC_MM, NoMultiMap=HC_noMM)) 
HumanPASMultiMap$WithMultiMap=as.numeric(as.character(HumanPASMultiMap$WithMultiMap))
HumanPASMultiMap$NoMultiMap=as.numeric(as.character(HumanPASMultiMap$NoMultiMap))
HumanPASMultiMapProp= HumanPASMultiMap%>% mutate(Difference=WithMultiMap-NoMultiMap, PercentLost=Difference/WithMultiMap)

```

check chimp171173 

```{r}
HumanPASMultiMapProp %>% filter(PAS=="chimp171173")
```

Plot  
Color top 10% 

```{r}
x=quantile(HumanPASMultiMapProp$PercentLost, seq(0,1, by=.05),na.rm = T)[20]

HumanPASMultiMapPropcol=HumanPASMultiMapProp %>% mutate(color=ifelse(PercentLost>=x, "red", "black")) %>% arrange(PercentLost)
```


```{r}
plot(HumanPASMultiMapPropcol$PercentLost,col=HumanPASMultiMapPropcol$color,ylab = "Percent of reads multimapping",main="Percent of reads lost to multimap filter in each human PAS",xlab="42,318 PAS Sorted")
legend("topleft", legend=c("Bottom 95%", "Top 5%"),col=c("black", "red"), pch=16,bty = 'n')
```
Look at this in chimp  


```{r}
ChimpPASMultiMap=as.data.frame(cbind(PAS=ChimpCountMM$PAS,WithMultiMap=CC_MM, NoMultiMap=CC_noMM)) 
ChimpPASMultiMap$WithMultiMap=as.numeric(as.character(ChimpPASMultiMap$WithMultiMap))
ChimpPASMultiMap$NoMultiMap=as.numeric(as.character(ChimpPASMultiMap$NoMultiMap))
ChimpPASMultiMapProp= ChimpPASMultiMap%>% mutate(Difference=WithMultiMap-NoMultiMap, PercentLost=Difference/WithMultiMap)


```

Plot  
Color top 5% 

```{r}
y=quantile(ChimpPASMultiMapProp$PercentLost, seq(0,1, by=.05),na.rm = T)[20]

ChimpPASMultiMapPropcol=ChimpPASMultiMapProp %>% mutate(color=ifelse(PercentLost>=y, "red", "black")) %>% arrange(PercentLost)
```
```{r}
plot(ChimpPASMultiMapPropcol$PercentLost,col=ChimpPASMultiMapPropcol$color,ylab = "Percent of reads multimapping",main="Percent of reads lost to multimap filter in each chimp PAS",xlab="42,318 PAS Sorted")
legend("topleft", legend=c("Bottom 95%", "Top 5%"),col=c("black", "red"), pch=16,bty = 'n')
```


Are these the same PAS :  

Look at example.  

```{r}
ChimpPASMultiMapProp %>% filter(PAS=="chimp171173")
```

So this shows that this PAS would not be lost in chimp    

```{r}
ChimpTop5=ChimpPASMultiMapPropcol %>% filter(color=="red") %>% select(PAS)
HumanTop5=HumanPASMultiMapPropcol %>% filter(color=="red") %>% select(PAS)

bothLostProp=intersect(HumanTop5$PAS,ChimpTop5$PAS)
OverlapProp <-intersect(HumanTop5$PAS, ChimpTop5$PAS) 
grid.newpage()
venn.plot <- draw.pairwise.venn(area1 = length(HumanTop5$PAS),
                           area2 = length(ChimpTop5$PAS),
                           cross.area = length(OverlapProp),
                           c("Human", "Chimp"), scaled = TRUE,
                           fill = c("darkorange", "darkgreen"),
                           cex = 1.5,
                           cat.cex = 1.5,
                           cat.pos = c(320, 25),
                           cat.dist = .05) 
```


This shows that multimapping is not equivalent between.  


Color top  2% (get up to 13% difference  )  

```{r}
y=quantile(ChimpPASMultiMapProp$PercentLost, seq(0,1, by=.01),na.rm = T)[99]

ChimpPASMultiMapPropcol2=ChimpPASMultiMapProp %>% mutate(color=ifelse(PercentLost>=y, "red", "black")) %>% arrange(PercentLost)

x=quantile(HumanPASMultiMapProp$PercentLost, seq(0,1, by=.01),na.rm = T)[99]

HumanPASMultiMapPropcol2=HumanPASMultiMapProp %>% mutate(color=ifelse(PercentLost>=y, "red", "black")) %>% arrange(PercentLost)

```



```{r}
plot(HumanPASMultiMapPropcol2$PercentLost,col=HumanPASMultiMapPropcol2$color,ylab = "Percent of reads multimapping",main="Percent of reads lost to multimap filter in each human PAS",xlab="42,318 PAS Sorted")
legend("topleft", legend=c("Bottom 98%", "Top 2%"),col=c("black", "red"), pch=16,bty = 'n')
```


```{r}
plot(ChimpPASMultiMapPropcol2$PercentLost,col=ChimpPASMultiMapPropcol2$color,ylab = "Percent of reads multimapping",main="Percent of reads lost to multimap filter in each chimp PAS",xlab="42,318 PAS Sorted")
legend("topleft", legend=c("Bottom 98%", "Top 2%"),col=c("black", "red"), pch=16,bty = 'n')
```

```{r}
ChimpTop2=ChimpPASMultiMapPropcol2 %>% filter(color=="red") %>% select(PAS)
HumanTop2=HumanPASMultiMapPropcol2 %>% filter(color=="red") %>% select(PAS)

bothLostProp=intersect(HumanTop2$PAS,ChimpTop2$PAS)
OverlapProp <-intersect(HumanTop2$PAS, ChimpTop2$PAS) 
grid.newpage()
venn.plot <- draw.pairwise.venn(area1 = length(HumanTop2$PAS),
                           area2 = length(ChimpTop2$PAS),
                           cross.area = length(OverlapProp),
                           c("Human", "Chimp"), scaled = TRUE,
                           fill = c("darkorange", "darkgreen"),
                           cex = 1.5,
                           cat.cex = 1.5,
                           cat.pos = c(320, 25),
                           cat.dist = .05) 
```

In this set there are more on the chimp side. 

Will need to choose a cutoff.  
