---
title: "Upsetter plot with all phenotypes double fiilter"
author: "Briana Mittleman"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this analysis I will use the UpSetR package to look at all of the differential gene regulation phenotype results in one plot. This should be easier to visualize than the venn diagrams.  


```{r}

library(UpSetR)
library(workflowr)
library(tidyverse)
library(RColorBrewer)
```



Input the datasets:

```{r}
#protein
Protein=read.table("../data/Khan_prot/HC_SigProtein.txt", header = T, stringsAsFactors = F)
#trans
Translation=read.table("../data/Wang_ribo/HC_SigTranslation.txt", header = T, stringsAsFactors = F)
#expression
nameID=read.table("../../genome_anotation_data/ensemble_to_genename.txt",sep="\t", header = T, stringsAsFactors = F)
DEgenes=read.table("../data/DiffExpression/DE_genes.txt", header = F,col.names = c("Gene_stable_ID"),stringsAsFactors = F) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(Gene.name)
#nuclear apa  
NucAPA=read.table("../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt",header = T,stringsAsFactors = F)

DSgenes=read.table("../data/DiffSplice_liftedJunc/orderedGeneListFixed.txt",stringsAsFactors = F, col.names = "DS")

```


Create a named list object  


```{r}
listInput_nucOnly <- list(DE=DEgenes$Gene.name, DS=DSgenes$DS, DAPA=NucAPA$gene, DT=Translation$Gene, DP=Protein$gene.symbol)


upset(fromList(listInput_nucOnly), order.by = "freq", keep.order = T,empty.intersections = "on")
```

Add colors for certain queries:

```{r}
#upset(movies, queries = list(list(query = intersects, params = list("Drama", 
#    "Comedy", "Action"), color = "orange", active = T), list(query = intersects, 
 #   params = list("Drama"), color = "red", active = F), list(query = intersects, 
 #   params = list("Action", "Drama"), active = T)))


upset(fromList(listInput_nucOnly), queries = list(list(query=intersects, params=list("DAPA", "DT", "DP"), color="red", active=T,query.name="APA,Ribo, Protein"),list(query=intersects, params=list("DE", "DT", "DP"), color="orange", active=T, query.name="Expression,Ribo, Protein"), list(query=intersects, params=list("DS", "DT", "DP"), color="green", active=T,query.name="Splicing ,Ribo, Protein"),list(query=intersects, params=list("DAPA", "DT"), color="blue", active=T, query.name="APA,Ribo") ,list(query=intersects, params=list("DAPA", "DP"), color="purple", active=T, query.name="APA, Protein")), order.by = "freq", query.legend = "bottom")

```
Remove splicing:  

```{r}

brewer.pal(6, "RdYlBu")
listInput_nosplice <- list(DE=DEgenes$Gene.name, DAPA=NucAPA$gene, DT=Translation$Gene, DP=Protein$gene.symbol)


upset(fromList(listInput_nosplice), queries = list(list(query=intersects, params=list("DAPA", "DT", "DP"), color="#D7191C", active=T,query.name="APA, Ribo, Protein"),list(query=intersects, params=list("DE", "DT", "DP"), color="#FDAE61", active=T, query.name="Expression,Ribo, Protein"),list(query=intersects, params=list("DAPA", "DT"), color="#FEE090", active=T, query.name="APA,Ribo") ,list(query=intersects, params=list("DAPA", "DP"), color="#ABD9E9", active=T, query.name="APA, Protein"),list(query=intersects, params=list("DAPA", "DE"), color="#2C7BB6", active=T, query.name="APA, Expression")), order.by = "freq", query.legend = "bottom")
```


Will use hypergeometric to test overlaps.  

https://lauren-blake.github.io/Regulatory_Evol/analysis/Tissue_specific_overlap.html#human-chimpanzee-rhesus-macaque-tissue-specific-overlap


phyper(success in sample, sucesss in possible, failure possible, sample size)

Success is the overlap, de, not DE, sample size is the apagenes tested in DE 

I need the full lists for this. Not just the significant ones.  


I will start with expression and apa. I need all of the genes tested for both.   

```{r}

DEgenestested=read.table("../data/DiffExpression/DE_Testedgenes.txt", header = F,col.names = c("Gene_stable_ID"),stringsAsFactors = F) %>% inner_join(nameID, by="Gene_stable_ID") %>% dplyr::select(Gene.name)
apaTested=read.table("../data/DiffIso_Nuclear_DF/TN_diff_isoform_allChrom.txt_significance.txt",sep="\t" ,col.names = c('status','loglr','df','p','cluster','p.adjust'),stringsAsFactors = F) %>% filter(status=="Success") %>% separate(cluster, into=c("chr", "Gene.name"),sep=":")

DEtestedandAPA=NucAPA %>%rename("Gene.name"=gene) %>% inner_join(DEgenestested, by="Gene.name") %>% nrow()
DeandAPA=NucAPA %>%rename("Gene.name"=gene) %>%  inner_join(DEgenes, by="Gene.name")%>% nrow()
NotDe= nrow(DEgenestested)-nrow(DEgenes)

x=DeandAPA
m=nrow(DEgenes)
n=NotDe
k=DEtestedandAPA


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))


#actual:
DeandAPA

#pval
phyper(DeandAPA, nrow(DEgenes), NotDe, DEtestedandAPA,lower.tail=F)
```

Translation:  
Success is the overlap, T, not T, sample size is the apagenes tested in DE 

```{r}
TranslationTested=read.table("../data/Wang_ribo/HC_AllTestedTranslation.txt",header = T,stringsAsFactors = F) %>% rename("gene"=Gene)
TranslationNotTE= TranslationTested %>% filter(HvC.FDR>=.05)
#actual overlap
TeandAPA=Translation %>% rename("gene"=Gene) %>% inner_join(NucAPA,by="gene")
TranslationTestedandAPA= TranslationTested %>% inner_join(NucAPA,by="gene")


x=nrow(TeandAPA)
m= nrow(Translation)
n=nrow(TranslationNotTE)
k=nrow(TranslationTestedandAPA)


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
nrow(TeandAPA)

#pval
phyper(nrow(TeandAPA), nrow(Translation), nrow(TranslationNotTE), nrow(TranslationTestedandAPA),lower.tail=F)
```

Protein  

Success is the overlap, dp, not no dp, sample size is the apagenes tested in dp  

```{r}
ProtTested=read.table("../data/Khan_prot/HC_AlltestedProtein.txt",header = T,stringsAsFactors = F) %>% rename("gene"=gene.symbol)
ProtNotPE= ProtTested %>% filter(HC.qvalues.protein>=.05)
#actual overlap
PEandAPA=Protein %>% rename("gene"=gene.symbol) %>% inner_join(NucAPA,by="gene")
ProtTestedandAPA= ProtTested %>% inner_join(NucAPA,by="gene")


x=nrow(PEandAPA)
m= nrow(Protein)
n=nrow(ProtNotPE)
k=nrow(ProtTestedandAPA)


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
nrow(PEandAPA)

#pval
phyper(nrow(PEandAPA), nrow(Protein), nrow(ProtNotPE), nrow(ProtTestedandAPA),lower.tail=F)
```


Use lauren's code for the 3 set  

Human-chimpanzee-rhesus macaque tissue specific overlap:  

m is the overlap of human+chimpanzee tissue-specific genes
n is the Total genes - (overlap of human+chimpanzee tissue-specific genes)
x/q is the overlap between human, chimpanzee, rhesus macaque tissue-specific genes
k is the total rhesus tissue-specific genes


phyper(x=overlap all pheno, m=overlap 2 (not apa) , n= total - overlap of 2 (not apa), k=total for apa)

```{r}
ApaProtTrans=PEandAPA  %>% inner_join(TeandAPA, by="gene")

PeandTE=Translation %>% inner_join(Protein) %>% rename("gene"=Gene)

#not dp and dt is all tested in both - pe and te set
PeandTEtested= ProtTested %>% full_join(TranslationTested, by="gene") %>% anti_join(PeandTE,by="gene")
  
apaTestedinboth= ProtTested %>% full_join(TranslationTested, by="gene") %>% inner_join(NucAPA, by="gene")

x=nrow(ApaProtTrans)
m= nrow(PeandTE)
n=nrow(PeandTEtested)
k=nrow(apaTestedinboth)


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
nrow(ApaProtTrans)

#pval


phyper(x,m,n,k,lower.tail=F)
```

Expression te and pe  


Success is the overlap, dp and dt, not dp and dt, sample size is the egenes tested in both dp and dt 


```{r}
Protein_g = Protein %>% rename("gene"=gene.symbol)
translation_g= Translation%>% rename("gene"=Gene)
DEgenes_g= DEgenes %>% rename("gene"=Gene.name)

EProtTrans=DEgenes %>% rename("gene"=Gene.name) %>% inner_join(Protein_g, by="gene") %>% inner_join(translation_g, by="gene")

PeandTE=Translation %>% inner_join(Protein, by = "ENSG") %>% rename("gene"=Gene)

#not dp and dt is all tested in both - pe and te set
PeandTEtested= ProtTested %>% full_join(TranslationTested, by="gene") %>% anti_join(PeandTE,by="gene")
  
expressioninboth= ProtTested %>% full_join(TranslationTested, by="gene") %>% inner_join(DEgenes_g, by="gene")

x=nrow(EProtTrans)
m= nrow(PeandTE)
n=nrow(PeandTEtested)
k=nrow(expressioninboth)


#expected
which(grepl(max(dhyper(1:x, m, n, k)), dhyper(1:x, m, n, k)))

#actual:
nrow(EProtTrans)

#pval
phyper(x,m,n,k,lower.tail=F)
```


I can make a table for this- I'll have the set, the actual, the expected, the pvalue. I can also add these numbers onto the figure above.  


Splicing results:  

```{r}
listInput_Splice <- list(DE=DEgenes$Gene.name, DS=DSgenes$DS, DT=Translation$Gene, DP=Protein$gene.symbol)
upset(fromList(listInput_Splice), order.by = "freq", keep.order = T,empty.intersections = "on")
```

```{r}
DeandAPAList=NucAPA %>%rename("Gene.name"=gene) %>%  inner_join(DEgenes, by="Gene.name") %>% rename("gene"=Gene.name)

write.table(DeandAPAList,"../data/DiffIso_Nuclear_DF/GeneswithDEanddAPA.txt", col.names = T,row.names = F,quote = F)
```

