---
title: "Visualize Usage percentages double filtered"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(workflowr)
library(tidyverse)
```

I want to write a loop that will create boxplots for all of the PAS in each gene. I need a script that takes one gene and then outputs the plot.  

I want PAS:Location as the x axis. Usage as the y axis. I will seperate the plots by species.  


I need a long dataframe with species, individual, PASid, gene, usage.  


Pull in the usage numbers: 

../data/Pheno_5perc_nuclear/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc_Nuclear.txt.gz

I need to make these numeric.  


```{bash,eval=F}
mkdir ../data/files4viz_nuclear_DF/
```


Vector with humans:

```{r}
humans=c('NA18498_N', 'NA18502_N' ,'NA18504_N', 'NA18510_N', 'NA18523_N')

chimps=c( 'NA18358_N', 'NA3622_N' ,'NA3659_N', 'NA4973_N', 'NApt30_N', 'NApt91_N')
```

LOC100288069_+_utr3-Both-human23

LOC107984841_+_utr3-Chimp-chimp242477
utr3-Chimp-chimp242477


I will double filter here as well.  
```{r,eval=F}
NuclearUsage=read.table("../data/Pheno_5perc_nuclear/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc_Nuclear.txt.gz", header = T)

PASMeta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% dplyr::rename("PASnum" = PAS)

Nuclear_gather=NuclearUsage %>% gather(key="Ind", value="Frac", -chrom) %>% separate(Frac, into=c("num", "denom"), sep="/") %>% mutate(numUsage=as.integer(num)/as.integer(denom)) %>% replace_na(list(numUsage=0)) %>% separate(Ind, into=c("sample", "fraction"), by="_") %>% mutate(species=ifelse(sample %in% humans, "Human","Chimp")) %>% separate(chrom, into=c("chr","start","end","PASid"), sep=":")%>% separate(PASid, into=c("Gene", "strand", "PASinf"), sep="_")%>% separate(PASinf, into=c("loc", "disc", "PASnum"), sep="-") 

%>% filter(PASnum %in% PASMeta$PASnum) 

%>% mutate(PAS=paste(PASnum,loc,sep="_")) %>% dplyr::select(chr,start,end, strand, PAS, species,Gene, sample,numUsage)
```

```{r}
HumanAnno=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndH=colnames(HumanAnno)[9:ncol(HumanAnno)]

HumanUsage=read.table("../Human/data/phenotype/ALLPAS_postLift_LocParsed_Human_Pheno_countOnlyNumeric.txt", col.names = IndH)
 

HumanMean=as.data.frame(cbind(HumanAnno[,1:8], Human=rowMeans(HumanUsage)))

HumanUsage_anno=as.data.frame(cbind(HumanAnno[,1:8],HumanUsage ))


ChimpAnno=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno.txt", header = T, stringsAsFactors = F) %>% tidyr::separate(chrom, sep = ":", into = c("chr", "start", "end", "id")) %>% tidyr::separate(id, sep="_", into=c("gene", "strand", "peak"))  %>% separate(peak,into=c("loc", "disc","PAS"), sep="-")
IndC=colnames(ChimpAnno)[9:ncol(ChimpAnno)]

ChimpUsage=read.table("../Chimp/data/phenotype/ALLPAS_postLift_LocParsed_Chimp_Pheno_countOnlyNumeric.txt", col.names = IndC)

ChimpMean=as.data.frame(cbind(ChimpAnno[,1:8], Chimp=rowMeans(ChimpUsage)))

ChimpUsage_anno=as.data.frame(cbind(ChimpAnno[,1:8],ChimpUsage ))

Meta=read.table("../data/PAS_doubleFilter/PAS_5perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F)

Nuclear_gather=ChimpUsage_anno %>% full_join(HumanUsage_anno,by=c("chr", "start", "end", "gene","strand","loc", "disc", "PAS")) %>% filter(PAS %in% Meta$PAS)  %>% select(-contains("_T")) %>% mutate(PAS=paste(PAS,loc,sep="_")) %>% select(-loc, -disc) %>% gather("Sample", "numUsage", -chr,-start,-end,-gene,-strand,-PAS ) %>%mutate(species=ifelse(Sample %in% humans, "Human","Chimp"))%>% dplyr::select(chr,start,end, strand, PAS, species,gene, Sample,numUsage)
```



example gene:  
DFFB
```{r}
Nuclear_gather_DFFB=Nuclear_gather %>% filter(gene=="DFFB")

Nuclear_gather_DFFB$start=as.integer(Nuclear_gather_DFFB$start)


ggplot(Nuclear_gather_DFFB,aes(x=(reorder(PAS, start)), y=numUsage, by=species, fill=species))+ geom_boxplot(notch =F, width=.5) + scale_fill_brewer(palette = "Dark2")+theme(axis.text.x = element_text(angle = 90), legend.position = "top",panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) + labs(y="Usage", title="Usage by species for DFFB",x="PAS") 
```


```{r}
write.table(Nuclear_gather,"../data/files4viz_nuclear_DF/NuclearPASUsage.txt", col.names = T, row.names = F, quote = F)
```



I will put this in an R script that takes a gene.  


test on DFFB
```{bash,eval=F}
Rscript PlotNuclearUsagebySpecies_DF.R -g DFFB
```

This works, now I can make it for every gene that was called as DAPA. I will make a list of all of the genes.  

genes are in ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt

```{bash,eval=F}
sed '1d' ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt > ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear_noHead.txt
sbatch makeNuclearDapaplots_DF.sh
```


Run for DIC genes as well



```{bash,eval=F}
PlotNuclearUsagebySpecies_DF_4DIC.R

mkdir ../data/DIC_Viz
```


```{r}
DiC5=read.table("../data/IndInfoContent/SimpsonMedianSignificance.txt",header = T, stringsAsFactors = F) %>% filter(sIC=="Yes") %>% select(gene)

write.table(DiC5,"../data/DIC_Viz/DiCGenes_5fdr.txt", col.names = F, row.names = F,quote = F)
```

```{bash,eval=F}
sbatch makedICPlots_DF.sh
```

