---
title: "Visualize Usage percentages double filtered"
author: "Briana Mittleman"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(workflowr)
library(tidyverse)
```

I want to write a loop that will create boxplots for all of the PAS in each gene. I need a script that takes one gene and then outputs the plot.  

I want PAS:Location as the x axis. Usage as the y axis. I will seperate the plots by species.  


I need a long dataframe with species, individual, PASid, gene, usage.  


Pull in the usage numbers: 

../data/Pheno_5perc_nuclear/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc_Nuclear.txt.gz

I need to make these numeric.  


```{bash,eval=F}
mkdir ../data/files4viz_nuclear_DF/
```


Vector with humans:

```{r}
humans=c('NA18498', 'NA18499', 'NA18502' ,'NA18504', 'NA18510', 'NA18523')

chimps=c( 'NA18358', 'NA3622' ,'NA3659', 'NA4973', 'NApt30', 'NApt91')
```

LOC100288069_+_utr3-Both-human23

LOC107984841_+_utr3-Chimp-chimp242477
utr3-Chimp-chimp242477


I will double filter here as well.  
```{r}
NuclearUsage=read.table("../data/Pheno_5perc_nuclear/ALLPAS_postLift_LocParsed_bothSpecies_pheno_5perc_Nuclear.txt.gz", header = T)

PASMeta=read.table("../data/PAS_doubleFilter/PAS_10perc_either_HumanCoord_BothUsage_meta_doubleFilter.txt", header = T, stringsAsFactors = F) %>% dplyr::rename("PASnum" = PAS)

Nuclear_gather=NuclearUsage %>% gather(key="Ind", value="Frac", -chrom) %>% separate(Frac, into=c("num", "denom"), sep="/") %>% mutate(numUsage=as.integer(num)/as.integer(denom)) %>% replace_na(list(numUsage=0)) %>% separate(Ind, into=c("sample", "fraction"), by="_") %>% mutate(species=ifelse(sample %in% humans, "Human","Chimp")) %>% separate(chrom, into=c("chr","start","end","PASid"), sep=":")%>% separate(PASid, into=c("Gene", "strand", "PASinf"), sep="_") %>% separate(PASinf, into=c("loc", "disc", "PASnum"), sep="-") %>% filter(PASnum %in% PASMeta$PASnum) %>% mutate(PAS=paste(PASnum,loc,sep="_")) %>% dplyr::select(chr,start,end, strand, PAS, species,Gene, sample,numUsage)
```



example gene:  
DFFB
```{r}
Nuclear_gather_DFFB=Nuclear_gather %>% filter(Gene=="DFFB")

Nuclear_gather_DFFB$start=as.integer(Nuclear_gather_DFFB$start)


ggplot(Nuclear_gather_DFFB,aes(x=(reorder(PAS, start)), y=numUsage, by=species, fill=species))+ geom_boxplot() + scale_fill_brewer(palette = "Dark2")+theme(axis.text.x = element_text(angle = 90)) + labs(y="Usage", title="Usage by species for DFFB",x="PAS")
```


```{r}
write.table(Nuclear_gather,"../data/files4viz_nuclear_DF/NuclearPASUsage.txt", col.names = T, row.names = F, quote = F)
```



I will put this in an R script that takes a gene.  


test on DFFB
```{bash,eval=F}
Rscript PlotNuclearUsagebySpecies_DF.R -g DFFB
```

This works, now I can make it for every gene that was called as DAPA. I will make a list of all of the genes.  

genes are in ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt

```{bash,eval=F}
sed '1d' ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear.txt > ../data/DiffIso_Nuclear_DF/SignifianceEitherGENES_Nuclear_noHead.txt

sbatch makeNuclearDapaplots_DF.sh
```

